---
title: 'OHI 2020: Mariculture'
author: "*Compiled on `r date()` by `r Sys.info()['user']`*"
output: 
  html_document:
    toc: true
    number_sections: true
    theme: cerulean
    highlight: haddock
    includes: 
      in_header: '../../../workflow/templates/ohi_hdr.html'
  pdf_document:
    toc: true
---

# Summary
This analysis converts FAO mariculture data into data used to calculate the OHI global mariculture status score. This also calculates the genetic escapee from mariculture pressure data.

# Updates from previous assessment

* New year of FAO mariculture yield data for 2018.
* 

***

# Data Source 

## Production data

**Reference**:      
http://www.fao.org/fishery/statistics/software/fishstatj/en#downlApp
Release date: March 2019
FAO Global Aquaculture Production Quantity 1950_2018
FAO metadata found [here](http://www.fao.org/fishery/statistics/global-aquaculture-production/en)

**Downloaded**: May 11, 2020

**Description**:  Quantity (tonnes) of mariculture for each country, species, year.

**Time range**: 1950-2018 

## Trujillo sustainability data

**Reference**:      
Mariculture Sustainability Index (Trujillo 2008)

Trujillo, Pablo. 2008. “Using a Mariculture Sustainability Index to Rank Countries’ Performance.” In Fisheries Centre Research Reports, edited by Jackie Alder and Daniel Pauly. University of British Columbia, Vancouver, Canada: Fisheries Centre Research Reports. https://circle.ubc.ca/handle/2429/40933. 

Original data table is located in the Trujillo 2008 paper saved in the OHI Zotero database. Report starts on page 28 of the link provided in the reference. 

**Description**:  

Original MSI rescaled from 1-10 to 0-1 for mariculture taxa groups.

## Seafood Watch sustainability data

**Reference**: 
https://www.seafoodwatch.org/-/m/sfw/pdf/whats%20new/complete%20recommendation%20list.pdf
Release date: August 3, 2020

**Downloaded**: July 22, 2020

**Description**: Monterey Bay Aquarium Seafood Watch aquaculture recommendations. Sustainability scored from 0-10.

***
  
# Methods

```{r setup, message=FALSE, warning=FALSE}

getwd()
knitr::opts_chunk$set(eval=FALSE)
  
## Load libraries, set directories
library(ohicore)  #devtools::install_github('ohi-science/ohicore@dev')
library(dplyr)
library(stringr)
library(tidyr)
library(tidyverse)

## Load FAO-specific user-defined functions
source('mar_fxs.R') # functions specific to mariculture dealing with compound countries
source('../../../workflow/R/fao_fxn.R') # function for cleaning FAO files
source('../../../workflow/R/common.R') # directory locations

```

# Import Raw Data: FAO Mariculture data

Mariculture production in tonnes.

```{r import-data}

mar <- read.csv(file.path(dir_M, 'git-annex/globalprep/_raw_data/FAO_mariculture/d2020/FAO_GlobalAquacultureProduction_Quantity_1950_2018.csv'), check.names=FALSE, stringsAsFactors=FALSE) ; head(mar) 

```

# Wrangle: 

## Tidy mariculture data

Filter freshwater mariculture, make long format, and clean FAO codes.

```{r tidy-data}

mar <- mar %>%
  rename(country = `Country (Country)`,
         FAO_name = `ASFIS species (ASFIS species)`, 
         fao = `FAO major fishing area (FAO major fishing area)`, 
         environment = `Environment (Environment)`)
table(mar$environment)  

## Include only marine environments
mar <- mar %>%
filter(environment %in% c("Brackishwater", "Marine"))  

## Convert to long format and clean FAO codes:
## For some reason, I can't provide the data range in gather programatically!
mar <- mar %>%
  select(-Unit) 

mar <- mar %>%
  gather(key="year", value="value", num_range("",1950:2018)) %>%
    fao_clean_data() 

```

## Update species names

Update species name in the `raw/species_list.csv` file with names in the `mar` dataset (exclude non-food species). I simplified the species_list.  I cut the "species" name columns because it wasn't clear what this was trying to accomplish and created potential error.  

```{r update-species}

## Commented out 'read.csv' lines are different versions of the dataset, representing the three different seaweed exclusion methods (exclude some seaweed, all seaweed, all seaweed nei). Use at the end for data checking. The first one is the one which is saved to the "output" folder. 
mar_sp <- read.csv('raw/species_list.csv', stringsAsFactors=FALSE) %>%
  select(FAO_name, exclude, alias, Taxon_code)
#mar_sp <- read.csv('raw/species_list.csv', stringsAsFactors=FALSE) %>%
#  select(FAO_name, exclude_no_seaweed, alias, Taxon_code)
#mar_sp <- read.csv('raw/species_list.csv', stringsAsFactors=FALSE) %>%
#  select(FAO_name, exclude_no_nei, alias, Taxon_code)

new.spp <- setdiff(mar$FAO_name, mar_sp$FAO_name)
new.spp # Check: if dim has 0 rows it means all match
## If there is a list of species, hand check species_list.csv to see whether to keep (exclude seaweeds and species harvested for ornamental/medicinal), check if synonyms match Trujillo names. 
## This is outlined in issue #8 and issue #81 for reference. You need to add the species output from this to species_list.csv and fill out the corresponding information. I.e. add FAO_name, exclude (whether to include in assessment or not. If they are only harvested for medicinal purposes, do not include), alias, and Taxon_code to species_list.csv.

## REMOVE SPECIES not relevant to mariculture goal (i.e., non-food species)
mar <- mar %>% left_join(mar_sp, by="FAO_name") 

## Filters out species that should be excluded from the MAR subgoal, depending on which of the 3 scenarios was read in. Searches for column name "exclude", "exclude_no_seaweed", or "exclude_no_nei"
if(sum(str_detect(names(mar_sp), "exclude$"))==1) {
  
 mar <- mar %>% filter(exclude==0)

 } else if (sum(str_detect(names(mar_sp), "exclude_no_seaweed"))==1) {
  
   mar <- mar %>% filter(exclude_no_seaweed==0)

   } else if (sum(str_detect(names(mar_sp), "exclude_no_nei"))==1) {
  
     mar <- mar %>% filter(exclude_no_nei==0)

     }
 
## Change names using species alias or FAO species name (global changes)
mar$species <- ifelse(!is.na(mar$alias), mar$alias, mar$FAO_name) 

## Sum production values for each group to account for duplicate rows after name change (remove NA values)
mar <- mar %>%
  filter(!is.na(value)) %>%
  group_by(country, fao, environment, species, year, Taxon_code) %>% 
    summarize(value = sum(value)) %>% 
  ungroup()

## Eliminate country-species data with zero production throughout the time-series (1950-recent)
mar <- mar %>%
  group_by(country, species) %>%
  mutate(total_value = sum(value)) %>%
  filter(total_value > 0) %>%
  select(-total_value) %>%
  ungroup()

```

## Convert country names to OHI regions

```{r convert-regions}

## Divide mariculture from countries that we report as separate regions (assume equal production in all regions)
  # Netherlands Antilles: Conch restoration among Aruba, Bonaire, Curacao
  # Channel Islands: Jersey and Guernsey
  # Bonaire/S.Eustatius/Saba
  # Yugoslavia SFR: no longer a country after 1992

mar <- mar %>%
  mutate(country = ifelse(country=="Réunion", "Reunion", country)) %>%  # This one is hard to get right; v2020: last year it was "R\xe9union", but this year it was "Réunion"
  mar_split()  # function in mar_fxs.R

mar_rgn <- name_2_rgn(df_in = mar, 
                       fld_name='country', 
                       flds_unique=c('species', 'fao', 'environment', 'Taxon_code', 'year')) 


## Sum values of regions with multiple subregions
mar_rgn <- mar_rgn %>%
  group_by(fao, environment, species, year, Taxon_code, rgn_id) %>% 
  summarize(value = sum(value)) %>%
  ungroup()

```

**Take a look at the tidied data for a single year and region**

```{r see-tidy-data}

data.frame(filter(mar_rgn, rgn_id==130) %>%
  filter(year==2013) %>%
  arrange(species))

```

For some regions, a specific species can be altered so that it matches more general Trujillo sustainability data.  In this case, I don't want the name changes to be global because some regions may have more specific species data. 

(Will explore this in the future, but will not implement this year).

```{r explore-Trujillo-changes}

# ## based on looking at the list, make a few name changes to match the regions Trujillo data
# # Chile name modification
# mar_rgn$species[mar_rgn$rgn_id==224 & mar_rgn$species == "Red abalone"] <- "Abalones nei"
# mar_rgn$species[mar_rgn$rgn_id==224 & mar_rgn$species == "Japanese abalone"] <- "Abalones nei"
# 
# # China
# mar_rgn$species[mar_rgn$rgn_id==209 & mar_rgn$species == "Areolate grouper"] <- "Groupers nei"
# mar_rgn$species[mar_rgn$rgn_id==209 & mar_rgn$species == "Greasy grouper"] <- "Groupers nei"
# mar_rgn$species[mar_rgn$rgn_id==209 & mar_rgn$species == "Hong Kong grouper"] <- "Groupers nei"
# mar_rgn$species[mar_rgn$rgn_id==209 & mar_rgn$species == "Orange-spotted grouper"] <- "Groupers nei"
# 
# # Honduras
# mar_rgn$species[mar_rgn$rgn_id==133 & mar_rgn$species == "Whiteleg shrimp"] <- "Penaeus shrimps nei"
# 
# # Italy
# mar_rgn$species[mar_rgn$rgn_id==84 & mar_rgn$species == "Pacific cupped oyster"] <- "Cupped oysters nei"
# 
# # New Zealand
# mar_rgn$species[mar_rgn$rgn_id==162 & mar_rgn$species == "Rainbow abalone"] <- "Abalones nei"
# 
# # Pakistan
# mar_rgn$species[mar_rgn$rgn_id==53 & mar_rgn$species == "Penaeus shrimps nei"] <- "Marine crustaceans nei"
# 
# # Philippines
# mar_rgn$species[mar_rgn$rgn_id==15 & mar_rgn$species == "Whiteleg shrimp"] <- "Penaeus shrimps nei"
# 
# # Portugal
# mar_rgn$species[mar_rgn$rgn_id==183 & mar_rgn$species == "Golden carpet shell"] <- "Marine molluscs nei"
# mar_rgn$species[mar_rgn$rgn_id==183 & mar_rgn$species == "Peppery furrow"] <- "Marine molluscs nei"
# mar_rgn$species[mar_rgn$rgn_id==183 & mar_rgn$species == "Solen razor clams nei"] <- "Razor clams nei"
# mar_rgn$species[mar_rgn$rgn_id==183 & mar_rgn$species == "Atlantic bluefin tuna"] <- "Marine fishes nei"
# mar_rgn$species[mar_rgn$rgn_id==183 & mar_rgn$species == "Meagre"] <- "Marine fishes nei"
# mar_rgn$species[mar_rgn$rgn_id==183 & mar_rgn$species == "Seabasses nei"] <- "Marine fishes nei"
# mar_rgn$species[mar_rgn$rgn_id==183 & mar_rgn$species == "Soles nei"] <- "Marine fishes nei"
# mar_rgn$species[mar_rgn$rgn_id==183 & mar_rgn$species == "White seabream"] <- "Marine fishes nei"
# 
# # Spain
# mar_rgn$species[mar_rgn$rgn_id==182 & mar_rgn$species == "Atlantic bluefin tuna"] <- "Tuna-like fishes nei"
# 
# # Turkey
# mar_rgn$species[mar_rgn$rgn_id==76 & mar_rgn$species == "Atlantic bluefin tuna"] <- "Tuna-like fishes nei"
# mar_rgn$species[mar_rgn$rgn_id==76 & mar_rgn$species == "European seabass"] <- "Seabasses nei"
# 
# ### sum values of regions with multiple subregions
# mar_rgn <- mar_rgn %>%
#   group_by(fao, environment, species, year, Taxon_code, rgn_id) %>%
#   summarize(value = sum(value)) %>%
#   ungroup()

```

# Gapfilling

## Fill in missing years after first year of harvest data with 0 values

For example: Production of blue shrimp in Maine starts in 1983 -- don't include years before that.

Checked to make sure that there weren't instances in which it made more sense to carry the previous year's data forward as a method of gapfilling. This didn't seem to be the case.

```{r gapfill-missing-0}

## Spread mar_rgn to create a value for every year-species-region (if missing it will be given a NA)
mar_rgn_spread <- spread(mar_rgn, year, value)
dim(mar_rgn_spread)

## Turn data frame back into long format
mar_rgn_gf <- gather(mar_rgn_spread, "year", "value", num_range("",1950:2018)) %>%
  arrange(rgn_id, species, year, Taxon_code, fao, environment) 

## NA values are converted to zero.
mar_rgn_gf <- mar_rgn_gf %>%
  mutate(year = as.numeric(as.character(year))) %>%
  mutate(value_w_0 = ifelse(is.na(value), 0, value)) %>%
  group_by(fao, environment, species, Taxon_code, rgn_id) %>% 
  mutate(cum_value = cumsum(value_w_0)) %>%
  ungroup() %>%
  filter(cum_value > 0) %>% # Eliminates years before mariculture began
  mutate(gap_0_fill = ifelse(is.na(value), "NA_to_zero", "0")) %>% # Record gapfill
  mutate(value = ifelse(is.na(value), 0, value)) %>% # Finally, convert all NAs in original column to 0
  select(-cum_value, -value_w_0)

```

See how may NA values were converted to 0

```{r table-NA-0}

table(mar_rgn_gf$gap_0_fill)
## 3790 of these out of 27810+3790 cases had NA values converted to 0 - 2018 assessment
## 3763 of these out of 28694+4291 cases had NA values converted to 0 - 2019 assessment
## 5656 of these out of 28137+5656 cases had NA values converted to 0 - 2020 assessment

```

Remove species-region-environment time series with less than four years of mariculture production > 0 tonnes (assume these are not established mariculture programs).

```{r remove-less-four-years}

mar_rgn_gf = mar_rgn_gf %>% 
  group_by(rgn_id, species, fao, environment) %>%
  mutate (not_0 = length(value[value>0])) %>% # Length of vector of years greater than 0
  filter (not_0>3) %>% # Filter for groups that have at least four years of mariculture production 
  ungroup() %>% 
  select(rgn_id, species, fao, environment, year, value, Taxon_code, gap_0_fill) 

```

Add a unique identifier per cultivated stock that describes each species, fao region, and environment grouping.

```{r add-identifier}

## Add a unique identifier per cultivated stock
identifier = mar_rgn_gf %>% 
  select(rgn_id, species, fao, environment) %>% 
  unique() %>% 
  mutate(species_code = 1:n())

mar_rgn_gf = left_join(mar_rgn_gf, identifier)
maric <- mar_rgn_gf

```

# Save file: 

Used to estimate total mariculture yield per country.

Saves the appropriate Mariculture-FP file depending on whether we are excluding species based on best judgment (exclude), all seaweeds (exclude_no_seaweed), or just seaweeds nei (exclude_no_nei). See method changes above.

```{r save-mariculture-yield}

if(sum(str_detect(names(mar_sp), "exclude$"))==1) {
 write.csv(maric, 'output/MAR_FP_data.csv', row.names=FALSE)
 } else if (sum(str_detect(names(mar_sp), "exclude_no_seaweed"))==1) {
   write.csv(maric, 'test/MAR_FP_data_no_seaweed.csv', row.names=FALSE)
   } else if (sum(str_detect(names(mar_sp), "exclude_no_nei"))==1) {
     write.csv(maric, 'test/MAR_FP_data_no_nei.csv', row.names=FALSE)
     }

```

# Import data: Trujillo sustainability scores

These data describe the sustainability and genetic escapes for country/species combinations (and, in a couple cases, environment and FAO region combinations). In cases where these data were not available for a specific county/species, we averaged the data across taxonomic groups to gapfill the missing data.

```{r trujillo-scores}

## Trujillo sustainability data:
sus = read.csv('raw/Truj_label_sust.csv', stringsAsFactors = FALSE, na.strings = NA)

```

# Wrangle

```{r re-add-scores}

## These need to be re-added (get cut when associated with region ids)
sus_no_rgn <- filter(sus, is.na(country))

```

## Convert country names to OHI region names.

```{r sus_convert-region}

sus_rgn <- name_2_rgn(df_in = sus, 
                       fld_name='country', 
                       flds_unique=c('species_fao', 'fao', 'environment', 'species_Truj'))

sus_rgn <- bind_rows(sus_rgn, sus_no_rgn) %>%
  unique()

```

**Check that the non-matches between Trujillo sustainability FAO spp (sus_rgn) and the FAO mariculture species in the wrangled FAO Aquaculture Production data table (maric) are not due to spelling errors or slightly different names. We want to include as many species that have sustainability scores as possible**

```{r sus_check-non-matches}

## Make sure same species are spelled the same in the two data tables (e.g. check that there are no extra spaces)
sort(setdiff(sus_rgn$species_fao, maric$species)) # Species that are no longer have mariculture industry or are not included due to being freshwater or non-food

sort(setdiff(maric$species, sus_rgn$species_fao)) # FAO species with no Trujillo data - there will probably be a long list

## Hand check each of the species output here. I mainly check to make sure there aren't obvious cases when the species is the same, but the names are presented slightly differently in the lists. For example, some species might have an extra space in the name somewhere. If any of these obvious differences exist, we need to go in and change the names to match. 
## Its not too surprising (unfortunately) that the FAO list has far more species than the sustainability list. But, I like to go over them to make sure that we are getting sustainability scores for as many species as we can.

```

# FAO mariculture and sustainability scores

Append sustainability score to the FAO mariculture data.

The following joins the sustainability scores to regions/species that have Trujillo data.

```{r add-sus-fao-data}

table(sus_rgn$match_type)

## Join taxa specific to country/species/environment
c_sp_env = sus_rgn %>% 
  filter(match_type == 'c_sp_env') %>% 
  select(rgn_id, species=species_fao, environment, Sust_c_sp_env = Maric_sustainability)

mar_sus <- maric %>%
  left_join(c_sp_env, by= c("species", "environment", "rgn_id"))

## Join taxa specific to country/species/fao region
c_sp_fao = sus_rgn %>% 
  filter(match_type == 'c_sp_fao') %>% 
  select(rgn_id, species=species_fao, fao, Sust_c_sp_fao = Maric_sustainability)

mar_sus <- mar_sus %>%
  left_join(c_sp_fao, by= c("species", "fao", "rgn_id"))

```

**Take a look at the data thus far**

```{r view-truj-sus-data}

head(data.frame(filter(mar_sus, rgn_id==218 & species == "Atlantic salmon")))
head(data.frame(filter(mar_sus, !is.na(Sust_c_sp_fao))))

```

```{r sus_join-country-species}

## Join taxa specific to country/species
c_sp = sus_rgn %>% 
  filter(match_type == 'c_sp') %>% 
  select(rgn_id, species=species_fao, Sust_c_sp = Maric_sustainability)

mar_sus <- mar_sus %>%
  left_join(c_sp, by= c("species", "rgn_id"))

```

**Now look at a summary after appending all the Trujillo data**

```{r truj-sus-summary}

summary(mar_sus)

```

Merge the three Trujillo type categories into a single sustainability score column in the following order:

1. Sust_c_sp_env: taxa specific to country/species/environment (smallest taxonomic level)
2. Sust_c_sp_fao:  taxa specific to country/species/fao region
3. Sust_c_sp: taxa specific to country/species (highest taxonomic level)

For example, if Sust_c_sp_env is missing, use Sust_c_sp_fao and so on. 

```{r truj-sus-merge}

mar_sus = mar_sus %>% 
  mutate(Sust = ifelse(!is.na(Sust_c_sp_env), Sust_c_sp_env, Sust_c_sp_fao)) %>%
  mutate(Sust = ifelse(is.na(Sust), Sust_c_sp, Sust)) %>%
  select(-Sust_c_sp_env, -Sust_c_sp_fao, -Sust_c_sp)

```

This joins the sustainability data that is gapfilled either at the species level (average of specific species/genera across regions) or at a higher course taxonomic levels and documents which data are gapfilled and how.

```{r truj-sus-join-gapfilled}

## Select observations gapfilled at the species/genera level:
gf_sp_sus <- sus_rgn %>% 
  filter(gapfill != "actuals" & match_type == "species") %>%
  select(species = species_fao, gapfill, Sust_gf_sp = Maric_sustainability)

## Check that there are no duplicated species_fao
gf_sp_sus[duplicated(gf_sp_sus$species), ]

## Match gapfilling values by species
mar_sus_gf = mar_sus %>%
  left_join(gf_sp_sus, by = 'species')

## Select observations gapfilled at the coarse taxon level:
gf_taxon_sus <- sus_rgn %>% 
  filter(gapfill != "actuals" & match_type == "taxon") %>%
  select(Taxon_code=taxon, Sust_gf_taxon = Maric_sustainability)

## Match gapfilling values by species
mar_sus_gf = mar_sus_gf %>%
  left_join(gf_taxon_sus, by = c('Taxon_code'))

```

**Take a look at the wrangled data**

```{r view-truj-sus-gapfilled}

summary(mar_sus_gf) # Should be no NAs in Sust_gf_taxon column
table(mar_sus_gf$gapfill)

```

Obtain a sustainability score for each record, and a book-keeping column of whether it's actual or gapfilled

For missing sustainability scores:

1. Use species-level Sustainability score
2. If no species-level scores, gapfill with taxon-level sustainability average

```{r truj-sus-gapfill-missing}

mar_sus_final = mar_sus_gf %>% 
  mutate(gapfill = ifelse(!is.na(Sust), "none", gapfill)) %>%
  mutate(Sust = ifelse(is.na(Sust), Sust_gf_sp, Sust)) %>% # Gapfill with species level
  mutate(gapfill = ifelse(is.na(Sust), "taxon_average", gapfill)) %>% # Add in taxon gapfill record
  mutate(Sust = ifelse(is.na(Sust), Sust_gf_taxon, Sust)) %>% # Gapfill with taxon level
  mutate(taxa_code = paste(species, species_code, sep="_")) %>%
  select(rgn_id, species, species_code, taxa_code, taxa_group=Taxon_code, year, gapfill_sus = gapfill, gapfill_fao = gap_0_fill, tonnes=value, Sust)

```

# Save Data:

```{r save-mar-harvest-data}

## Save mariculture harvest data
mar_harvest_tonnes = mar_sus_final %>%
  select(rgn_id, taxa_code, taxa_group, year, tonnes)

anyDuplicated(mar_harvest_tonnes) # Check for duplication

## Saves the appropriate Mariculture Harvest Tonnes file depending on whether we are excluding species based on best judgment (exclude), all seaweeds (exclude_no_seaweed), or just seaweeds nei (exclude_no_nei). See method changes above.
if(sum(str_detect(names(mar_sp), "exclude$"))==1) {
 write.csv(mar_harvest_tonnes, 'output/mar_harvest_tonnes.csv', row.names=F)
 } else if (sum(str_detect(names(mar_sp), "exclude_no_seaweed"))==1) {
   write.csv(mar_harvest_tonnes, 'test/mar_harvest_tonnes_no_seaweed.csv', row.names=F)
   } else if (sum(str_detect(names(mar_sp), "exclude_no_nei"))==1) {
     write.csv(mar_harvest_tonnes, 'test/mar_harvest_tonnes_no_nei.csv', row.names=F)
     }

## Save gapfill data for mariculture harvest
mar_harvest_tonnes_gf = mar_sus_final %>%
  select(rgn_id, taxa_code, taxa_group, year, tonnes=gapfill_fao)


## Saves the appropriate Mariculture Harvest Gapfill file depending on whether we are excluding species based on best judgment (exclude), all seaweeds (exclude_no_seaweed), or just seaweeds nei (exclude_no_nei). See method changes above
if(sum(str_detect(names(mar_sp), "exclude$"))==1) {
 write.csv(mar_harvest_tonnes_gf, 'output/mar_harvest_tonnes_gf.csv', row.names=F)
 } else if (sum(str_detect(names(mar_sp), "exclude_no_seaweed"))==1) {
   write.csv(mar_harvest_tonnes_gf, 'test/mar_harvest_tonnes_gf_no_seaweed.csv', row.names=F)
   } else if (sum(str_detect(names(mar_sp), "exclude_no_nei"))==1) {
     write.csv(mar_harvest_tonnes_gf, 'test/mar_harvest_tonnes_gf_no_nei.csv', row.names=F)
     }


## Save sustainability scores data for 2012
mar_sustainability_score = mar_sus_final %>% 
  mutate(year = 2012) %>% # Only 2012 sustainability scores exist (Trujillo)
  select(rgn_id, year, taxa_code, sust_coeff = Sust) %>% 
  unique()

anyDuplicated(mar_sustainability_score)


## Saves the appropriate Sustainability Score file depending on whether we are excluding species based on best judgment (exclude), all seaweeds (exclude_no_seaweed), or just seaweeds nei (exclude_no_nei). See method changes above
if(sum(str_detect(names(mar_sp), "exclude$"))==1) {
 write.csv(mar_sustainability_score, 'output/mar_sustainability.csv', row.names=F)
 } else if (sum(str_detect(names(mar_sp), "exclude_no_seaweed"))==1) {
   write.csv(mar_sustainability_score, 'test/mar_sustainability_no_seaweed.csv', row.names=F)
   } else if (sum(str_detect(names(mar_sp), "exclude_no_nei"))==1) {
     write.csv(mar_sustainability_score, 'test/mar_sustainability_no_nei.csv', row.names=F)
     }


## Save gapfill data for sustainability scores
mar_sustainability_score_gf = mar_sus_final %>% 
  select(rgn_id, taxa_code, sust_coeff = gapfill_sus) %>% 
  unique()


## Saves the appropriate Sustainability Score Gapfill file depending on whether we are excluding species based on best judgment (exclude), all seaweeds (exclude_no_seaweed), or just seaweeds nei (exclude_no_nei). See method changes above
if(sum(str_detect(names(mar_sp), "exclude$"))==1) {
 write.csv(mar_sustainability_score_gf, 'output/mar_sustainability_gf.csv', row.names=F)
 } else if (sum(str_detect(names(mar_sp), "exclude_no_seaweed"))==1) {
   write.csv(mar_sustainability_score_gf, 'test/mar_sustainability_gf_no_seaweed.csv', row.names=F)
   } else if (sum(str_detect(names(mar_sp), "exclude_no_nei"))==1) {
     write.csv(mar_sustainability_score_gf, 'test/mar_sustainability_gf_no_nei.csv', row.names=F)
     }

```

# Wrangle: Genetic Escapes Data

Combine genetic escapes data to mariculture data. 

* Data from Trujillo data table. 
* These data are used as a pressure layer to describe the risk of genetic escapees due to mariculture.
* First merge with the species data (no gapfilling) for each country/species/fao region combination.

```{r wrangle-genetic-escapes-data}

## Can eliminate the environment category because these have the same scores
esc = sus_rgn %>% 
  filter(!is.na(Genetic.escapees)) %>%
  mutate(match_type = ifelse(match_type == "c_sp_env", "c_sp", match_type)) %>%
  group_by(rgn_id, species=species_fao, fao, match_type, taxon, gapfill) %>%
  summarize(Genetic.escapees = mean(Genetic.escapees)) %>%
  ungroup()
  
## Join taxa specific to country/species/fao
c_sp_fao = esc %>% 
  filter(match_type == 'c_sp_fao') %>% 
  select(rgn_id, species, fao, Esc_c_sp_fao = Genetic.escapees) 

mar_esc <- maric %>%
  left_join(c_sp_fao, by= c("species", "fao", "rgn_id"))

head(data.frame(filter(mar_esc, !is.na(Esc_c_sp_fao))))

## Join taxa specific to country/species
c_sp = esc %>% 
  filter(match_type == 'c_sp') %>% 
  select(rgn_id, species, Esc_c_sp = Genetic.escapees) # fao is blank

mar_esc <- mar_esc %>%
  left_join(c_sp, by= c("species", "rgn_id"))

```

**Look at a summary of the wrangled data table**

```{r summary-genetic-escapes}

summary(mar_esc)

```

Merge the different match types (esc_c_sp_fao, esc_c_sp) into a single sustainability score

1. First use Esc_c_sp_fao
2. If no Esc_c_sp_fao value, use Esc_c_sp

```{r merge-genetic-escapes-scores}

mar_esc = mar_esc %>% 
  mutate(Escapees = ifelse(!is.na(Esc_c_sp_fao), Esc_c_sp_fao, Esc_c_sp)) %>%
  select(-Esc_c_sp_fao, -Esc_c_sp)

```

# Gapfill 

Join the sustainability data that is gapfilled either at the species level (average of specific species/genera across regions) or at a higher course taxonomic levels and documents which data are gapfilled and how.
```{r gapfill-genetic-escapes}

## Select observations gapfilled at the species/genera level:
gf_species_esc <- esc %>% 
  filter(gapfill != "actuals" & match_type == "species") %>%
  select(species, gapfill, Esc_gf_sp = Genetic.escapees)

## Check that there are no duplicated species_fao
gf_species_esc[duplicated(gf_species_esc$species), ]

## Match gapfilling values by species
mar_esc_gf = mar_esc %>%
  left_join(gf_species_esc, by = 'species')

## Select observations gapfilled at the coarse taxon level:
gf_taxon_sus <- esc %>% 
  filter(gapfill != "actuals" & match_type == "taxon") %>%
  select(Taxon_code=taxon, Esc_gf_taxon = Genetic.escapees)

## Match gapfilling values by species
mar_esc_gf = mar_esc_gf %>%
  left_join(gf_taxon_sus, by = c('Taxon_code'))

```

**Take a look at wrangled data**

```{r summary-genetic-escapes-gapfilled}

summary(mar_esc_gf)

table(mar_esc_gf$gapfill)

```

## Record Gapfilling Methods

Obtain a sustainability score for each record, and a book-keeping column of whether it's actual or gap-filled.

For gapfill column

1. If no missing escapee data, record "none"
2. If NA, then gapfill with `Esc_gf_sp` (gapfill for species-level already recorded)
3. If no `Esc_gf_sp` value, then record "taxon_average" as the gapfill method
4. Fill in missing value with `Esc_gf_taxon` value

```{r record-gen-esc-gapfill}

tonnes_esc = mar_esc_gf %>% 
  mutate(gapfill = ifelse(!is.na(Escapees), "none", gapfill)) %>% 
  mutate(Escapees = ifelse(is.na(Escapees), Esc_gf_sp, Escapees)) %>%
  mutate(gapfill = ifelse(is.na(Escapees), "taxon_average", gapfill)) %>%
  mutate(Escapees = ifelse(is.na(Escapees), Esc_gf_taxon, Escapees)) %>%
  select(rgn_id, species, species_code, year, gapfill_escapees = gapfill, tonnes=value, Escapees)

```

**Take a look at the summary**

```{r summary_gapfilled-gen-esc}

summary(tonnes_esc)

anyDuplicated(tonnes_esc) # Check for duplicates

```

Final formatting of the escapee data. This is used as a pressure layer.

For each region/year: take a weighted average of the genetic escape probability for each taxa based on tonnes mariculture
```{r gen-esc-formatting}

genEscapes <- tonnes_esc %>%
  group_by(rgn_id, year) %>%
  summarize(genEscapes = weighted.mean(Escapees, tonnes, na.rm=TRUE))

```

Obtain gapfill information from `genEscapes`

Gapfill values in output file are the proportion of data that is gapfilled.

```{r obtain-gapfill-info-genEscapes}

## Obtain corresponding gapfilling information for each region (average of gapfilled data, weighted by tonnes of mariculture).
genEscapes_gf <- tonnes_esc %>%
  mutate(gapfill = ifelse(gapfill_escapees=="none", 0, 1)) %>%
  group_by(rgn_id, year) %>%
  summarize(genEscapes = weighted.mean(gapfill, tonnes, na.rm=TRUE)) %>%
  ungroup() %>%
  filter(year==2018) %>%
  select(rgn_id, pressures.score=genEscapes) %>%
  mutate(pressures.score=ifelse(pressures.score=="NaN", NA, pressures.score)) %>%
  data.frame()

```

## Save gapfill data

```{r save-gen-esc-gapfilled}

## Saves the appropriate Genetic Escapes Gapfill file depending on whether we are excluding species based on best judgment (exclude), all seaweeds (exclude_no_seaweed), or just seaweeds nei (exclude_no_nei). See method changes above
if(sum(str_detect(names(mar_sp), "exclude$"))==1) {
 write.csv(genEscapes_gf, 'output/GenEsc_gf.csv', row.names=FALSE)
 } else if (sum(str_detect(names(mar_sp), "exclude_no_seaweed"))==1) {
   write.csv(genEscapes_gf, 'test/GenEsc_gf_no_seaweed.csv', row.names=FALSE)
   } else if (sum(str_detect(names(mar_sp), "exclude_no_nei"))==1) {
     write.csv(genEscapes_gf, 'test/GenEsc_gf_no_nei.csv', row.names=FALSE)
     }

```

Obtain escapee data layers from `genEscapes`

```{r create-gen-esc-layers}

# Create the escapee data layers:
data <- genEscapes %>%
    select(rgn_id, year, pressure_score = genEscapes)

```

## Save escapee pressure layer 

```{r save-gen-esc-pressure-layer}

## Saves the appropriate Genetic Escapes data file depending on whether we are excluding species based on best judgment (exclude), all seaweeds (exclude_no_seaweed), or just seaweeds nei (exclude_no_nei). See method changes above
if(sum(str_detect(names(mar_sp), "exclude$"))==1) {
 write.csv(data, 'output/GenEsc.csv', row.names=FALSE)  
 } else if (sum(str_detect(names(mar_sp), "exclude_no_seaweed"))==1) {
   write.csv(data, 'test/GenEsc_no_seaweed.csv', row.names=FALSE) 
   } else if (sum(str_detect(names(mar_sp), "exclude_no_nei"))==1) {
    write.csv(data, 'test/GenEsc_no_nei.csv', row.names=FALSE)  
     }

```

## Data check

Comparing this year's data to previous year's data. Expect small variation from year to year. Plot to view differences.

```{r data-check}

## Compare genetic escapes pressure scores for Russia; saw large changes between 2018 and 2017 assessment years. No big changes between 2019 and 2018 assessment years. No changes between 2020 and 2019 assessment years.
old <- read.csv("../v2019/output/GenEsc.csv") %>%
  filter(rgn_id == 73) %>% 
  select(rgn_id, year, prs_score_old=pressure_score)

new <- read.csv("../v2020/output/GenEsc.csv") %>%
  filter(rgn_id == 73) %>% 
  select(rgn_id, year, pressure_score)

compare <- full_join(old, new, by=c("rgn_id","year")); View(compare)

plot(compare$prs_score_old, compare$pressure_score);abline(0,1,col="red")



## Compare yield data for Russia
mar_old <- read.csv("../v2019/output/MAR_FP_data.csv") %>% 
  filter(rgn_id == 73, year == 2017) %>% 
  select(rgn_id, species, fao, old_value = value)
mar_new <- read.csv("output/MAR_FP_data.csv") %>% 
  filter(rgn_id == 73, year == 2018) %>% 
  select(rgn_id, species, fao, value)

yield <- mar_old %>% 
  full_join(mar_new, by = c("rgn_id","species","fao")); View(yield)



## Compare genetic escapes pressure scores for Vietnam; no change in pressures between 2020 and 2019 assessment years.
old <- read.csv("../v2019/output/GenEsc.csv") %>%
  filter(rgn_id == 207) %>% 
  select(rgn_id, year, prs_score_old=pressure_score)

new <- read.csv("../v2020/output/GenEsc.csv") %>%
  filter(rgn_id == 207) %>% 
  select(rgn_id, year, pressure_score)

compare <- full_join(old, new, by=c("rgn_id","year")); View(compare)

plot(compare$prs_score_old, compare$pressure_score);abline(0,1,col="red")



## Compare yield data for Vietnam
mar_old <- read.csv("../v2019/output/MAR_FP_data.csv") %>% 
  filter(rgn_id == 207, year == 2017) %>% 
  select(rgn_id, species, fao, environment, old_value = value)
mar_new <- read.csv("output/MAR_FP_data.csv") %>% 
  filter(rgn_id == 207, year == 2018) %>% 
  select(rgn_id, species, fao, environment, value)

yield <- mar_old %>% 
  full_join(mar_new, by = c("rgn_id","species","fao", "environment")); View(yield)



## Compare genetic escapes pressure scores for Iceland; saw large changes between 2018 and 2017 assessment years, we do not see a big difference between 2019 and 2018 assessment years or 2020 and 2019 assessment years. 
old <- read.csv("../v2019/output/GenEsc.csv") %>%
  filter(rgn_id == 143) %>% 
  select(rgn_id, year, prs_score_old=pressure_score)

new <- read.csv("../v2020/output/GenEsc.csv") %>%
  filter(rgn_id == 143) %>% 
  select(rgn_id, year, pressure_score)

compare <- full_join(old, new, by=c("rgn_id","year")); View(compare)

plot(compare$prs_score_old, compare$pressure_score);abline(0,1,col="red")



## Compare yield data for Iceland. Atlantic Salmon increased by a lot between 2018 and 2019 assessment years. Atlantic Salmon stayed relatively stable between 2020 and 2019 assessment years. However, Senegalese sole has an NA value in 2019 assessment; maybe a gapfilling error from last year?
mar_old <- read.csv("../v2019/output/MAR_FP_data.csv") %>% 
  filter(rgn_id == 143, year == 2017) %>% 
  select(rgn_id, species, fao, environment, old_value = value)
mar_new <- read.csv("output/MAR_FP_data.csv") %>% 
  filter(rgn_id == 143, year == 2018) %>% 
  select(rgn_id, species, fao, environment, value)

yield <- mar_old %>% 
  full_join(mar_new, by = c("rgn_id","species","fao", "environment")); View(yield)



## Compare genetic escapes pressure scores for Belize; saw large changes between 2018 and 2017 assessment years, we do not see big differences between 2019 and 2018 assessment years or 2020 and 2019 assessment years. 
old <- read.csv("../v2019/output/GenEsc.csv") %>%
  filter(rgn_id == 164) %>% 
  select(rgn_id, year, prs_score_old=pressure_score)

new <- read.csv("../v2020/output/GenEsc.csv") %>%
  filter(rgn_id == 164) %>% 
  select(rgn_id, year, pressure_score)

compare <- full_join(old, new, by=c("rgn_id","year")); View(compare)

plot(compare$prs_score_old, compare$pressure_score);abline(0,1,col="red")

## Compare yield data for Belize; yield of Whiteleg shrimp was halved between 2019 and 2020 assessment years.
mar_old <- read.csv("../v2019/output/MAR_FP_data.csv") %>% 
  filter(rgn_id == 164, year == 2017) %>% 
  select(rgn_id, species, fao, old_value = value)
mar_new <- read.csv("output/MAR_FP_data.csv") %>% 
  filter(rgn_id == 164, year == 2018) %>% 
  select(rgn_id, species, fao, value)

yield <- mar_old %>% 
  full_join(mar_new, by = c("rgn_id","species","fao")); View(yield)


## Compare genetic escapes pressure scores for France; saw large changes between 2018 and 2017 assessment years, we do not see big differences between 2019 and 2018 assessment years or 2020 and 2019 assessment years. 
old <- read.csv("../v2019/output/GenEsc.csv") %>%
  filter(rgn_id == 179) %>% 
  select(rgn_id, year, prs_score_old=pressure_score)

new <- read.csv("../v2020/output/GenEsc.csv") %>%
  filter(rgn_id == 179) %>% 
  select(rgn_id, year, pressure_score)

compare <- full_join(old, new, by=c("rgn_id","year")); View(compare)

plot(compare$prs_score_old, compare$pressure_score);abline(0,1,col="red")


## Compare yield data for France
mar_old <- read.csv("../v2019/output/MAR_FP_data.csv") %>% 
  filter(rgn_id == 179, year == 2017) %>% 
  select(rgn_id, species, fao, old_value = value)
mar_new <- read.csv("output/MAR_FP_data.csv") %>% 
  filter(rgn_id == 179, year == 2018) %>% 
  select(rgn_id, species, fao, value)

yield <- mar_old %>% 
  full_join(mar_new, by = c("rgn_id","species","fao")); View(yield)

plot(yield$old_value, yield$value);abline(0,1,col="red")


## Compare genetic escapes pressure scores for Malta; saw large changes between 2019 and 2020 assessment years. 
old <- read.csv("../v2019/output/GenEsc.csv") %>%
  filter(rgn_id == 68) %>% 
  select(rgn_id, year, prs_score_old=pressure_score)

new <- read.csv("../v2020/output/GenEsc.csv") %>%
  filter(rgn_id == 68) %>% 
  select(rgn_id, year, pressure_score)

compare <- full_join(old, new, by=c("rgn_id","year")); View(compare)

plot(compare$prs_score_old, compare$pressure_score);abline(0,1,col="red")


## Compare yield data for Malta
mar_old <- read.csv("../v2019/output/MAR_FP_data.csv") %>% 
  filter(rgn_id == 68, year == 2017) %>% 
  select(rgn_id, species, fao, old_value = value)
mar_new <- read.csv("output/MAR_FP_data.csv") %>% 
  filter(rgn_id == 68, year == 2018) %>% 
  select(rgn_id, species, fao, value)

yield <- mar_old %>% 
  full_join(mar_new, by = c("rgn_id","species","fao")); View(yield)

plot(yield$old_value, yield$value);abline(0,1,col="red")


## Compare new/old Mariculture sustainability scores
tmp_old <- read.csv("../v2019/raw/Truj_label_sust.csv") %>% 
  filter(country == "Russian Federation") %>% 
  select(country, species_Truj, sust_old = Maric_sustainability, gen_old = Genetic.escapees)
  
tmp_new <- read.csv("../v2020/raw/Truj_label_sust.csv") %>% 
  filter(country == "Russian Federation") %>% 
  select(country, species_Truj, Maric_sustainability, Genetic.escapees)

test <- tmp_new %>% 
  full_join(tmp_old, by=c("country", "species_Truj"))
View(test)

plot(test$sust_old, test$Maric_sustainability);abline(0,1,col="red")


## Compare new/old genetic escapees scores
plot(test$gen_old, test$Genetic.escapees);abline(0,1,col="red")

```


## Data Check

Compare the three exclusion methods for seaweed species.

*Method 1:Exclude some seaweed species (`exclude`)
*Method 2 (original): Exclude all seaweed species (`exclude_no_sw`)
*Method 3: Exclude all seaweed nei species (`exclude_no_nei`)

```{r data-check-seaweed}

## Read in the three different versions of wrangled data
## Best Professional Judgement (BPJ)
exclude <- read.csv("../v2020/output/GenEsc.csv") %>%
  filter(year == 2018) %>% # compare the most recent shared year between old and new
  select(rgn_id, prs_score_BPJ=pressure_score)

## Exclude all seaweeds (no seaweed)
exclude_no_sw <- read.csv("../v2020/test/GenEsc_no_seaweed.csv") %>%
  filter(year == 2018) %>% # compare the most recent shared year between old and new
  select(rgn_id, prs_score_no_sw=pressure_score)

## Exclude only nei seaweeds (no nei)
exclude_no_nei <- read.csv("../v2020/test/GenEsc_no_nei.csv") %>%
  filter(year == 2018) %>% # compare the most recent shared year between old and new
  select(rgn_id, prs_score_no_nei=pressure_score)


## Compare No Seaweed to Best Professional Judgement
compare <- full_join(exclude, exclude_no_sw, by="rgn_id")
compare <- na.omit(compare)
## Compare genetic escapes pressure scores
plot(compare$prs_score_BPJ, compare$prs_score_no_sw); abline(0,1,col="red")

## Compare No Nei to No Seaweed (not much difference)
compare <- full_join(exclude_no_nei, exclude_no_sw, by="rgn_id")
compare <- na.omit(compare)
## Compare genetic escapes pressure scores
plot(compare$prs_score_no_nei, compare$prs_score_no_sw); abline(0,1,col="red")

## Compare No Nei to BPJ
compare <- full_join(exclude_no_nei, exclude, by="rgn_id")
compare <- na.omit(compare)
## Compare genetic escapes pressure scores
plot(compare$prs_score_BPJ, compare$prs_score_no_nei); abline(0,1,col="red")


```

# New Sustainability Scores from Seafood Watch Data

## Import data: Seafood Watch sustainability scores

These data describe the sustainability country/species combinations. In cases where these data were not available for a specific county/species, we averaged the data across taxonomic groups to gapfill the missing data.

```{r sw-scores}

## Load in Seafood Watch sustainability scores data from mazu:
sw_sus <- read.csv(file.path(dir_M, 'git-annex/globalprep/_raw_data/seafood_watch_mar_sustainability/d2020/Seafood-Watch_aquaculture-recs_July-2020.csv'), check.names = FALSE, stringsAsFactors = FALSE, na.strings = c("NA", ""))

head(sw_sus)

```

## Wrangle

### Tidy Seafood Watch sustainability data

Rename columns to match with MAR data and fill in species column

```{r tidy_sw-sus}

## Rename columns
sw_sus <- sw_sus %>%
  rename(report_title = 'Report Title',
         start_year = 'Start year',
         sw_species = 'Common name',
         genus = 'Genus',
         spp = 'Species',
         fao_species = 'FAO Common name',
         region = 'Region',
         country = 'Country',
         state_territory = 'State/Territory',
         sub_region = 'Sub-Region',
         water_body = 'Body of Water',
         parent_method = 'Parent Method',
         method = 'Method',
         score = 'Overall Score',
         rec = 'Overall Recommendation'
         ) %>% 
  dplyr::select(report_title, start_year, sw_species, genus, spp, fao_species, region, country, state_territory, sub_region, water_body, parent_method, method, score, rec)

## Change species names using FAO species name (fao_species); if NA, use common name (sw_species)
sw_sus$species <- ifelse(!is.na(sw_sus$fao_species), sw_sus$fao_species, sw_sus$sw_species)

```

### Keep NA countries

```{r keep-na-countries_sw-sus}

## These need to be re-added later (get cut when associated with region ids)
sw_sus_no_rgn <- filter(sw_sus, is.na(country))
  # 203 entries with no country

```

### Convert country names to OHI region IDs.

```{r sw-sus_convert-region}

## Change country names to match OHI region names
sw_sus <- sw_sus %>% 
  mutate(country = ifelse(country=="Korea, the Republic of", "South Korea", country)) %>% # Data removed for not having a match; change name to match
  mutate(country = ifelse(country=="United Kingdom of Great Britain and Northern Ireland (the)", "United Kingdom", country))  # Data removed for not having a match; change name to match

## Convert country names to OHI region IDs. (ohicore/R/name_2_rgn.R)
sw_sus_rgn <- name_2_rgn(df_in = sw_sus, 
                       fld_name='country', 
                       flds_unique=c('fao_species', 'sw_species', 'region', 'score'),
                       keep_fld_name = TRUE) # 'country' now shows the original Seafood Watch data name; 'rgn_name' is what we want to use from now on
  # Goes from 330 obs. to 127 obs. (because 203 obs. have no country associated)


## Re-add NA countries
sw_sus_rgn <- bind_rows(sw_sus_rgn, sw_sus_no_rgn) %>%
  unique()
  # Back to 330 obs.

```

**Check that the non-matches between Seafood Watch sustainability data species (sw_sus_rgn) and the FAO mariculture species in the wrangled FAO Aquaculture Production data table (maric) are not due to spelling errors or slightly different names. We want to include as many species that have sustainability scores as possible**

```{r sw-sus_check-non-matches}

## Make sure same species are spelled the same in the two data tables (e.g. check that there are no extra spaces)
sort(setdiff(sw_sus_rgn$species, maric$species)) # Seafood Watch species with no FAO data

sort(setdiff(maric$species, sw_sus_rgn$species)) # FAO species with no Seafood Watch data - there will probably be a long list

## Hand check each of the species output here. I mainly check to make sure there aren't obvious cases when the species is the same, but the names are presented slightly differently in the lists. For example, some species might have an extra space in the name somewhere. If any of these obvious differences exist, we need to go in and change the names to match. 
## Its not too surprising (unfortunately) that the FAO list has far more species than the sustainability list. But, I like to go over them to make sure that we are getting sustainability scores for as many species as we can.

```

```{r sw-sus_fix-non-matches}

## Rename species in Seafood Watch data to match FAO species data
sw_sus_rgn$species <- gsub("Abalone", "Abalones nei", sw_sus_rgn$species)
sw_sus_rgn$species <- gsub("Barramundi", "Barramundi(=Giant seaperch)", sw_sus_rgn$species)
sw_sus_rgn$species <- gsub("Coho salmon", "Coho(=Silver) salmon", sw_sus_rgn$species)
sw_sus_rgn$species <- gsub("Manila clam", "Japanese carpet shell", sw_sus_rgn$species) # Japanese carpet shell is another name for Manila clam (thanks animal crossing :P )
sw_sus_rgn$species <- gsub("New Zealand Mussel", "New Zealand mussel", sw_sus_rgn$species)
sw_sus_rgn$species <- gsub("Hard clam (unspecified)", "Northern quahog(=Hard clam)", sw_sus_rgn$species) # Hard clams are also known as quahogs, round clams, or hard-shell clams

```

# FAO mariculture and Seafood Watch sustainability scores

```{r match-type_sw-sus-data}

## Add column "match_type" to categorize obs. that 1) have a country associated, 2) are "global" Seafood Watch data, or 3) only distinguished by water body
sw_sus_rgn <- sw_sus_rgn %>% 
  dplyr::mutate(match_type = dplyr::case_when(!is.na(country) ~ "sw_sp_c", # Add match type specific to species/country
                                str_detect(report_title, regex("Global")) ~ "sw_sp_g", # Add match type specific to species/global Seafood Watch data
                                TRUE ~ "sw_sp_w" # Add match type specific to species/water body
  ))

table(sw_sus_rgn$match_type)

```

```{r gapfill-sw-sus-match-rgn-id}

## Add column "gapfill" to track which match_type uses the Seafood Watch sustainability data directly, or needs to be gapfilled
sw_sus_rgn <- sw_sus_rgn %>% 
  dplyr::mutate(gapfill = dplyr::case_when(match_type == "sw_sp_c" ~ "actuals",
                                           match_type == "sw_sp_g" ~ "sp_global",
                                           match_type == "sw_sp_w" ~ "sp_waterbody"))

```

```{r gf_sw_sp_w}

## Add a line for each country that borders each water body
  # The only water body is the Mediterranean; add all countries that border it (source: https://www.medqsr.org/mediterranean-marine-and-coastal-environment#:~:text=Today%2021%20countries%2C%20with%20surface,Syria%2C%20Tunisia%2C%20and%20Turkey.)
gf_sw_sp_w <- sw_sus_rgn %>% 
  filter(match_type == 'sw_sp_w') %>% 
  
  # Atlantic bluefin tuna - Marine net pen
  add_row(rgn_id = 82, rgn_name = 'Albania', species = 'Atlantic bluefin tuna', score = 4.42, match_type = 'sw_sp_w', gapfill = 'sp_waterbody', .after = 1) %>% 
  add_row(rgn_id = 84, rgn_name = 'Algeria', species = 'Atlantic bluefin tuna', score = 4.42, match_type = 'sw_sp_w', gapfill = 'sp_waterbody', .after = 1) %>%
  add_row(rgn_id = 232, rgn_name = 'Bosnia and Herzegovina', species = 'Atlantic bluefin tuna', score = 4.42, match_type = 'sw_sp_w', gapfill = 'sp_waterbody', .after = 1) %>%
  add_row(rgn_id = 187, rgn_name = 'Croatia', species = 'Atlantic bluefin tuna', score = 4.42, match_type = 'sw_sp_w', gapfill = 'sp_waterbody', .after = 1) %>%
  add_row(rgn_id = 81, rgn_name = 'Cyprus', species = 'Atlantic bluefin tuna', score = 4.42, match_type = 'sw_sp_w', gapfill = 'sp_waterbody', .after = 1) %>%
  add_row(rgn_id = 214, rgn_name = 'Egypt', species = 'Atlantic bluefin tuna', score = 4.42, match_type = 'sw_sp_w', gapfill = 'sp_waterbody', .after = 1) %>%
  add_row(rgn_id = 179, rgn_name = 'France', species = 'Atlantic bluefin tuna', score = 4.42, match_type = 'sw_sp_w', gapfill = 'sp_waterbody', .after = 1) %>%
  add_row(rgn_id = 80, rgn_name = 'Greece', species = 'Atlantic bluefin tuna', score = 4.42, match_type = 'sw_sp_w', gapfill = 'sp_waterbody', .after = 1) %>%
  add_row(rgn_id = 79, rgn_name = 'Israel', species = 'Atlantic bluefin tuna', score = 4.42, match_type = 'sw_sp_w', gapfill = 'sp_waterbody', .after = 1) %>%
  add_row(rgn_id = 184, rgn_name = 'Italy', species = 'Atlantic bluefin tuna', score = 4.42, match_type = 'sw_sp_w', gapfill = 'sp_waterbody', .after = 1) %>%
  add_row(rgn_id = 78, rgn_name = 'Lebanon', species = 'Atlantic bluefin tuna', score = 4.42, match_type = 'sw_sp_w', gapfill = 'sp_waterbody', .after = 1) %>%
  add_row(rgn_id = 67, rgn_name = 'Libya', species = 'Atlantic bluefin tuna', score = 4.42, match_type = 'sw_sp_w', gapfill = 'sp_waterbody', .after = 1) %>%
  add_row(rgn_id = 68, rgn_name = 'Malta', species = 'Atlantic bluefin tuna', score = 4.42, match_type = 'sw_sp_w', gapfill = 'sp_waterbody', .after = 1) %>%
  add_row(rgn_id = 185, rgn_name = 'Monaco', species = 'Atlantic bluefin tuna', score = 4.42, match_type = 'sw_sp_w', gapfill = 'sp_waterbody', .after = 1) %>%
  add_row(rgn_id = 186, rgn_name = 'Montenegro', species = 'Atlantic bluefin tuna', score = 4.42, match_type = 'sw_sp_w', gapfill = 'sp_waterbody', .after = 1) %>%
  add_row(rgn_id = 62, rgn_name = 'Morocco', species = 'Atlantic bluefin tuna', score = 4.42, match_type = 'sw_sp_w', gapfill = 'sp_waterbody', .after = 1) %>%
  add_row(rgn_id = 188, rgn_name = 'Slovenia', species = 'Atlantic bluefin tuna', score = 4.42, match_type = 'sw_sp_w', gapfill = 'sp_waterbody', .after = 1) %>%
  add_row(rgn_id = 182, rgn_name = 'Spain', species = 'Atlantic bluefin tuna', score = 4.42, match_type = 'sw_sp_w', gapfill = 'sp_waterbody', .after = 1) %>%
  add_row(rgn_id = 77, rgn_name = 'Syria', species = 'Atlantic bluefin tuna', score = 4.42, match_type = 'sw_sp_w', gapfill = 'sp_waterbody', .after = 1) %>%
  add_row(rgn_id = 61, rgn_name = 'Tunisia', species = 'Atlantic bluefin tuna', score = 4.42, match_type = 'sw_sp_w', gapfill = 'sp_waterbody', .after = 1) %>%
  add_row(rgn_id = 76, rgn_name = 'Turkey', species = 'Atlantic bluefin tuna', score = 4.42, match_type = 'sw_sp_w', gapfill = 'sp_waterbody', .after = 1) %>% 

  # Gilthead seabream - Marine net pen
  add_row(rgn_id = 82, rgn_name = 'Albania', species = 'Gilthead seabream', score = 4.42, match_type = 'sw_sp_w', gapfill = 'sp_waterbody', .after = 23) %>% 
  add_row(rgn_id = 84, rgn_name = 'Algeria', species = 'Gilthead seabream', score = 4.42, match_type = 'sw_sp_w', gapfill = 'sp_waterbody', .after = 23) %>%
  add_row(rgn_id = 232, rgn_name = 'Bosnia and Herzegovina', species = 'Gilthead seabream', score = 4.42, match_type = 'sw_sp_w', gapfill = 'sp_waterbody', .after = 23) %>%
  add_row(rgn_id = 187, rgn_name = 'Croatia', species = 'Gilthead seabream', score = 4.42, match_type = 'sw_sp_w', gapfill = 'sp_waterbody', .after = 23) %>%
  add_row(rgn_id = 81, rgn_name = 'Cyprus', species = 'Gilthead seabream', score = 4.42, match_type = 'sw_sp_w', gapfill = 'sp_waterbody', .after = 23) %>%
  add_row(rgn_id = 214, rgn_name = 'Egypt', species = 'Gilthead seabream', score = 4.42, match_type = 'sw_sp_w', gapfill = 'sp_waterbody', .after = 23) %>%
  add_row(rgn_id = 179, rgn_name = 'France', species = 'Gilthead seabream', score = 4.42, match_type = 'sw_sp_w', gapfill = 'sp_waterbody', .after = 23) %>%
  add_row(rgn_id = 80, rgn_name = 'Greece', species = 'Gilthead seabream', score = 4.42, match_type = 'sw_sp_w', gapfill = 'sp_waterbody', .after = 23) %>%
  add_row(rgn_id = 79, rgn_name = 'Israel', species = 'Gilthead seabream', score = 4.42, match_type = 'sw_sp_w', gapfill = 'sp_waterbody', .after = 23) %>%
  add_row(rgn_id = 184, rgn_name = 'Italy', species = 'Gilthead seabream', score = 4.42, match_type = 'sw_sp_w', gapfill = 'sp_waterbody', .after = 23) %>%
  add_row(rgn_id = 78, rgn_name = 'Lebanon', species = 'Gilthead seabream', score = 4.42, match_type = 'sw_sp_w', gapfill = 'sp_waterbody', .after = 23) %>%
  add_row(rgn_id = 67, rgn_name = 'Libya', species = 'Gilthead seabream', score = 4.42, match_type = 'sw_sp_w', gapfill = 'sp_waterbody', .after = 23) %>%
  add_row(rgn_id = 68, rgn_name = 'Malta', species = 'Gilthead seabream', score = 4.42, match_type = 'sw_sp_w', gapfill = 'sp_waterbody', .after = 23) %>%
  add_row(rgn_id = 185, rgn_name = 'Monaco', species = 'Gilthead seabream', score = 4.42, match_type = 'sw_sp_w', gapfill = 'sp_waterbody', .after = 23) %>%
  add_row(rgn_id = 186, rgn_name = 'Montenegro', species = 'Gilthead seabream', score = 4.42, match_type = 'sw_sp_w', gapfill = 'sp_waterbody', .after = 23) %>%
  add_row(rgn_id = 62, rgn_name = 'Morocco', species = 'Gilthead seabream', score = 4.42, match_type = 'sw_sp_w', gapfill = 'sp_waterbody', .after = 23) %>%
  add_row(rgn_id = 188, rgn_name = 'Slovenia', species = 'Gilthead seabream', score = 4.42, match_type = 'sw_sp_w', gapfill = 'sp_waterbody', .after = 23) %>%
  add_row(rgn_id = 182, rgn_name = 'Spain', species = 'Gilthead seabream', score = 4.42, match_type = 'sw_sp_w', gapfill = 'sp_waterbody', .after = 23) %>%
  add_row(rgn_id = 77, rgn_name = 'Syria', species = 'Gilthead seabream', score = 4.42, match_type = 'sw_sp_w', gapfill = 'sp_waterbody', .after = 23) %>%
  add_row(rgn_id = 61, rgn_name = 'Tunisia', species = 'Gilthead seabream', score = 4.42, match_type = 'sw_sp_w', gapfill = 'sp_waterbody', .after = 23) %>%
  add_row(rgn_id = 76, rgn_name = 'Turkey', species = 'Gilthead seabream', score = 4.42, match_type = 'sw_sp_w', gapfill = 'sp_waterbody', .after = 23) %>% 
  
  # Meagre - Marine net pen
  add_row(rgn_id = 82, rgn_name = 'Albania', species = 'Meagre', score = 4.29, match_type = 'sw_sp_w', gapfill = 'sp_waterbody', .after = 45) %>% 
  add_row(rgn_id = 84, rgn_name = 'Algeria', species = 'Meagre', score = 4.29, match_type = 'sw_sp_w', gapfill = 'sp_waterbody', .after = 45) %>%
  add_row(rgn_id = 232, rgn_name = 'Bosnia and Herzegovina', species = 'Meagre', score = 4.29, match_type = 'sw_sp_w', gapfill = 'sp_waterbody', .after = 45) %>%
  add_row(rgn_id = 187, rgn_name = 'Croatia', species = 'Meagre', score = 4.29, match_type = 'sw_sp_w', gapfill = 'sp_waterbody', .after = 45) %>%
  add_row(rgn_id = 81, rgn_name = 'Cyprus', species = 'Meagre', score = 4.29, match_type = 'sw_sp_w', gapfill = 'sp_waterbody', .after = 45) %>%
  add_row(rgn_id = 214, rgn_name = 'Egypt', species = 'Meagre', score = 4.29, match_type = 'sw_sp_w', gapfill = 'sp_waterbody', .after = 45) %>%
  add_row(rgn_id = 179, rgn_name = 'France', species = 'Meagre', score = 4.29, match_type = 'sw_sp_w', gapfill = 'sp_waterbody', .after = 45) %>%
  add_row(rgn_id = 80, rgn_name = 'Greece', species = 'Meagre', score = 4.29, match_type = 'sw_sp_w', gapfill = 'sp_waterbody', .after = 45) %>%
  add_row(rgn_id = 79, rgn_name = 'Israel', species = 'Meagre', score = 4.29, match_type = 'sw_sp_w', gapfill = 'sp_waterbody', .after = 45) %>%
  add_row(rgn_id = 184, rgn_name = 'Italy', species = 'Meagre', score = 4.29, match_type = 'sw_sp_w', gapfill = 'sp_waterbody', .after = 45) %>%
  add_row(rgn_id = 78, rgn_name = 'Lebanon', species = 'Meagre', score = 4.29, match_type = 'sw_sp_w', gapfill = 'sp_waterbody', .after = 45) %>%
  add_row(rgn_id = 67, rgn_name = 'Libya', species = 'Meagre', score = 4.29, match_type = 'sw_sp_w', gapfill = 'sp_waterbody', .after = 45) %>%
  add_row(rgn_id = 68, rgn_name = 'Malta', species = 'Meagre', score = 4.29, match_type = 'sw_sp_w', gapfill = 'sp_waterbody', .after = 45) %>%
  add_row(rgn_id = 185, rgn_name = 'Monaco', species = 'Meagre', score = 4.29, match_type = 'sw_sp_w', gapfill = 'sp_waterbody', .after = 45) %>%
  add_row(rgn_id = 186, rgn_name = 'Montenegro', species = 'Meagre', score = 4.29, match_type = 'sw_sp_w', gapfill = 'sp_waterbody', .after = 45) %>%
  add_row(rgn_id = 62, rgn_name = 'Morocco', species = 'Meagre', score = 4.29, match_type = 'sw_sp_w', gapfill = 'sp_waterbody', .after = 45) %>%
  add_row(rgn_id = 188, rgn_name = 'Slovenia', species = 'Meagre', score = 4.29, match_type = 'sw_sp_w', gapfill = 'sp_waterbody', .after = 45) %>%
  add_row(rgn_id = 182, rgn_name = 'Spain', species = 'Meagre', score = 4.29, match_type = 'sw_sp_w', gapfill = 'sp_waterbody', .after = 45) %>%
  add_row(rgn_id = 77, rgn_name = 'Syria', species = 'Meagre', score = 4.29, match_type = 'sw_sp_w', gapfill = 'sp_waterbody', .after = 45) %>%
  add_row(rgn_id = 61, rgn_name = 'Tunisia', species = 'Meagre', score = 4.29, match_type = 'sw_sp_w', gapfill = 'sp_waterbody', .after = 45) %>%
  add_row(rgn_id = 76, rgn_name = 'Turkey', species = 'Meagre', score = 4.29, match_type = 'sw_sp_w', gapfill = 'sp_waterbody', .after = 45) %>% 
  
# European seabass - Marine net pen
  add_row(rgn_id = 82, rgn_name = 'Albania', species = 'European seabass', score = 4.45, match_type = 'sw_sp_w', gapfill = 'sp_waterbody', .after = 67) %>% 
  add_row(rgn_id = 84, rgn_name = 'Algeria', species = 'European seabass', score = 4.45, match_type = 'sw_sp_w', gapfill = 'sp_waterbody', .after = 67) %>%
  add_row(rgn_id = 232, rgn_name = 'Bosnia and Herzegovina', species = 'European seabass', score = 4.45, match_type = 'sw_sp_w', gapfill = 'sp_waterbody', .after = 67) %>%
  add_row(rgn_id = 187, rgn_name = 'Croatia', species = 'European seabass', score = 4.45, match_type = 'sw_sp_w', gapfill = 'sp_waterbody', .after = 67) %>%
  add_row(rgn_id = 81, rgn_name = 'Cyprus', species = 'European seabass', score = 4.45, match_type = 'sw_sp_w', gapfill = 'sp_waterbody', .after = 67) %>%
  add_row(rgn_id = 214, rgn_name = 'Egypt', species = 'European seabass', score = 4.45, match_type = 'sw_sp_w', gapfill = 'sp_waterbody', .after = 67) %>%
  add_row(rgn_id = 179, rgn_name = 'France', species = 'European seabass', score = 4.45, match_type = 'sw_sp_w', gapfill = 'sp_waterbody', .after = 67) %>%
  add_row(rgn_id = 80, rgn_name = 'Greece', species = 'European seabass', score = 4.45, match_type = 'sw_sp_w', gapfill = 'sp_waterbody', .after = 67) %>%
  add_row(rgn_id = 79, rgn_name = 'Israel', species = 'European seabass', score = 4.45, match_type = 'sw_sp_w', gapfill = 'sp_waterbody', .after = 67) %>%
  add_row(rgn_id = 184, rgn_name = 'Italy', species = 'European seabass', score = 4.45, match_type = 'sw_sp_w', gapfill = 'sp_waterbody', .after = 67) %>%
  add_row(rgn_id = 78, rgn_name = 'Lebanon', species = 'European seabass', score = 4.45, match_type = 'sw_sp_w', gapfill = 'sp_waterbody', .after = 67) %>%
  add_row(rgn_id = 67, rgn_name = 'Libya', species = 'European seabass', score = 4.45, match_type = 'sw_sp_w', gapfill = 'sp_waterbody', .after = 67) %>%
  add_row(rgn_id = 68, rgn_name = 'Malta', species = 'European seabass', score = 4.45, match_type = 'sw_sp_w', gapfill = 'sp_waterbody', .after = 67) %>%
  add_row(rgn_id = 185, rgn_name = 'Monaco', species = 'European seabass', score = 4.45, match_type = 'sw_sp_w', gapfill = 'sp_waterbody', .after = 67) %>%
  add_row(rgn_id = 186, rgn_name = 'Montenegro', species = 'European seabass', score = 4.45, match_type = 'sw_sp_w', gapfill = 'sp_waterbody', .after = 67) %>%
  add_row(rgn_id = 62, rgn_name = 'Morocco', species = 'European seabass', score = 4.45, match_type = 'sw_sp_w', gapfill = 'sp_waterbody', .after = 67) %>%
  add_row(rgn_id = 188, rgn_name = 'Slovenia', species = 'European seabass', score = 4.45, match_type = 'sw_sp_w', gapfill = 'sp_waterbody', .after = 67) %>%
  add_row(rgn_id = 182, rgn_name = 'Spain', species = 'European seabass', score = 4.45, match_type = 'sw_sp_w', gapfill = 'sp_waterbody', .after = 67) %>%
  add_row(rgn_id = 77, rgn_name = 'Syria', species = 'European seabass', score = 4.45, match_type = 'sw_sp_w', gapfill = 'sp_waterbody', .after = 67) %>%
  add_row(rgn_id = 61, rgn_name = 'Tunisia', species = 'European seabass', score = 4.45, match_type = 'sw_sp_w', gapfill = 'sp_waterbody', .after = 67) %>%
  add_row(rgn_id = 76, rgn_name = 'Turkey', species = 'European seabass', score = 4.45, match_type = 'sw_sp_w', gapfill = 'sp_waterbody', .after = 67)

```

```{r gf_sw_sp_g}

## Research species according to method and create new rows matching up species score with regions that are found to produce that type of aquaculture
gf_sw_sp_g <- sw_sus_rgn %>% 
  filter(match_type == 'sw_sp_g') %>% 
  
  # Atlantic cod - Indoor recirculating tank (without wastewater treatment)
  add_row(rgn_id = 218, rgn_name = 'Canada', species = 'Atlantic cod', score = 6.61, match_type = 'sw_sp_g', gapfill = 'sp_global', .after = 1) %>% # Source: http://www.fishbase.us/report/fao/FAOAquacultureList.php?scientific=Gadus+morhua
  add_row(rgn_id = 143, rgn_name = 'Iceland', species = 'Atlantic cod', score = 6.61, match_type = 'sw_sp_g', gapfill = 'sp_global', .after = 1) %>% # Source: http://www.fishbase.us/report/fao/FAOAquacultureList.php?scientific=Gadus+morhua
  add_row(rgn_id = 223, rgn_name = 'Norway', species = 'Atlantic cod', score = 6.61, match_type = 'sw_sp_g', gapfill = 'sp_global', .after = 1) %>% # Source: http://www.fishbase.us/report/fao/FAOAquacultureList.php?scientific=Gadus+morhua
  add_row(rgn_id = 180, rgn_name = 'United Kingdom', species = 'Atlantic cod', score = 6.61, match_type = 'sw_sp_g', gapfill = 'sp_global', .after = 1) %>% # Source: http://www.fishbase.us/report/fao/FAOAquacultureList.php?scientific=Gadus+morhua
  
  # Mussels - Bottom culture (regions only include those listed in 'maric' for 'Sea mussels nei')
  add_row(rgn_id = 25, rgn_name = 'Thailand', species = 'Mussels', score = 8.10, match_type = 'sw_sp_g', gapfill = 'sp_global', .after = 6) %>% # Source: https://www.seafoodwatch.org/-/m/sfw/pdf/reports/m/mba_seafoodwatch_farmedmussels.pdf
  add_row(rgn_id = 73, rgn_name = 'Russia', species = 'Mussels', score = 8.10, match_type = 'sw_sp_g', gapfill = 'sp_global', .after = 6) %>% # Source: https://www.seafoodwatch.org/-/m/sfw/pdf/reports/m/mba_seafoodwatch_farmedmussels.pdf
  add_row(rgn_id = 102, rgn_name = 'South Africa', species = 'Mussels', score = 8.10, match_type = 'sw_sp_g', gapfill = 'sp_global', .after = 6) %>% # Source: https://www.seafoodwatch.org/-/m/sfw/pdf/reports/m/mba_seafoodwatch_farmedmussels.pdf
  add_row(rgn_id = 135, rgn_name = 'Mexico', species = 'Mussels', score = 8.10, match_type = 'sw_sp_g', gapfill = 'sp_global', .after = 6) %>% # Source: https://www.seafoodwatch.org/-/m/sfw/pdf/reports/m/mba_seafoodwatch_farmedmussels.pdf
  add_row(rgn_id = 136, rgn_name = 'Guatemala', species = 'Mussels', score = 8.10, match_type = 'sw_sp_g', gapfill = 'sp_global', .after = 6) %>% # Source: https://www.seafoodwatch.org/-/m/sfw/pdf/reports/m/mba_seafoodwatch_farmedmussels.pdf
  add_row(rgn_id = 147, rgn_name = 'French Polynesia', species = 'Mussels', score = 8.10, match_type = 'sw_sp_g', gapfill = 'sp_global', .after = 6) %>% # Source: https://www.seafoodwatch.org/-/m/sfw/pdf/reports/m/mba_seafoodwatch_farmedmussels.pdf
  add_row(rgn_id = 180, rgn_name = 'United Kingdom', species = 'Mussels', score = 8.10, match_type = 'sw_sp_g', gapfill = 'sp_global', .after = 6) %>% # Source: https://www.seafoodwatch.org/-/m/sfw/pdf/reports/m/mba_seafoodwatch_farmedmussels.pdf
  add_row(rgn_id = 182, rgn_name = 'Spain', species = 'Mussels', score = 8.10, match_type = 'sw_sp_g', gapfill = 'sp_global', .after = 6) %>% # Source: https://www.seafoodwatch.org/-/m/sfw/pdf/reports/m/mba_seafoodwatch_farmedmussels.pdf
  add_row(rgn_id = 183, rgn_name = 'Portugal', species = 'Mussels', score = 8.10, match_type = 'sw_sp_g', gapfill = 'sp_global', .after = 6) %>% # Source: https://www.seafoodwatch.org/-/m/sfw/pdf/reports/m/mba_seafoodwatch_farmedmussels.pdf
  add_row(rgn_id = 209, rgn_name = 'China', species = 'Mussels', score = 8.10, match_type = 'sw_sp_g', gapfill = 'sp_global', .after = 6) %>% # Source: https://www.seafoodwatch.org/-/m/sfw/pdf/reports/m/mba_seafoodwatch_farmedmussels.pdf
  add_row(rgn_id = 219, rgn_name = 'Saint Pierre and Miquelon', species = 'Mussels', score = 8.10, match_type = 'sw_sp_g', gapfill = 'sp_global', .after = 6) %>% # Source: https://www.seafoodwatch.org/-/m/sfw/pdf/reports/m/mba_seafoodwatch_farmedmussels.pdf
  
  # Basket Cockle - Off-bottom culture
  add_row(rgn_id = 163, rgn_name = 'United States', species = 'Basket Cockle', score = 7.00, match_type = 'sw_sp_g', gapfill = 'sp_global', .after = 18) %>% # Source: https://www.seafoodwatch.org/-/m/c1eb959a63b645b9b8ddefff5cd6967f.pdf
  add_row(rgn_id = 209, rgn_name = 'China', species = 'Basket Cockle', score = 7.00, match_type = 'sw_sp_g', gapfill = 'sp_global', .after = 18) %>% # Source: https://www.seafoodwatch.org/-/m/c1eb959a63b645b9b8ddefff5cd6967f.pdf
  add_row(rgn_id = 218, rgn_name = 'Canada', species = 'Basket Cockle', score = 7.00, match_type = 'sw_sp_g', gapfill = 'sp_global', .after = 18) %>% # Source: https://www.seafoodwatch.org/-/m/c1eb959a63b645b9b8ddefff5cd6967f.pdf
  add_row(rgn_id = 207, rgn_name = 'Vietnam', species = 'Basket Cockle', score = 7.00, match_type = 'sw_sp_g', gapfill = 'sp_global', .after = 18) %>% # Source: https://www.seafoodwatch.org/-/m/c1eb959a63b645b9b8ddefff5cd6967f.pdf
  add_row(rgn_id = 20, rgn_name = 'South Korea', species = 'Basket Cockle', score = 7.00, match_type = 'sw_sp_g', gapfill = 'sp_global', .after = 18) %>% # Source: https://www.seafoodwatch.org/-/m/c1eb959a63b645b9b8ddefff5cd6967f.pdf
  add_row(rgn_id = 210, rgn_name = 'Japan', species = 'Basket Cockle', score = 7.00, match_type = 'sw_sp_g', gapfill = 'sp_global', .after = 18) %>% # Source: https://www.seafoodwatch.org/-/m/c1eb959a63b645b9b8ddefff5cd6967f.pdf
  
  # Abalones nei - Open bottom culture
  add_row(rgn_id = 210, rgn_name = 'Japan', species = 'Abalones nei', score = 6.23, match_type = 'sw_sp_g', gapfill = 'sp_global', .after = 25) %>% # Source: https://www.seafoodwatch.org/-/m/3894b7bcd4e34144b85179d9455ffa91.pdf
  add_row(rgn_id = 163, rgn_name = 'United States', species = 'Abalones nei', score = 6.23, match_type = 'sw_sp_g', gapfill = 'sp_global', .after = 25) %>% # Source: https://www.seafoodwatch.org/-/m/3894b7bcd4e34144b85179d9455ffa91.pdf
  add_row(rgn_id = 209, rgn_name = 'China', species = 'Abalones nei', score = 6.23, match_type = 'sw_sp_g', gapfill = 'sp_global', .after = 25) %>% # Source: https://www.seafoodwatch.org/-/m/3894b7bcd4e34144b85179d9455ffa91.pdf
  add_row(rgn_id = 16, rgn_name = 'Australia', species = 'Abalones nei', score = 6.23, match_type = 'sw_sp_g', gapfill = 'sp_global', .after = 25) %>% # Source: https://www.seafoodwatch.org/-/m/3894b7bcd4e34144b85179d9455ffa91.pdf
  add_row(rgn_id = 135, rgn_name = 'Mexico', species = 'Abalones nei', score = 6.23, match_type = 'sw_sp_g', gapfill = 'sp_global', .after = 25) %>% # Source: https://www.seafoodwatch.org/-/m/3894b7bcd4e34144b85179d9455ffa91.pdf
  add_row(rgn_id = 224, rgn_name = 'Chile', species = 'Abalones nei', score = 6.23, match_type = 'sw_sp_g', gapfill = 'sp_global', .after = 25) %>% # Source: https://www.seafoodwatch.org/-/m/3894b7bcd4e34144b85179d9455ffa91.pdf
  add_row(rgn_id = 162, rgn_name = 'New Zealand', species = 'Abalones nei', score = 6.23, match_type = 'sw_sp_g', gapfill = 'sp_global', .after = 25) %>% # Source: https://www.seafoodwatch.org/-/m/3894b7bcd4e34144b85179d9455ffa91.pdf
  add_row(rgn_id = 102, rgn_name = 'South Africa', species = 'Abalones nei', score = 6.23, match_type = 'sw_sp_g', gapfill = 'sp_global', .after = 25) %>% # Source: https://www.seafoodwatch.org/-/m/3894b7bcd4e34144b85179d9455ffa91.pdf
  
  # European seabass - Indoor recirculating tank (without wastewater treatment)	
  add_row(rgn_id = 179, rgn_name = 'France', species = 'European seabass', score = 6.61, match_type = 'sw_sp_g', gapfill = 'sp_global', .after = 34) %>% # Source: http://www.fao.org/fishery/culturedspecies/Dicentrarchus_labrax/en
  add_row(rgn_id = 184, rgn_name = 'Italy', species = 'European seabass', score = 6.61, match_type = 'sw_sp_g', gapfill = 'sp_global', .after = 34) %>% # Source: http://www.fao.org/fishery/culturedspecies/Dicentrarchus_labrax/en
  add_row(rgn_id = 80, rgn_name = 'Greece', species = 'European seabass', score = 6.61, match_type = 'sw_sp_g', gapfill = 'sp_global', .after = 34) %>% # Source: http://www.fao.org/fishery/culturedspecies/Dicentrarchus_labrax/en
  add_row(rgn_id = 76, rgn_name = 'Turkey', species = 'European seabass', score = 6.61, match_type = 'sw_sp_g', gapfill = 'sp_global', .after = 34) %>% # Source: http://www.fao.org/fishery/culturedspecies/Dicentrarchus_labrax/en
  add_row(rgn_id = 182, rgn_name = 'Spain', species = 'European seabass', score = 6.61, match_type = 'sw_sp_g', gapfill = 'sp_global', .after = 34) %>% # Source: http://www.fao.org/fishery/culturedspecies/Dicentrarchus_labrax/en
  add_row(rgn_id = 187, rgn_name = 'Croatia', species = 'European seabass', score = 6.61, match_type = 'sw_sp_g', gapfill = 'sp_global', .after = 34) %>% # Source: http://www.fao.org/fishery/culturedspecies/Dicentrarchus_labrax/en
  add_row(rgn_id = 214, rgn_name = 'Egypt', species = 'European seabass', score = 6.61, match_type = 'sw_sp_g', gapfill = 'sp_global', .after = 34) %>% # Source: http://www.fao.org/fishery/culturedspecies/Dicentrarchus_labrax/en
  add_row(rgn_id = 176, rgn_name = 'Germany', species = 'European seabass', score = 6.61, match_type = 'sw_sp_g', gapfill = 'sp_global', .after = 34) %>% # Source: http://www.fao.org/fishery/culturedspecies/Dicentrarchus_labrax/en
  add_row(rgn_id = 183, rgn_name = 'Portugal', species = 'European seabass', score = 6.61, match_type = 'sw_sp_g', gapfill = 'sp_global', .after = 34) %>% # Source: http://www.fao.org/fishery/culturedspecies/Dicentrarchus_labrax/en
  add_row(rgn_id = 188, rgn_name = 'Slovenia', species = 'European seabass', score = 6.61, match_type = 'sw_sp_g', gapfill = 'sp_global', .after = 34) %>% # Source: http://www.fao.org/fishery/culturedspecies/Dicentrarchus_labrax/en
  add_row(rgn_id = 232, rgn_name = 'Bosnia and Herzegovina', species = 'European seabass', score = 6.61, match_type = 'sw_sp_g', gapfill = 'sp_global', .after = 34) %>% # Source: http://www.fao.org/fishery/culturedspecies/Dicentrarchus_labrax/en
  add_row(rgn_id = 62, rgn_name = 'Morocco', species = 'European seabass', score = 6.61, match_type = 'sw_sp_g', gapfill = 'sp_global', .after = 34) %>% # Source: http://www.fao.org/fishery/culturedspecies/Dicentrarchus_labrax/en
  add_row(rgn_id = 84, rgn_name = 'Algeria', species = 'European seabass', score = 6.61, match_type = 'sw_sp_g', gapfill = 'sp_global', .after = 34) %>% # Source: http://www.fao.org/fishery/culturedspecies/Dicentrarchus_labrax/en
  add_row(rgn_id = 61, rgn_name = 'Tunisia', species = 'European seabass', score = 6.61, match_type = 'sw_sp_g', gapfill = 'sp_global', .after = 34) %>% # Source: http://www.fao.org/fishery/culturedspecies/Dicentrarchus_labrax/en
  add_row(rgn_id = 68, rgn_name = 'Malta', species = 'European seabass', score = 6.61, match_type = 'sw_sp_g', gapfill = 'sp_global', .after = 34) %>% # Source: http://www.fao.org/fishery/culturedspecies/Dicentrarchus_labrax/en
  add_row(rgn_id = 81, rgn_name = 'Cyprus', species = 'European seabass', score = 6.61, match_type = 'sw_sp_g', gapfill = 'sp_global', .after = 34) %>% # Source: http://www.fao.org/fishery/culturedspecies/Dicentrarchus_labrax/en
  add_row(rgn_id = 79, rgn_name = 'Israel', species = 'European seabass', score = 6.61, match_type = 'sw_sp_g', gapfill = 'sp_global', .after = 34) %>% # Source: http://www.fao.org/fishery/culturedspecies/Dicentrarchus_labrax/en
  
  # Sand gaper - Bottom culture
  add_row(rgn_id = 218, rgn_name = 'Canada', species = 'Sand gaper', score = 7.00, match_type = 'sw_sp_g', gapfill = 'sp_global', .after = 52) %>% # Source: https://www.seafoodwatch.org/-/m/b564a45ea2cb4034b73cdb4b106b4b86.pdf
  add_row(rgn_id = 163, rgn_name = 'United States', species = 'Sand gaper', score = 7.00, match_type = 'sw_sp_g', gapfill = 'sp_global', .after = 52) %>% # Source: https://www.seafoodwatch.org/-/m/7e33b7917e0940d6b5d1e27f0711ae01.pdf
  
  # Blood cockle - Off-bottom culture
  add_row(rgn_id = 163, rgn_name = 'United States', species = 'Blood cockle', score = 7.00, match_type = 'sw_sp_g', gapfill = 'sp_global', .after = 55) %>% # Source: https://www.seafoodwatch.org/-/m/sfw/pdf/reports/c/mba_seafoodwatch_farmedclamsreport.pdf
  add_row(rgn_id = 209, rgn_name = 'China', species = 'Blood cockle', score = 7.00, match_type = 'sw_sp_g', gapfill = 'sp_global', .after = 55) %>% # Source: https://www.seafoodwatch.org/-/m/sfw/pdf/reports/c/mba_seafoodwatch_farmedclamsreport.pdf
  add_row(rgn_id = 218, rgn_name = 'Canada', species = 'Blood cockle', score = 7.00, match_type = 'sw_sp_g', gapfill = 'sp_global', .after = 55) %>% # Source: https://www.seafoodwatch.org/-/m/sfw/pdf/reports/c/mba_seafoodwatch_farmedclamsreport.pdf
  add_row(rgn_id = 207, rgn_name = 'Vietnam', species = 'Blood cockle', score = 7.00, match_type = 'sw_sp_g', gapfill = 'sp_global', .after = 55) %>% # Source: https://www.seafoodwatch.org/-/m/sfw/pdf/reports/c/mba_seafoodwatch_farmedclamsreport.pdf
  
  # Rex sole - Indoor recirculating tank (with wastewater)
  add_row(rgn_id = 163, rgn_name = 'United States', species = 'Rex sole', score = 6.90, match_type = 'sw_sp_g', gapfill = 'sp_global', .after = 60) %>% # Source: https://www.seafoodwatch.org/seafood-recommendations/groups/sole?q=Rex%20Sole&t=Rex%20Sole&type=rex&method=farmed&o=284100531
  
  # Black sea bass - Indoor recirculating tank (with wastewater)
  add_row(rgn_id = 163, rgn_name = 'United States', species = 'Black sea bass', score = 6.90, match_type = 'sw_sp_g', gapfill = 'sp_global', .after = 62) %>% # Source: http://www.fishbase.us/report/fao/FAOAquacultureList.php?scientific=Centropristis+striata
  
  # Abalones nei - Indoor flowthrough tank
  add_row(rgn_id = 163, rgn_name = 'United States', species = 'Abalones nei', score = 6.95, match_type = 'sw_sp_g', gapfill = 'sp_global', .after = 64) %>% # Source: https://www.seafoodwatch.org/-/m/3894b7bcd4e34144b85179d9455ffa91.pdf
  add_row(rgn_id = 16, rgn_name = 'Australia', species = 'Abalones nei', score = 6.95, match_type = 'sw_sp_g', gapfill = 'sp_global', .after = 64) %>% # Source: https://www.seafoodwatch.org/-/m/3894b7bcd4e34144b85179d9455ffa91.pdf
  add_row(rgn_id = 209, rgn_name = 'China', species = 'Abalones nei', score = 6.95, match_type = 'sw_sp_g', gapfill = 'sp_global', .after = 64) %>% # Source: https://www.seafoodwatch.org/-/m/3894b7bcd4e34144b85179d9455ffa91.pdf
  add_row(rgn_id = 135, rgn_name = 'Mexico', species = 'Abalones nei', score = 6.95, match_type = 'sw_sp_g', gapfill = 'sp_global', .after = 64) %>% # Source: https://www.seafoodwatch.org/-/m/3894b7bcd4e34144b85179d9455ffa91.pdf
  add_row(rgn_id = 224, rgn_name = 'Chile', species = 'Abalones nei', score = 6.95, match_type = 'sw_sp_g', gapfill = 'sp_global', .after = 64) %>% # Source: https://www.seafoodwatch.org/-/m/3894b7bcd4e34144b85179d9455ffa91.pdf
  add_row(rgn_id = 210, rgn_name = 'Japan', species = 'Abalones nei', score = 6.95, match_type = 'sw_sp_g', gapfill = 'sp_global', .after = 64) %>% # Source: https://www.seafoodwatch.org/-/m/3894b7bcd4e34144b85179d9455ffa91.pdf
  add_row(rgn_id = 162, rgn_name = 'New Zealand', species = 'Abalones nei', score = 6.95, match_type = 'sw_sp_g', gapfill = 'sp_global', .after = 64) %>% # Source: https://www.seafoodwatch.org/-/m/3894b7bcd4e34144b85179d9455ffa91.pdf
  add_row(rgn_id = 102, rgn_name = 'South Africa', species = 'Abalones nei', score = 6.95, match_type = 'sw_sp_g', gapfill = 'sp_global', .after = 64) %>% # Source: https://www.seafoodwatch.org/-/m/3894b7bcd4e34144b85179d9455ffa91.pdf
  
  # Arctic char - Indoor recirculating tank (with wastewater)
  add_row(rgn_id = 175, rgn_name = 'Denmark', species = 'Arctic char', score = 6.90, match_type = 'sw_sp_g', gapfill = 'sp_global', .after = 73) %>% # Source: http://www.fishbase.us/report/fao/FAOAquacultureList.php?scientific=Salvelinus+alpinus
  add_row(rgn_id = 179, rgn_name = 'France', species = 'Arctic char', score = 6.90, match_type = 'sw_sp_g', gapfill = 'sp_global', .after = 73) %>% # Source: http://www.fishbase.us/report/fao/FAOAquacultureList.php?scientific=Salvelinus+alpinus
  add_row(rgn_id = 143, rgn_name = 'Iceland', species = 'Arctic char', score = 6.90, match_type = 'sw_sp_g', gapfill = 'sp_global', .after = 73) %>% # Source: http://www.fishbase.us/report/fao/FAOAquacultureList.php?scientific=Salvelinus+alpinus
  add_row(rgn_id = 181, rgn_name = 'Ireland', species = 'Arctic char', score = 6.90, match_type = 'sw_sp_g', gapfill = 'sp_global', .after = 73) %>% # Source: http://www.fishbase.us/report/fao/FAOAquacultureList.php?scientific=Salvelinus+alpinus
  add_row(rgn_id = 180, rgn_name = 'United Kingdom', species = 'Arctic char', score = 6.90, match_type = 'sw_sp_g', gapfill = 'sp_global', .after = 73) %>% # Source: http://www.fishbase.us/report/fao/FAOAquacultureList.php?scientific=Salvelinus+alpinus
  add_row(rgn_id = 163, rgn_name = 'United States', species = 'Arctic char', score = 6.90, match_type = 'sw_sp_g', gapfill = 'sp_global', .after = 73) %>% # Source: http://www.fishbase.us/report/fao/FAOAquacultureList.php?scientific=Salvelinus+alpinus
  
  # Chilean mussel - Off-bottom culture
  add_row(rgn_id = 224, rgn_name = 'Chile', species = 'Chilean mussel', score = 8.10, match_type = 'sw_sp_g', gapfill = 'sp_global', .after = 80) %>% # Source: https://www.seafoodwatch.org/-/m/sfw/pdf/reports/m/mba_seafoodwatch_farmedmussels.pdf
  
  # Lined scallop - Bottom culture
  add_row(rgn_id = 209, rgn_name = 'China', species = 'Lined scallop', score = 8.86, match_type = 'sw_sp_g', gapfill = 'sp_global', .after = 82) %>% # Source: https://www.seafoodwatch.org/-/m/sfw/pdf/reports/s/mba_seafoodwatch_farmedscallopsreport.pdf
  add_row(rgn_id = 218, rgn_name = 'Canada', species = 'Lined scallop', score = 8.86, match_type = 'sw_sp_g', gapfill = 'sp_global', .after = 82) %>% # Source: https://www.seafoodwatch.org/-/m/sfw/pdf/reports/s/mba_seafoodwatch_farmedscallopsreport.pdf
  add_row(rgn_id = 163, rgn_name = 'United States', species = 'Lined scallop', score = 8.86, match_type = 'sw_sp_g', gapfill = 'sp_global', .after = 82) %>% # Source: https://www.seafoodwatch.org/-/m/sfw/pdf/reports/s/mba_seafoodwatch_farmedscallopsreport.pdf
  add_row(rgn_id = 171, rgn_name = 'Brazil', species = 'Lined scallop', score = 8.86, match_type = 'sw_sp_g', gapfill = 'sp_global', .after = 82) %>% # Source: https://www.seafoodwatch.org/-/m/sfw/pdf/reports/s/mba_seafoodwatch_farmedscallopsreport.pdf
  add_row(rgn_id = 137, rgn_name = 'Ecuador', species = 'Lined scallop', score = 8.86, match_type = 'sw_sp_g', gapfill = 'sp_global', .after = 82) %>% # Source: https://www.seafoodwatch.org/-/m/sfw/pdf/reports/s/mba_seafoodwatch_farmedscallopsreport.pdf
  add_row(rgn_id = 223, rgn_name = 'Norway', species = 'Lined scallop', score = 8.86, match_type = 'sw_sp_g', gapfill = 'sp_global', .after = 82) %>% # Source: https://www.seafoodwatch.org/-/m/sfw/pdf/reports/s/mba_seafoodwatch_farmedscallopsreport.pdf
  
  # Razor clam - Bottom culture
  add_row(rgn_id = 163, rgn_name = 'United States', species = 'Razor clam', score = 7.00, match_type = 'sw_sp_g', gapfill = 'sp_global', .after = 89) %>% # Source: https://www.seafoodwatch.org/-/m/sfw/pdf/reports/c/mba_seafoodwatch_farmedclamsreport.pdf
  add_row(rgn_id = 209, rgn_name = 'China', species = 'Razor clam', score = 7.00, match_type = 'sw_sp_g', gapfill = 'sp_global', .after = 89) %>% # Source: https://www.seafoodwatch.org/-/m/sfw/pdf/reports/c/mba_seafoodwatch_farmedclamsreport.pdf
  add_row(rgn_id = 218, rgn_name = 'Canada', species = 'Razor clam', score = 7.00, match_type = 'sw_sp_g', gapfill = 'sp_global', .after = 89) %>% # Source: https://www.seafoodwatch.org/-/m/sfw/pdf/reports/c/mba_seafoodwatch_farmedclamsreport.pdf
  add_row(rgn_id = 207, rgn_name = 'Vietnam', species = 'Razor clam', score = 7.00, match_type = 'sw_sp_g', gapfill = 'sp_global', .after = 89) %>% # Source: https://www.seafoodwatch.org/-/m/sfw/pdf/reports/c/mba_seafoodwatch_farmedclamsreport.pdf
  add_row(rgn_id = 20, rgn_name = 'South Korea', species = 'Razor clam', score = 7.00, match_type = 'sw_sp_g', gapfill = 'sp_global', .after = 89) %>% # Source: https://www.seafoodwatch.org/-/m/sfw/pdf/reports/c/mba_seafoodwatch_farmedclamsreport.pdf
  add_row(rgn_id = 210, rgn_name = 'Japan', species = 'Razor clam', score = 7.00, match_type = 'sw_sp_g', gapfill = 'sp_global', .after = 89) %>% # Source: https://www.seafoodwatch.org/-/m/sfw/pdf/reports/c/mba_seafoodwatch_farmedclamsreport.pdf
  
  # NOT INCLUDED (because freshwater species): Giant river prawn - Indoor recirculating tank (without wastewater treatment)
  
  # Mussels - Off-bottom culture (regions only include those listed in 'maric' for 'Sea mussels nei')
  add_row(rgn_id = 25, rgn_name = 'Thailand', species = 'Mussels', score = 8.10, match_type = 'sw_sp_g', gapfill = 'sp_global', .after = 97) %>% # Source: https://www.seafoodwatch.org/-/m/sfw/pdf/reports/m/mba_seafoodwatch_farmedmussels.pdf
  add_row(rgn_id = 73, rgn_name = 'Russia', species = 'Mussels', score = 8.10, match_type = 'sw_sp_g', gapfill = 'sp_global', .after = 97) %>% # Source: https://www.seafoodwatch.org/-/m/sfw/pdf/reports/m/mba_seafoodwatch_farmedmussels.pdf
  add_row(rgn_id = 102, rgn_name = 'South Africa', species = 'Mussels', score = 8.10, match_type = 'sw_sp_g', gapfill = 'sp_global', .after = 97) %>% # Source: https://www.seafoodwatch.org/-/m/sfw/pdf/reports/m/mba_seafoodwatch_farmedmussels.pdf
  add_row(rgn_id = 135, rgn_name = 'Mexico', species = 'Mussels', score = 8.10, match_type = 'sw_sp_g', gapfill = 'sp_global', .after = 97) %>% # Source: https://www.seafoodwatch.org/-/m/sfw/pdf/reports/m/mba_seafoodwatch_farmedmussels.pdf
  add_row(rgn_id = 136, rgn_name = 'Guatemala', species = 'Mussels', score = 8.10, match_type = 'sw_sp_g', gapfill = 'sp_global', .after = 97) %>% # Source: https://www.seafoodwatch.org/-/m/sfw/pdf/reports/m/mba_seafoodwatch_farmedmussels.pdf
  add_row(rgn_id = 147, rgn_name = 'French Polynesia', species = 'Mussels', score = 8.10, match_type = 'sw_sp_g', gapfill = 'sp_global', .after = 97) %>% # Source: https://www.seafoodwatch.org/-/m/sfw/pdf/reports/m/mba_seafoodwatch_farmedmussels.pdf
  add_row(rgn_id = 180, rgn_name = 'United Kingdom', species = 'Mussels', score = 8.10, match_type = 'sw_sp_g', gapfill = 'sp_global', .after = 97) %>% # Source: https://www.seafoodwatch.org/-/m/sfw/pdf/reports/m/mba_seafoodwatch_farmedmussels.pdf
  add_row(rgn_id = 182, rgn_name = 'Spain', species = 'Mussels', score = 8.10, match_type = 'sw_sp_g', gapfill = 'sp_global', .after = 97) %>% # Source: https://www.seafoodwatch.org/-/m/sfw/pdf/reports/m/mba_seafoodwatch_farmedmussels.pdf
  add_row(rgn_id = 183, rgn_name = 'Portugal', species = 'Mussels', score = 8.10, match_type = 'sw_sp_g', gapfill = 'sp_global', .after = 97) %>% # Source: https://www.seafoodwatch.org/-/m/sfw/pdf/reports/m/mba_seafoodwatch_farmedmussels.pdf
  add_row(rgn_id = 209, rgn_name = 'China', species = 'Mussels', score = 8.10, match_type = 'sw_sp_g', gapfill = 'sp_global', .after = 97) %>% # Source: https://www.seafoodwatch.org/-/m/sfw/pdf/reports/m/mba_seafoodwatch_farmedmussels.pdf
  add_row(rgn_id = 219, rgn_name = 'Saint Pierre and Miquelon', species = 'Mussels', score = 8.10, match_type = 'sw_sp_g', gapfill = 'sp_global', .after = 97) %>% # Source: https://www.seafoodwatch.org/-/m/sfw/pdf/reports/m/mba_seafoodwatch_farmedmussels.pdf
  
  # Snubnose pompano - Indoor recirculating tank (with wastewater) (regions only include those listed in 'maric' for 'Snubnose pompano')
  add_row(rgn_id = 15, rgn_name = 'Philippines', species = 'Snubnose pompano', score = 6.90, match_type = 'sw_sp_g', gapfill = 'sp_global', .after = 109) %>% # Sources: https://www.seafoodwatch.org/seafood-recommendations/groups/pompano?type=snubnose&o=1409791156, http://www.fishbase.us/report/fao/FAOAquacultureList.php?scientific=Trachinotus+blochii
  add_row(rgn_id = 206, rgn_name = 'Malaysia', species = 'Snubnose pompano', score = 6.90, match_type = 'sw_sp_g', gapfill = 'sp_global', .after = 109) %>% # Sources: https://www.seafoodwatch.org/seafood-recommendations/groups/pompano?type=snubnose&o=1409791156, http://www.fishbase.us/report/fao/FAOAquacultureList.php?scientific=Trachinotus+blochii
  add_row(rgn_id = 208, rgn_name = 'Singapore', species = 'Snubnose pompano', score = 6.90, match_type = 'sw_sp_g', gapfill = 'sp_global', .after = 109) %>% # Sources: https://www.seafoodwatch.org/seafood-recommendations/groups/pompano?type=snubnose&o=1409791156, http://www.fishbase.us/report/fao/FAOAquacultureList.php?scientific=Trachinotus+blochii
  add_row(rgn_id = 209, rgn_name = 'China', species = 'Snubnose pompano', score = 6.90, match_type = 'sw_sp_g', gapfill = 'sp_global', .after = 109) %>% # Sources: https://www.seafoodwatch.org/seafood-recommendations/groups/pompano?type=snubnose&o=1409791156, http://www.fishbase.us/report/fao/FAOAquacultureList.php?scientific=Trachinotus+blochii
  
  # Hard clam (unspecified) - Off-bottom culture
  add_row(rgn_id = 163, rgn_name = 'United States', species = 'Hard clam (unspecified)', score = 7.00, match_type = 'sw_sp_g', gapfill = 'sp_global', .after = 114) %>% # Source: https://www.seafoodwatch.org/-/m/sfw/pdf/reports/c/mba_seafoodwatch_farmedclamsreport.pdf
  add_row(rgn_id = 209, rgn_name = 'China', species = 'Hard clam (unspecified)', score = 7.00, match_type = 'sw_sp_g', gapfill = 'sp_global', .after = 114) %>% # Source: https://www.seafoodwatch.org/-/m/sfw/pdf/reports/c/mba_seafoodwatch_farmedclamsreport.pdf
  add_row(rgn_id = 218, rgn_name = 'Canada', species = 'Hard clam (unspecified)', score = 7.00, match_type = 'sw_sp_g', gapfill = 'sp_global', .after = 114) %>% # Source: https://www.seafoodwatch.org/-/m/sfw/pdf/reports/c/mba_seafoodwatch_farmedclamsreport.pdf
  add_row(rgn_id = 207, rgn_name = 'Vietnam', species = 'Hard clam (unspecified)', score = 7.00, match_type = 'sw_sp_g', gapfill = 'sp_global', .after = 114) %>% # Source: https://www.seafoodwatch.org/-/m/sfw/pdf/reports/c/mba_seafoodwatch_farmedclamsreport.pdf
  add_row(rgn_id = 20, rgn_name = 'South Korea', species = 'Hard clam (unspecified)', score = 7.00, match_type = 'sw_sp_g', gapfill = 'sp_global', .after = 114) %>% # Source: https://www.seafoodwatch.org/-/m/sfw/pdf/reports/c/mba_seafoodwatch_farmedclamsreport.pdf
  add_row(rgn_id = 210, rgn_name = 'Japan', species = 'Hard clam (unspecified)', score = 7.00, match_type = 'sw_sp_g', gapfill = 'sp_global', .after = 114) %>% # Source: https://www.seafoodwatch.org/-/m/sfw/pdf/reports/c/mba_seafoodwatch_farmedclamsreport.pdf
  
  # Striped bass - Indoor recirculating tank (with wastewater)
  add_row(rgn_id = 135, rgn_name = 'Mexico', species = 'Striped bass', score = 6.90, match_type = 'sw_sp_g', gapfill = 'sp_global', .after = 121) %>% # Sources: http://www.fishbase.us/report/fao/FAOAquacultureList.php?scientific=Morone+saxatilis, https://www.seafoodwatch.org/seafood-recommendations/groups/bass?type=striped&o=620802890
  
  # Beluga - Indoor recirculating tank (with wastewater)
  add_row(rgn_id = 71, rgn_name = 'Bulgaria', species = 'Beluga', score = 6.90, match_type = 'sw_sp_g', gapfill = 'sp_global', .after = 123) %>% # Sources: http://www.fishbase.us/report/fao/FAOAquacultureList.php?scientific=Huso+huso, https://www.seafoodwatch.org/seafood-recommendations/groups/sturgeon?q=Sturgeon&location=domestic&type=beluga&o=983567610
  add_row(rgn_id = 75, rgn_name = 'Ukraine', species = 'Beluga', score = 6.90, match_type = 'sw_sp_g', gapfill = 'sp_global', .after = 123) %>% # Sources: http://www.fishbase.us/report/fao/FAOAquacultureList.php?scientific=Huso+huso, https://www.seafoodwatch.org/seafood-recommendations/groups/sturgeon?q=Sturgeon&location=domestic&type=beluga&o=983567610
  
  # NOT INCLUDED (because freshwater species): Giant river prawn - Indoor recirculating tank (with wastewater)

  # Venus clam - Off-bottom culture
  add_row(rgn_id = 163, rgn_name = 'United States', species = 'Venus clam', score = 7.00, match_type = 'sw_sp_g', gapfill = 'sp_global', .after = 127) %>% # Source: https://www.seafoodwatch.org/-/m/sfw/pdf/reports/c/mba_seafoodwatch_farmedclamsreport.pdf
  add_row(rgn_id = 209, rgn_name = 'China', species = 'Venus clam', score = 7.00, match_type = 'sw_sp_g', gapfill = 'sp_global', .after = 127) %>% # Source: https://www.seafoodwatch.org/-/m/sfw/pdf/reports/c/mba_seafoodwatch_farmedclamsreport.pdf
  add_row(rgn_id = 218, rgn_name = 'Canada', species = 'Venus clam', score = 7.00, match_type = 'sw_sp_g', gapfill = 'sp_global', .after = 127) %>% # Source: https://www.seafoodwatch.org/-/m/sfw/pdf/reports/c/mba_seafoodwatch_farmedclamsreport.pdf
  add_row(rgn_id = 207, rgn_name = 'Vietnam', species = 'Venus clam', score = 7.00, match_type = 'sw_sp_g', gapfill = 'sp_global', .after = 127) %>% # Source: https://www.seafoodwatch.org/-/m/sfw/pdf/reports/c/mba_seafoodwatch_farmedclamsreport.pdf
  add_row(rgn_id = 20, rgn_name = 'South Korea', species = 'Venus clam', score = 7.00, match_type = 'sw_sp_g', gapfill = 'sp_global', .after = 127) %>% # Source: https://www.seafoodwatch.org/-/m/sfw/pdf/reports/c/mba_seafoodwatch_farmedclamsreport.pdf
  add_row(rgn_id = 210, rgn_name = 'Japan', species = 'Venus clam', score = 7.00, match_type = 'sw_sp_g', gapfill = 'sp_global', .after = 127) %>% # Source: https://www.seafoodwatch.org/-/m/sfw/pdf/reports/c/mba_seafoodwatch_farmedclamsreport.pdf
  
  # Iceland scallop - Bottom culture
  add_row(rgn_id = 209, rgn_name = 'China', species = 'Iceland scallop', score = 8.86, match_type = 'sw_sp_g', gapfill = 'sp_global', .after = 134) %>% # Source: https://www.seafoodwatch.org/-/m/sfw/pdf/reports/s/mba_seafoodwatch_farmedscallopsreport.pdf
  add_row(rgn_id = 218, rgn_name = 'Canada', species = 'Iceland scallop', score = 8.86, match_type = 'sw_sp_g', gapfill = 'sp_global', .after = 134) %>% # Source: https://www.seafoodwatch.org/-/m/sfw/pdf/reports/s/mba_seafoodwatch_farmedscallopsreport.pdf
  add_row(rgn_id = 163, rgn_name = 'United States', species = 'Iceland scallop', score = 8.86, match_type = 'sw_sp_g', gapfill = 'sp_global', .after = 134) %>% # Source: https://www.seafoodwatch.org/-/m/sfw/pdf/reports/s/mba_seafoodwatch_farmedscallopsreport.pdf
  add_row(rgn_id = 171, rgn_name = 'Brazil', species = 'Iceland scallop', score = 8.86, match_type = 'sw_sp_g', gapfill = 'sp_global', .after = 134) %>% # Source: https://www.seafoodwatch.org/-/m/sfw/pdf/reports/s/mba_seafoodwatch_farmedscallopsreport.pdf
  add_row(rgn_id = 137, rgn_name = 'Ecuador', species = 'Iceland scallop', score = 8.86, match_type = 'sw_sp_g', gapfill = 'sp_global', .after = 134) %>% # Source: https://www.seafoodwatch.org/-/m/sfw/pdf/reports/s/mba_seafoodwatch_farmedscallopsreport.pdf
  add_row(rgn_id = 223, rgn_name = 'Norway', species = 'Iceland scallop', score = 8.86, match_type = 'sw_sp_g', gapfill = 'sp_global', .after = 134) %>% # Source: https://www.seafoodwatch.org/-/m/sfw/pdf/reports/s/mba_seafoodwatch_farmedscallopsreport.pdf
  
  # Meagre - Indoor recirculating tank (with wastewater)
  add_row(rgn_id = 183, rgn_name = 'Portugal', species = 'Meagre', score = 6.90, match_type = 'sw_sp_g', gapfill = 'sp_global', .after = 141) %>% # Source: https://www.seafoodwatch.org/seafood-recommendations/groups/meagre?o=1301679247 (NO: Egypt, Turkey, Croatia, Cyprus, France, Greece, Italy and Spain)
  
  # Japanese amberjack - Indoor recirculating tank (with wastewater)
  add_row(rgn_id = 20, rgn_name = 'South Korea', species = 'Japanese amberjack', score = 6.90, match_type = 'sw_sp_g', gapfill = 'sp_global', .after = 143) # Source: http://www.fishbase.us/report/fao/FAOAquacultureList.php?scientific=Seriola+quinqueradiata, https://www.seafoodwatch.org/seafood-recommendations/groups/amberjack?type=yellowtail&q=Amberjack,%20Yellowtail&o=262057062 (NO: Japan (farmed in marine net pens; but no way to differentiate))
  
  #
  #add_row(rgn_id = , rgn_name = '', species = '', score = , match_type = 'sw_sp_g', gapfill = 'sp_global', .after = ) %>% # Source:
  # NEEDS TO BE FINISHED
  


  ## Species that need a name change to match 'maric':
    # Mussels -> Sea mussels nei
    # Hard clam (unspecified) -> Clams, etc. nei (? doesn't specify hard clams though)
  
  
  #select(rgn_id, species, score, report_title, genus, spp, parent_method, method)

```

Append Seafood Watch sustainability scores data to the FAO mariculture data.

The following joins the Seafood Watch sustainability scores to regions/species that have FAO mariculture data.

```{r add-sw-sus-fao-data}

## Join data specific to species/country
sw_sp_c = sw_sus_rgn %>% 
  filter(match_type == 'sw_sp_c') %>% 
  select(rgn_id, species, Sust_sw_sp_c = score)

mar_sw_sus <- maric %>%
  left_join(sw_sp_c, by= c("species", "rgn_id"))


## Join data specific to species/water body
sw_sp_w = gf_sw_sp_w %>% 
  select(rgn_id, species, Sust_sw_sp_w = score)

mar_sw_sus <- mar_sw_sus %>%
  left_join(sw_sp_w, by= c("species", "rgn_id"))


## Join data specific to species/global Seafood Watch data
sw_sp_g = gf_sw_sp_g %>% 
  select(rgn_id, species, Sust_sw_sp_g = score)

mar_sw_sus <- mar_sw_sus %>%
  left_join(sw_sp_g, by= c("species", "rgn_id"))

```

Since some regions have multiple sustainability scores for the same species due to multiple aquaculture methods, but we don't know what proportions of which methods are used, we take the average of the sustainability scores in these instances.

```{r avg-sw-sus-types}

## Average sustainability scores of the three "Sust_" types:
mar_sw_sus <- mar_sw_sus %>% 
  dplyr::mutate(Sust_mean = rowMeans(dplyr::select(., starts_with("Sust")), na.rm = TRUE))

```

Merge the three Seafood Watch type categories into a single sustainability score column in the following order:

1. Sust_mean: average of the three "Sust_" columns
2. Sust_sw_sp_c: specific to species/country
3. Sust_sw_sp_w: specific to species/water body
4. Sust_sw_sp_g: specific to species/global Seafood Watch data

For example, if Sust_mean is missing, use Sust_sw_sp_c and so on. 

```{r sw-sus-merge}

mar_sw_sus = mar_sw_sus %>% 
  dplyr::mutate(Sust = ifelse(!is.na(Sust_mean), Sust_mean, Sust_sw_sp_c)) %>%
  dplyr::mutate(Sust = ifelse(is.na(Sust), Sust_sw_sp_w, Sust)) %>%
  dplyr::mutate(Sust = ifelse(is.na(Sust), Sust_sw_sp_g, Sust)) %>%
  dplyr::select(-Sust_mean, -Sust_sw_sp_c, -Sust_sw_sp_w, -Sust_sw_sp_g)

```

Average sustainability scores within regions with more than score (due to more than one aquaculture method):

```{r sw-sus-avg}

mar_sw_sus <- mar_sw_sus %>% 
  dplyr::group_by(rgn_id, species) %>% 
  dplyr::mutate(Sust_avg = mean(Sust, na.rm=TRUE)) %>% 
  dplyr::ungroup()

```

Get rid of duplicates for region/species/year:

```{r sw-sus-dup}

mar_sw_sus <- mar_sw_sus %>% 
  dplyr::distinct(rgn_id, species, environment, year, .keep_all = TRUE)
  # Went from 30682 obs. to 28507 obs.

```

**Now look at a summary after appending all the Seafood Watch data**

```{r sw-sus-summary}

summary(mar_sw_sus)
# Sust_avg still has 24951 NAs out of 28507 total obs.; there are 3556 obs. with sustainability data

```

**LAURA STOPPED HERE 8.28.2020**
- need to gapfill by averaging species within a taxonomic category


This joins the sustainability data that is gapfilled either at the species level (average of specific species/genera across regions) or at a higher course taxonomic levels and documents which data are gapfilled and how.

```{r sw-sus-join-gapfilled}

# ## Select observations gapfilled at the species/water body level:
# gf_sw_w_sus <- sw_sus_rgn %>% 
#   filter(gapfill != "actuals" & match_type == "sw_sp_w") %>%
#   select(species, gapfill, Sust_gf_sw_w = score)
# 
# ## Check that there are no duplicated species
# gf_sw_w_sus[duplicated(gf_sw_w_sus$species), ]
# 
# ## Match gapfilling values by species
# mar_sw_sus_gf = mar_sw_sus %>%
#   left_join(gf_sw_w_sus, by = 'species')
# 
# 
# ## Select observations gapfilled at the species/global level:
# gf_sw_g_sus <- sw_sus_rgn %>% 
#   filter(gapfill != "actuals" & match_type == "sw_sp_g") %>%
#   select(species, Sust_gf_sw_g = score)
# 
# ## Check that there are no duplicated species
# gf_sw_g_sus[duplicated(gf_sw_g_sus$species), ]
#   # There are duplicated species because multiple aquaculture methods are listed for multiple "global" species
# 
# ## Match gapfilling values by species
# mar_sw_sus_gf = mar_sw_sus_gf %>%
#   left_join(gf_sw_g_sus, by = 'species')

```

**Take a look at the wrangled data**

```{r view-sw-sus-gapfilled}

# summary(mar_sw_sus_gf)
# 
# table(mar_sw_sus_gf$gapfill)

```

Obtain a sustainability score for each record, and a book-keeping column of whether it's actual or gapfilled

For missing sustainability scores:

1. Use species-level Sustainability score
2. If no species-level scores, gapfill with genus-level sustainability average

```{r sw-sus-gapfill-missing}

# mar_sw_sus_final = mar_sw_sus_gf %>% 
#   mutate(gapfill = ifelse(!is.na(Sust), "none", gapfill)) %>%
#   mutate(Sust = ifelse(is.na(Sust), Sust_gf_sp, Sust)) %>% # Gapfill with species level
#   mutate(gapfill = ifelse(is.na(Sust), "taxon_average", gapfill)) %>% # Add in taxon gapfill record
#   mutate(Sust = ifelse(is.na(Sust), Sust_gf_taxon, Sust)) %>% # Gapfill with taxon level
#   mutate(taxa_code = paste(species, species_code, sep="_")) %>%
#   select(rgn_id, species, species_code, taxa_code, taxa_group=Taxon_code, year, gapfill_sus = gapfill, gapfill_fao = gap_0_fill, tonnes=value, Sust)

```

# Save Data:

```{r save-mar-sw-harvest-data}

# ## Save mariculture harvest data
# mar_harvest_tonnes = mar_sus_final %>%
#   select(rgn_id, taxa_code, taxa_group, year, tonnes)
# 
# anyDuplicated(mar_harvest_tonnes) # Check for duplication
# 
# ## Saves the appropriate Mariculture Harvest Tonnes file depending on whether we are excluding species based on best judgment (exclude), all seaweeds (exclude_no_seaweed), or just seaweeds nei (exclude_no_nei). See method changes above.
# if(sum(str_detect(names(mar_sp), "exclude$"))==1) {
#  write.csv(mar_harvest_tonnes, 'output/mar_harvest_tonnes.csv', row.names=F)
#  } else if (sum(str_detect(names(mar_sp), "exclude_no_seaweed"))==1) {
#    write.csv(mar_harvest_tonnes, 'test/mar_harvest_tonnes_no_seaweed.csv', row.names=F)
#    } else if (sum(str_detect(names(mar_sp), "exclude_no_nei"))==1) {
#      write.csv(mar_harvest_tonnes, 'test/mar_harvest_tonnes_no_nei.csv', row.names=F)
#      }
# 
# ## Save gapfill data for mariculture harvest
# mar_harvest_tonnes_gf = mar_sus_final %>%
#   select(rgn_id, taxa_code, taxa_group, year, tonnes=gapfill_fao)
# 
# 
# ## Saves the appropriate Mariculture Harvest Gapfill file depending on whether we are excluding species based on best judgment (exclude), all seaweeds (exclude_no_seaweed), or just seaweeds nei (exclude_no_nei). See method changes above
# if(sum(str_detect(names(mar_sp), "exclude$"))==1) {
#  write.csv(mar_harvest_tonnes_gf, 'output/mar_harvest_tonnes_gf.csv', row.names=F)
#  } else if (sum(str_detect(names(mar_sp), "exclude_no_seaweed"))==1) {
#    write.csv(mar_harvest_tonnes_gf, 'test/mar_harvest_tonnes_gf_no_seaweed.csv', row.names=F)
#    } else if (sum(str_detect(names(mar_sp), "exclude_no_nei"))==1) {
#      write.csv(mar_harvest_tonnes_gf, 'test/mar_harvest_tonnes_gf_no_nei.csv', row.names=F)
#      }
# 
# 
# ## Save sustainability scores data for 2012
# mar_sustainability_score = mar_sus_final %>% 
#   mutate(year = 2012) %>% # Only 2012 sustainability scores exist (Trujillo)
#   select(rgn_id, year, taxa_code, sust_coeff = Sust) %>% 
#   unique()
# 
# anyDuplicated(mar_sustainability_score)
# 
# 
# ## Saves the appropriate Sustainability Score file depending on whether we are excluding species based on best judgment (exclude), all seaweeds (exclude_no_seaweed), or just seaweeds nei (exclude_no_nei). See method changes above
# if(sum(str_detect(names(mar_sp), "exclude$"))==1) {
#  write.csv(mar_sustainability_score, 'output/mar_sustainability.csv', row.names=F)
#  } else if (sum(str_detect(names(mar_sp), "exclude_no_seaweed"))==1) {
#    write.csv(mar_sustainability_score, 'test/mar_sustainability_no_seaweed.csv', row.names=F)
#    } else if (sum(str_detect(names(mar_sp), "exclude_no_nei"))==1) {
#      write.csv(mar_sustainability_score, 'test/mar_sustainability_no_nei.csv', row.names=F)
#      }
# 
# 
# ## Save gapfill data for sustainability scores
# mar_sustainability_score_gf = mar_sus_final %>% 
#   select(rgn_id, taxa_code, sust_coeff = gapfill_sus) %>% 
#   unique()
# 
# 
# ## Saves the appropriate Sustainability Score Gapfill file depending on whether we are excluding species based on best judgment (exclude), all seaweeds (exclude_no_seaweed), or just seaweeds nei (exclude_no_nei). See method changes above
# if(sum(str_detect(names(mar_sp), "exclude$"))==1) {
#  write.csv(mar_sustainability_score_gf, 'output/mar_sustainability_gf.csv', row.names=F)
#  } else if (sum(str_detect(names(mar_sp), "exclude_no_seaweed"))==1) {
#    write.csv(mar_sustainability_score_gf, 'test/mar_sustainability_gf_no_seaweed.csv', row.names=F)
#    } else if (sum(str_detect(names(mar_sp), "exclude_no_nei"))==1) {
#      write.csv(mar_sustainability_score_gf, 'test/mar_sustainability_gf_no_nei.csv', row.names=F)
#      }

```



***


# Potential Aquaculture Calculations for New Reference Point

[REFERENCE RMD FILE: https://raw.githubusercontent.com/OHI-Science/ohiprep_v2019/gh-pages/globalprep/mar/v2019/reference_point/CountryProductionEstimate.Rmd]

# Summary
This analysis produces potential tonnes of aquaculture from growth potential estimates of finfish and bivalves. These aquaculture numbers will be used for the reference point for the mariculture subgoal. 

***

# Data Source 

## Growth Potential (Phi) data 

**Reference**:      
https://knb.ecoinformatics.org/view/doi:10.5063/F1CF9N69
Rebecca Gentry, Halley Froehlich, Dietmar Grimm, Peter Kareiva, Michael Parke, et al. SNAPP - Mapping the Global Potential for Marine Aquaculture. Knowledge Network for Biocomplexity. doi:10.5063/F1CF9N69. 

**Downloaded**: 7/3/2019

**Description**: Growth Potential estimate raster for global cells 

***
  
# Methods
```{r setup ref, message=FALSE, warning=FALSE}
knitr::opts_chunk$set(eval=FALSE)
#######regression inputs from VBGF_Fish.r

# This is a heavily modified script from KNB:
# https://knb.ecoinformatics.org/view/doi:10.5063/F1CF9N69

# From this paper:
# https://www.nature.org/content/dam/tnc/nature/en/documents/Mapping_the_global_potential_for_marine_aquaculture.pdf


### libraries useful for data wrangling
library(dplyr)
library(tidyr)
library(tidyverse)

## libraries useful for spatial data
library(raster)       
library(rgdal)        
library(sf)         
library(fasterize)

## data visualization
library(RColorBrewer)
library(ggplot2)
library(rasterVis)    
library(maps)

## path management
library(here)


## some OHI files
source('http://ohi-science.org/ohiprep_v2019/workflow/R/common.R')
dir_git <- file.path(here(), "globalprep/mar/v2019")
```

## Establish variables and coefficients
```{r}
## Variables needed to get from PHI to tonnes of production
## MRF: assumes 1 farm is 1km2  ##
# (seems easier to just units of km2, rather than farm)

F_estCoef <- c(7.6792, (-5.8198)) #from regression estimated in VBGF_Fish_Final.r
B_estCoef <- c(2.9959,(-1.6659)) #from regression estimate in VBGF_Bivalves.r
density <- 20 #juveniles per m3
cagesize <- 9000 #m3
cagesperfarm <- 24 #located atleast 1 km apart.. cagesperkm2
bivperfarm <- 130000000 #(100 longlines/km2) * 4 km * (100 bivalves seeded/0.0003 km) = 133333333 bivalve/km2
weight35cm <- 554.8  ## in grams see VBGF_Fish_Final. Paper reports 548 grams 
```

## Finfish Production
```{r}
## Global tiff file of PHI (Growth Potential) estimates
FishPhiALLConstraints <- raster(file.path(dir_M, "git-annex/globalprep/mar/v2019/Spatial_Data/NewLayersWOHypoxia/FishPhiALLConstraints95LT2.tiff"))
plot(FishPhiALLConstraints)
FishPhiVector=getValues(FishPhiALLConstraints)


## Convert Phi raster to number of years it takes to grow a 35 cm fish

LogFishYears <- calc(FishPhiALLConstraints, fun=function(x){F_estCoef[1]+F_estCoef[2]*log(x)})
LogFishYears
plot(LogFishYears)

FishYears <- calc(LogFishYears, fun=function(x){exp(x)})

FishYears
plot(FishYears)
writeRaster(FishYears,file.path(dir_M, "git-annex/globalprep/mar/v2019/Spatial_Data/NewLayersWOHypoxia/FishYearsbyCell.tif"), overwrite=TRUE)

FishYears <- raster(file.path(dir_M, "git-annex/globalprep/mar/v2019/Spatial_Data/NewLayersWOHypoxia/FishYearsbyCell.tif"))
FishYearsVector=getValues(FishYears)



#OHI 2018 regions (original analysis used older regions file)

## call spatial file from sourced file
regions_shape()
OHIcountries <- regions %>%
  filter(type_w_ant == "eez")
OHIcountries <- st_transform(OHIcountries, crs(FishPhiALLConstraints))
OHIcountries_raster <- fasterize(OHIcountries, FishPhiALLConstraints, field="rgn_id")  
CountryVector=getValues(OHIcountries_raster)




### area of each cell (each cell is different given lat/long coordinate reference system)
areaPerCell <- area(FishPhiALLConstraints, weights=FALSE, na.rm=TRUE)
areaPerCellVector <- getValues(areaPerCell)

### Make a dataframe with raster values that includes cells: Country, area, Phi, and Years to Harvest
productionDF <- data.frame(CellID = 1:933120000,
                           Country = CountryVector,
                           AreaKM2 = areaPerCellVector,
                           PhiPrime = FishPhiVector, 
                           YearsToHarvest = FishYearsVector)

head(productionDF)

summary(FishYearsVector)
summary(areaPerCellVector) ##they seem to match



## calculate production for each cell
productionDFFishCells <- productionDF %>%
  filter(!is.na(YearsToHarvest)) %>%
  mutate(F_yieldperfarmMT = (weight35cm * density * cagesize * cagesperfarm)/1000000) %>%  # MRF: units yieldperkm2? 554.8 grams * 20 juv/m3 * 9000 m3 * 24 cages/km2 = grams/km2
  mutate(F_yieldpercellperyear = (F_yieldperfarmMT/YearsToHarvest) * AreaKM2) %>%
  arrange(YearsToHarvest) %>%
  mutate(YieldCumSum = cumsum(F_yieldpercellperyear)) %>%
  mutate(AreaCumSum = cumsum(AreaKM2))

write.csv(productionDFFishCells, file.path(dir_M, "git-annex/globalprep/mar/v2019/int/fish/productionDFFishCells.csv")) #save to mazu because so large. This is functionally a raster file. 

productionDFFishCells <- read.csv(file.path(dir_M, "git-annex/globalprep/mar/v2019/int/fish/productionDFFishCells.csv"))

head(productionDFFishCells)
summary(productionDFFishCells)
str(productionDFFishCells)
##cumsum area is 11,402,629 km2 -  # matches the paper 
##cumprod is 15,950,000,000MT -   # matches the paper 

### how many of these cells are not in a country?
sum(is.na(productionDFFishCells$Country))
dim(productionDFFishCells)

## MRF: with new OHI regions: 544,569 are not in a country..probably a lot are in conflicted areas


## Calculate production if 1% of top production area is used:
productionByCountryFishDF <- productionDFFishCells %>%
  filter(!is.na(Country)) %>%
  dplyr::select(-YieldCumSum, -AreaCumSum, -X) %>%
  arrange(YearsToHarvest) %>%
  mutate(ID = Country) %>%
  dplyr::arrange(ID) %>%
  group_by(ID) %>%
  mutate(CountryYieldCumSum = cumsum(F_yieldpercellperyear)) %>%
  mutate(CountryAreaCumSum = cumsum(AreaKM2)) %>%
  mutate(MaxProdPerCountry = max(CountryYieldCumSum)) %>% 
  mutate(OnePercentDevPerCountry = .01*max(CountryAreaCumSum)) #calculating 1 percent of area

write.csv(productionByCountryFishDF,file.path(dir_M, "git-annex/globalprep/mar/v2019/int/fish/FishProdByCountryByCell.csv"), row.names = FALSE) #save to mazu because so large. This is functionally a raster file. 

productionByCountryFishDF <- read.csv(file.path(dir_M, "git-annex/globalprep/mar/v2019/int/fish/FishProdByCountryByCell.csv"))

## MRF: For each area identify amount of area that corresponds to 1% of production area, 
## MRF: assume maximum production within the country for the 1% of area
CountryProdSummary <- productionByCountryFishDF %>%
  dplyr::arrange(YearsToHarvest) %>%
  dplyr::arrange(ID) %>%
  group_by(ID) %>%
  filter(CountryAreaCumSum <= OnePercentDevPerCountry) %>%
  mutate(ProdPerCountryOnePercent = max(CountryYieldCumSum)) %>%
  slice(1)


write.csv(CountryProdSummary, file.path(dir_git, "/int/FishProdByCountrySummary.csv"), row.names = FALSE) #save to github

CountryProdSummary <- read.csv(file.path(dir_git, "int/FishProdByCountrySummary.csv"))

# MRF: get fasted YearsToHarvest for each country
CountryProdSummaryNop <- productionByCountryFishDF %>%
  dplyr::arrange(YearsToHarvest) %>%
  dplyr::arrange(ID) %>%
  group_by(ID) %>%
  slice(1)



# Add country names

region_data()
CountryLabel <- rgns_eez %>%
  dplyr::select(ID = rgn_id, rgn_name)


## Final data
## I think the relevant value we want for the reference point is in this table:  ProdPerCountryOnePercent. Indonesia number for fish matches up with paper. > 24 million tonnes of fish if 1% of aquaculture potential developed. 
CountryProdSummaryFAO <- CountryProdSummary %>%
  ungroup %>%
  dplyr::select(ID:ProdPerCountryOnePercent) %>%
  full_join(CountryLabel, by= "ID")

write.csv(CountryProdSummaryFAO, file.path(dir_git, "/int/FishProdByCountrySummaryFAO.csv"), row.names = FALSE) #save to github

sum(CountryProdSummaryFAO$MaxProdPerCountry, na.rm = TRUE)
#15451023277 number matches paper. > 15 billion tonnes 
```


## Bivalve Production
```{r}
##Now for Bivalves 
BivalvePhiALLConstraints=raster(file.path(dir_M, "git-annex/globalprep/mar/v2019/Spatial_Data/NewLayersWOHypoxia/BivalvePhiALLConstraints95LT1.tif"))

plot(BivalvePhiALLConstraints)
BivalvePhiVector <- getValues(BivalvePhiALLConstraints)


#OHI 2018 regions (original analysis used older regions file)

## call spatial file from sourced file
regions_shape()

OHIcountries <- regions %>%
  filter(type_w_ant == "eez")

OHIcountries <- st_transform(OHIcountries, crs(FishPhiALLConstraints))

OHIcountries_raster <- fasterize(OHIcountries, FishPhiALLConstraints, field="rgn_id")  

CountryVector <- getValues(OHIcountries_raster)

#make the value of each cell the years it takes to grow a 4 cm bivlave
LogBivalveYears <- calc(BivalvePhiALLConstraints, fun = function(x){B_estCoef[1] + B_estCoef[2]*(x)})
LogBivalveYears
plot(LogBivalveYears)

BivalveYears=calc(LogBivalveYears,fun=function(x){exp(x)})

BivalveYears
plot(BivalveYears)

writeRaster(BivalveYears,file.path(dir_M, "git-annex/globalprep/mar/v2019/Spatial_Data/NewLayersWOHypoxia/BivalveYearsByCell.tif"), overwrite=TRUE)

BivalveYears <- raster(file.path(dir_M, "git-annex/globalprep/mar/v2019/Spatial_Data/NewLayersWOHypoxia/BivalveYearsByCell.tif"))

BivalveYearsVector <- getValues(BivalveYears)


###now load in area values for each cell
#areaPerCell=raster("Spatial_Data/MiddleFiles/AreaBivalveLT1.grd")
areaPerCell <- area(BivalvePhiALLConstraints, weights=FALSE, na.rm=TRUE)
areaPerCell

areaPerCellVector=getValues(areaPerCell)


productionDFBiv <- data.frame(CellID=1:933120000,
                           Country=CountryVector,
                           AreaKM2=areaPerCellVector,
                           PhiPrime=BivalvePhiVector, 
                           YearsToHarvest=BivalveYearsVector)

head(productionDFBiv)

productionDFBivCells <- productionDFBiv %>%
  filter(!is.na(YearsToHarvest)) %>%
  mutate(B_yieldperfarmInd = bivperfarm) %>%
  mutate(B_yieldpercellperyear = (B_yieldperfarmInd / YearsToHarvest) * AreaKM2) %>%
  arrange(YearsToHarvest) %>%
  mutate(YieldCumSum = cumsum(B_yieldpercellperyear)) %>%
  mutate(AreaCumSum = cumsum(AreaKM2))

head(productionDFBivCells)
summary(productionDFBivCells)
str(productionDFBivCells)

write.csv(productionDFBivCells, file.path(dir_M, "/git-annex/globalprep/mar/v2019/int/productionDFBivCells.csv")) #save to mazu because so large. This is functionally a raster file. 

productionDFBivCells <- read.csv(file.path(dir_M, "/git-annex/globalprep/mar/v2019/int/bivalve/productionDFBivCells.csv"))

productionByCountryBivDF <- productionDFBivCells %>%
  filter(!is.na(Country)) %>%
  dplyr::select(-YieldCumSum, -X) %>%
  dplyr::select(-AreaCumSum) %>%
  arrange(YearsToHarvest) %>%
  mutate(ID = Country) %>%
  dplyr::arrange(ID) %>%
  group_by(ID) %>%
  mutate(CountryYieldCumSum = cumsum(B_yieldpercellperyear)) %>%
  mutate(CountryAreaCumSum = cumsum(AreaKM2)) %>%
  mutate(MaxPhi = max(PhiPrime)) %>%
  mutate(averagePhi = mean(PhiPrime)) %>%
  mutate(averageWeightedPhi = sum(PhiPrime*AreaKM2)/(max(CountryAreaCumSum))) %>%
  mutate(MaxDevPerCountry = max(CountryAreaCumSum)) %>%
  mutate(MaxProdPerCountry = max(CountryYieldCumSum)) %>%
  mutate(OnePercentDevPerCountry = .01*max(CountryAreaCumSum))

head(productionByCountryBivDF)

write.csv(productionByCountryBivDF,file.path(dir_M, "git-annex/globalprep/mar/v2019/int/bivalve/BivProdByCountryByCell.csv"), row.names = FALSE) #save to mazu because so large. This is functionally a raster file. 

productionByCountryBivDF <- read.csv(file.path(dir_M, "git-annex/globalprep/mar/v2019/int/bivalve/BivProdByCountryByCell.csv")) 


CountryProdSummary <- productionByCountryBivDF %>%
  filter(CountryAreaCumSum <= OnePercentDevPerCountry) %>%
  mutate(ProdPerCountryOnePercent = max(CountryYieldCumSum)) %>%
  slice(1)

write.csv(CountryProdSummary,file.path(dir_git, "int/BivalveProdByCountrySummary.csv"), row.names = FALSE) #save to github


region_data()
CountryLabel <- rgns_eez %>%
  dplyr::select(ID = rgn_id, rgn_name)

head(CountryLabel)

## I think the relevant value we want for the reference point is in this table:  ProdPerCountryOnePercent. Indonesia number for bivalves is higher than the paper, however the paper does use the phrase "over 3.9*10^11 tonnes". Data says 4.7 * 10^11 million tonnes of bivalves if 1% of aquaculture potential developed, which is greater than 3.9 * 10^11 tonnes.  
CountryProdSummaryFAO <- CountryProdSummary %>%
  ungroup %>%
  dplyr::select(ID:ProdPerCountryOnePercent) %>%
  full_join(CountryLabel, by = "ID")

write.csv(CountryProdSummaryFAO, file.path(dir_git, "int/BivalveProdByCountrySummaryFAO.csv"), row.names = FALSE) #save to github 

sum(CountryProdSummaryFAO$MaxDevPerCountry, na.rm = TRUE)
#1491404 km2. Matches paper ~1,500,000 km2
```


## Unit Conversion and Gap Filling
To compare potential vs harvest, we need to convert bivalve units to metric tonnes, they are currently in units of individual bivalves.  Some figures:

* Scallops: [40-50 per kg (~22g ea)](http://www.thefoodcoach.com.au/food/?FoodID=241)
* Mussels: [15-20 per 1 lb blue (~25g ea)](https://www.pacseafood.com/products/details/seafood,shellfish/mussels/)
* Oysters: [12-15 per 1 lb (~33g ea) aquaculture halfshell](http://www.mobjackbayseafood.com/products.html)
* Clams: [20-30 per 1 lb Manila (~18g ea), 8-14 per 1 lb littlenecks (~40g ea)](https://www.pacseafood.com/products/details/seafood,shellfish/clams/)

Averaging these gives about 27.5g per piece.

``` {r}
aq_mass_per_pc <- 0.0275 * 1e-3 ### mass of bivalve piece in tonnes

pot_b <- read_csv(file.path(dir_git, 'int/BivalveProdByCountrySummaryFAO.csv')) %>%
  mutate(potential_prod_one_percent_b = ProdPerCountryOnePercent * aq_mass_per_pc,
         potential_prod_max_b = MaxProdPerCountry * aq_mass_per_pc,
         aq_type = 'shellfish',
         units   = 'tonnes') %>%
  dplyr::select("rgn_id" = "ID", rgn_name, potential_prod_one_percent_b, potential_prod_max_b)

pot_f <- read_csv(file.path(dir_git, 'int/FishProdByCountrySummaryFAO.csv')) %>%
  mutate(aq_type = 'finfish',
         units   = 'tonnes') %>%
  dplyr::select("rgn_id" = "ID", rgn_name, "potential_prod_one_percent_f" = "ProdPerCountryOnePercent", "potential_prod_max_f" = "MaxProdPerCountry")

pot_aq_int <- full_join(pot_b, pot_f) ## makes sense. 1% of potential AREA... not 1% of potential PRODUCTION

write_csv(pot_aq, file.path(dir_git, 'int/aq_potential_int.csv')) ##intermediate file that might come in handy in the future

DT::datatable(pot_aq_int)


pot_aq_final <- pot_aq_int %>%
  mutate(potential_prod_one_percent_b = replace_na(potential_prod_one_percent_b, 0),
         potential_prod_one_percent_f = replace_na(potential_prod_one_percent_f, 0)) %>%
  mutate(potential_mar_tonnes = potential_prod_one_percent_b + potential_prod_one_percent_f) %>%
  arrange(rgn_id) %>%
  dplyr::select(rgn_id, potential_mar_tonnes) %>%
  mutate(year = 2019)

write_csv(pot_aq_final, file.path(dir_git, 'output/aq_potential_final.csv'))

DT::datatable(pot_aq_final)

## make gapfilling dataset 
pot_aq_final_gf <- pot_aq_final %>%
  mutate(gapfilled = case_when(
    potential_mar_tonnes == 0 ~ 1,
    potential_mar_tonnes > 0 ~ 0
  ), 
  method = case_when(
    potential_mar_tonnes == 0 ~ "missing regions given 0 value",
    potential_mar_tonnes > 0 ~ ""
  )) %>%
  dplyr::select(rgn_id, gapfilled, method, year)

write_csv(pot_aq_final_gf, file.path(dir_git, "output/aq_potential_gf.csv"))
```
***

