---
title: "Explore NP Changes"
author: "Gage Clawson"
date: "7/7/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r setup, warning=FALSE, message=FALSE}

knitr::opts_chunk$set(eval=FALSE)

## load libraries, set directories
library(ohicore)  #devtools::install_github('ohi-science/ohicore@dev')
library(dplyr)
library(stringr)
library(tidyr)
library(zoo)  
library(ggplot2)
library(here)
library(tidyverse)
library(plotly)
library(readr)


## Load FAO-specific user-defined functions
source(here('workflow/R/fao_fxn.R')) # function for cleaning FAO files
source(here('workflow/R/common.R')) # directory locations
source(here('globalprep/np/v2020/R/np_fxn.R'))
source(here('globalprep/mar/v2020/mar_fxs.R')) # functions specific to mariculture dealing with compound countries

```


Simultaneously read and process FAO commodities value and quantity data.

```{r}

# ## NOTE: This can be run as a loop, but the "value" and "quant" datasets need to be run individually to make sure
# ##  there are no problems (after this check, they can be looped for efficiency)
# 
# ## describe where the raw data are located:
# dir_fao_data <- file.path(dir_M, 'git-annex/globalprep/_raw_data/FAO_commodities/d2020')
# 
# ## list files included in d2020 folder (value and quant datasets)
# files <- list.files(dir_fao_data, pattern=glob2rx('*.csv'), full.names=TRUE)
# 
# ## To compare to old data:
# # dir_fao_data <- file.path(dir_M, 'git-annex/globalprep/_raw_data/FAO_commodities/d2018')
# # files <- list.files(dir_fao_data, pattern=glob2rx('*.csv'), full.names=T)
# 
# ## loop
# for (f in files){ # f = files[2]
#   cat(sprintf('\n\n\n====\nfile: %s\n', basename(f)))
#   
#   
#   d <- read.csv(f, check.names=FALSE, strip.white=TRUE, stringsAsFactors = FALSE) # stringsAsFactors=T
#   # checks names syntactically, strips leading and trailing whitespace, prevents conversion of characters to factors 
#   
#   ## Specifies that units are tonnes if we are reading in the Commodities Quantity data csv, and usd if we are reading in the Commodities Value data csv
#   units <- c('tonnes','usd')[str_detect(f, c('quant','value'))] # detect unit name using lowercase American English
# 
#   ## gather into long format and clean up FAO-specific data foibles
#   ## warning: attributes are not identical across measure variables; they will be dropped: this is fine
#   m <- d %>% 
#     rename(country   = `Country (Country)`,
#            commodity = `Commodity (Commodity)`,
#            trade     = `Trade flow (Trade flow)`) %>%
#     gather(year, value, -country, -commodity, -trade, -Unit)
#   
#   ## Include only the "Exports" data:
#   m <- m %>%
#     filter(trade == "Exports")
# 
#   m <- m %>%
#     fao_clean_data() %>%  # swaps out FAO-specific codes. NOTE: optional parameter 'sub_0_0' can be passed to control how a '0 0' code is interpreted.
#     select(-trade, -Unit) %>% # eliminate 'trade' column
#   arrange(country, commodity, is.na(value), year)
# 
#   
#   ## Products join: attach product categories from com2prod, and
#   ##   filter out all entries that do not match a product category.
#   ## Note: commodity_lookup is user-defined function to compare 
#   ##   commodities in data vs commodities in lookup table
#   
#   ## load lookup for converting commodities to products
#   com2prod <- read.csv(here('globalprep/np/v2020/raw/commodities2products.csv'), na.strings='')
#   
#   ## version used in 2019:
#   ##    read.csv(here('globalprep/np/v2019/raw/commodities2products.csv'), na.strings='')
#   ## version used in 2018:
#   ##    com2prod <- read.csv('raw/commodities2products.csv', na.strings='')
#   ## version used in 2015: use when testing....
#   ##    com2prod <- read.csv('../v2014_test/commodities2products.csv', na.strings='')
#     
#   ## Check the current commodity-to-product lookup table.  If necessary, make changes to     "raw/commodities2products.csv"
#   np_commodity_lookup(m, com2prod)
#     
#   ## inner_join will attach product names to matching commodities according to
#   ##    lookup table 'com2prod', and eliminate all commodities that do not appear in the lookup table.
#   m <- m %>%
#       inner_join(com2prod, by='commodity')
#     
#     
#   ## Special case: user-defined function deals with 
#   ##   breaking up Antilles into separate reported rgns
#   m <- np_split_antilles(m)
#     
#   ## Some changes to region names that aren't working in name_2_rgn()
#   m <- m %>%
#     mutate(country = ifelse(country == "Côte d'Ivoire", "Ivory Coast", country)) %>%
#     mutate(country = ifelse(country == "C<f4>te d'Ivoire	", "Ivory Coast", country)) %>%
#     mutate(country = ifelse(country == "C\xf4te d'Ivoire", "Ivory Coast", country)) %>%
#     mutate(country = ifelse(country == "Cura<e7>ao","Curacao", country)) %>%
#     mutate(country = ifelse(country == "Curaçao","Curacao", country)) %>%
#     mutate(country = ifelse(country == "Cura\xe7ao","Curacao", country)) %>%
#     mutate(country = ifelse(country == "R\xe9union", "Reunion", country)) %>% 
#     mutate(country = ifelse(country == "Réunion", "Reunion", country)) %>% 
#     mutate(country = ifelse(country == "R<e9>union", "Reunion", country)) %>% 
#     filter(country != "Azerbaijan") # landlocked, but not being removed by name_2_rgn?
#                
#     
#   m_rgn <- name_2_rgn(df_in = m,
#                       fld_name='country', 
#                       flds_unique=c('commodity', 'product', 'year'))
#     
# # v2020 duplicates: China, Hong Kong SAR, Macao SAR, Guadeloupe, Serbia and Montenegro, etc - these are addressed below in the group_by/summarize pipe  
#   
#   ## combine composite regions
#   ## When summarizing the dataset, this function provides a modified way to sum the value column while maintaining NA values when both variables are NA (rather than turning to zero values). The function will sum non-NA values normally.
#   sum_function <- function(x) {
#     if (sum(is.na(x)) == length(x)) 
#       return(NA)
#     return(sum(x, na.rm = T))}
#   
#   m_rgn <- m_rgn %>%
#     group_by(rgn_id, rgn_name, commodity, product, year) %>%
#     summarize(value = sum_function(value)) %>%
#     ungroup()
# 
#   ## units: rename value field to units based on filename
#   names(m_rgn)[names(m_rgn) == 'value'] <- units  
#   
#   ## output to .csv - should create two csvs (tonnes.csv and usd.csv)
#   harvest_out <- sprintf(here('globalprep/np/v2020/int/%s.csv'), units)
#   write.csv(m_rgn, harvest_out, row.names = FALSE, na = '')
# }

```

Read in relevant csvs
```{r}
## Read in quant dataset from intermediate folder
h_tonnes <- read.csv(here('globalprep/np/v2020/int/tonnes.csv'))

## Read in value dataset from intermediate folder
h_usd <- read.csv(here('globalprep/np/v2020/int/usd.csv'))

## concatenates h_tonnes and h_usd data
## h includes rgn_name, rgn_id, commodity, product, year, tonnes, usd.
h <- h_usd %>%
    full_join(h_tonnes, by=c('rgn_name', 'rgn_id', 'commodity', 'product', 'year')) %>%
    mutate(commodity = as.character(commodity)) %>%
    arrange(rgn_id, product, commodity, year)

## clips out years prior to first reporting year, for each commodity per region
h <- h %>% np_harvest_preclip()

```

```{r}
## graph totals of each product

graph_data_1 <- h %>%
  group_by(product) %>%
  summarise(sum_usd = sum(usd, na.rm = TRUE), sum_tonnes = sum(tonnes, na.rm = TRUE)) %>%
  ungroup()
  
ggplot(graph_data_1, aes(x = product, y = sum_usd)) +
  geom_col() + 
  theme_bw()


ggplot(graph_data_1, aes(x = product, y = sum_tonnes)) +
  geom_col() + 
  theme_bw()


graph_data_2 <- h %>%
  group_by(rgn_id, rgn_name, product) %>%
  summarise(sum_usd = sum(usd, na.rm = TRUE), sum_tonnes = sum(tonnes, na.rm = TRUE)) %>%
  arrange(-sum_usd)

test <- graph_data_2 %>%
  group_by(rgn_id, rgn_name) %>%
  summarise(max = max(sum_usd)) %>%
  ungroup() %>%
  left_join(graph_data_2, by = c("rgn_id", "rgn_name")) %>%
  filter(max == sum_usd) %>%
  group_by(product) %>%
  summarise(max_count_usd = n())
  

  ggplot(test, aes(x = product, y = max_count_usd, label = max_count_usd)) +
  geom_col() + 
  theme_bw() +
  geom_label() + 
  labs(title = "Count of regions with maximum usd per product ")
```

# Seaweeds

**Steps**
1. Explore using FAO aquaculture dataset for seaweed instead of FAO commodities dataset
  - compare summed values
2. New reference point: no net loss of seaweed production within a country

## Import Raw Data: FAO Mariculture data

Mariculture production in tonnes.

```{r}

mar <- read.csv(file.path(dir_M, 'git-annex/globalprep/_raw_data/FAO_mariculture/d2020/FAO_GlobalAquacultureProduction_Quantity_1950_2018.csv'), check.names=FALSE, stringsAsFactors=FALSE) ; head(mar) 

```

## Wrangle: 

### Tidy mariculture data

Filter freshwater mariculture, make long format, and clean FAO codes.

```{r}

mar <- mar %>%
  rename(country = `Country (Country)`,
         FAO_name = `ASFIS species (ASFIS species)`, 
         fao = `FAO major fishing area (FAO major fishing area)`, 
         environment = `Environment (Environment)`)
table(mar$environment)  

## include only marine environments
mar <- mar %>%
filter(environment %in% c("Brackishwater", "Marine"))  

## convert to long format and clean FAO codes:
mar <- mar %>%
  select(-Unit) 

mar <- mar %>%
  gather(key="year", value="value", num_range("",1950:2018)) %>%
    fao_clean_data() 

```

### Update species names

Update species name in the `raw/species_list.csv` file with names in the `mar` dataset (exclude non-food species). I simplified the species_list.  I cut the "species" name columns because it wasn't clear what this was trying to accomplish and created potential error.  

```{r}

## Read in 'species_list.csv' (originally from 'globalprep/mar/v2020/raw'). Filter out for 'Taxon_code = AL' (only seaweed species). Rename 'exclude' columns to 'include' since we're including previously excluded seaweed species (not primarily used as human food). Therefore, "0" means exclude, and "1" means include. Four scenarios for comparison purposes: 1) "include" - best guess of non-human food species; 2) "include_no_food" - if species is used for human food at all, exclude it; 3) "include_all_seaweed" - including all seaweeds in NP goal (including seaweeds used primarily for human food), 4) "include_all_nei" - including all seaweeds not elsewhere included in the dataset.
seaweed_sp <- read.csv('raw/species_list.csv', stringsAsFactors=FALSE) %>% 
  filter(Taxon_code == 'AL') %>% 
  rename(include = exclude, 
         include_all_seaweed = exclude_no_seaweed, 
         include_all_nei = exclude_no_nei) %>% 
  mutate(include_no_food = include) %>% 
  dplyr::select(1:2, include_no_food, everything())
  

## Save in 'globalprep/np/v2020/explore/raw' as 'species_list_np_seaweeds.csv'. Export and edit csv file directly to go through each species and confirm if appropriate to include in NP goal, adding to 'notes_2' column why each species is included/excluded and what proportion (with a reference).
write.csv(seaweed_sp,"raw/species_list_np_seaweeds.csv", row.names = FALSE)

## Read in edited 'species_list_np_seaweeds_edited.csv'.
seaweeds <- read.csv('raw/species_list_np_seaweeds_edited.csv', stringsAsFactors = FALSE)

```

Changes made to species_list_np_seaweeds.csv:

FAO_name                   |  change in "include" value  | explanation
-------------------------- | --------------------------- | -------------
[Chondracanthus chamissoi] | 0 --> 0.8                   | Used in carrageenan
[Porphyra columbina]       | 0 --> 0.01                  | Porphyra spp. are produced almost exclusively for direct human consumption
[Spirulina maxima]         | 0 --> 0.3                   | More than 70% of Spirulina market is for human consumption, mainly as health food
Aquatic plants nei         | no change (0)               | The important seaweeds for human consumption are already listed
Babberlocks                | no change (0)               | Only used for human food according to FAO
Bright green nori          | 0 --> 0.01                  | Used for abalone feed
Dark green nori            | 0 --> 0.01                  | Used for abalone feed
Sea lettuces nei           | 0 --> 0.5                   | Used for feed in aquarium trade
Brown seaweeds             | 0 --> 0.5                   | Main uses: food and alginate
Caulerpa seaweeds          | 0 --> 0.01                  | Used for abalone feed
[Meristotheca senegalense] | no change (0)               | Only used for human food
Dulse                      | 0 --> 0.1                   | Found to be a valuable food for red abalone
Elkhorn sea moss           | 1 --> 0.99                  | Primarily used to produce carrageenan, but also edible
Eucheuma seaweeds nei      | 1 --> 0.9                   | Also eaten raw in some countries
Fragile codium             | 1 --> 0.1                   | Medicinal uses
Fusiform sargassum         | no change (0)               | Only used for human food
Gelidium seaweeds          | 1 --> 0.8                   | Primarily used for agar, but also consumed as food
Giant kelp                 | 1 --> 0.9                   | Primarily used for alginate, but also consumed as food
Gracilaria seaweeds        | 1 --> 0.75                  | Primarily used for agar, but also consumed as food (Ogo)
Green laver                | 0 --> 0.01                  | Used for abalone feed
Harpoon seaweeds           | 1 --> 0.75                  | Also used for human food (ingredient in poke)
Japanese isinglass         | 1 --> 0.95                  | Primarily used for agar, but also consumed as food, especially in Korea
Japanese kelp              | 0 --> 0.1                   | Primarily used for human food (kombu), but also for alginate
Kelp nei                   | 1 --> 0.5                   | Main uses: food and alginate
Laver (Nori)               | 0 --> 0.01                  | Used for abalone feed
Mozuku                     | 0 --> 0.01                  | Used for homeopathic purposes
Nori nei                   | 0 --> 0.01                  | Used for abalone feed
Red seaweeds               | 1 --> 0.75                  | Main uses: food and 2 sources of hydrocolloids (agar and carrageenan)
Sea belt                   | 1 --> 0.95                  | Primarily used for agar, but food market growing
Seaweeds nei               | 0 --> 0.15                  | Most seaweeds are used for human food
Spiny eucheuma             | 1 --> 0.9                   | Primarily used for carrageenan, but also consumed as food
Tangle                     | 1 --> 0.99                  | Primarily used for alginate, but also consumed as food
Wakame                     | 0 --> 0.01                  | Used for abalone feed
Wakame nei                 | 0 --> 0.01                  | Used for abalone feed
Warty gracilaria           | 1 --> 0.95                  | Primarily used for agar, but increasingly used for abalone feed
Giant kelps nei            | 1 --> 0.9                   | Primarily used for alginate, but also consumed as food
Green seaweeds             | 1 --> 0.75                  | Mixed non-food uses, but also consumed as food
Coarse seagrape            | no change (0)               | Only used for human food
[Sargassum spp]            | 1 --> 0.99                  | Primarily used for alginate, but also consumed as food
Spirulina nei.             | 1 --> 0.3                   | More than 70% of Spirulina market is for human consumption, mainly as health food
[Dunaliella salina]        | no change (1)               | Used in cosmetics and dietary supplements
[Capsosiphon fulvescens]   | 1 --> 0.1                   | Used as traditional medicine
Slender wart weed          | 1 --> 0.75                  | Primarily used for agar and fish feed, but also consumed as food


```{r}

## Commented out 'read.csv' lines are different versions of the dataset, representing the three different seaweed exclusion methods (exclude some seaweed, all seaweed, all seaweed nei). Use at the end for data checking.The first one is the one which is saved to the "output" folder. 
mar_sp <- read.csv('raw/species_list.csv', stringsAsFactors=FALSE)%>%
  select(FAO_name, exclude, alias, Taxon_code)
#mar_sp <- read.csv('raw/species_list.csv', stringsAsFactors=FALSE) %>%
#  select(FAO_name, exclude_no_seaweed, alias, Taxon_code)
#mar_sp <- read.csv('raw/species_list.csv', stringsAsFactors=FALSE) %>%
#  select(FAO_name, exclude_no_nei, alias, Taxon_code)

new.spp <- setdiff(mar$FAO_name, mar_sp$FAO_name)
new.spp # check: if dim has 0 rows it means all match
## if there is a list of species, hand check species_list.csv to see whether to keep (exclude seaweeds and species harvested for ornamental/medicinal), check if synonyms match Trujillo names. 
## This is outlined in issue #8 and issue #81 for reference. You need to add the species output from this to species_list.csv and fill out the corresponding information. I.e. add FAO_name, exlude (whether to include in assessment or not. If they are only harvested for medicinal purposes, do not include), alias, and Taxon_code to species_list.csv.

## REMOVE SPECIES not relevant to mariculuture goal (i.e., non-food species)
mar <- mar %>% left_join(mar_sp, by="FAO_name") 

## Filters out species that should be excluded from the MAR sub-goal, depending on which of the 3 scenarios was read in. Searches for column name "exclude", "exclude_no_seaweed", or "exclude_no_nei"
if(sum(str_detect(names(mar_sp), "exclude$"))==1) {
  
 mar <- mar %>% filter(exclude==0)

 } else if (sum(str_detect(names(mar_sp), "exclude_no_seaweed"))==1) {
  
   mar <- mar %>% filter(exclude_no_seaweed==0)

   } else if (sum(str_detect(names(mar_sp), "exclude_no_nei"))==1) {
  
     mar <- mar %>% filter(exclude_no_nei==0)

     }
 
## change names using species alias or FAO species name (global changes)
mar$species <- ifelse(!is.na(mar$alias), mar$alias, mar$FAO_name) 

## sum production values for each group to account for duplicate rows after name change (remove NA values)
mar <- mar %>%
  filter(!is.na(value)) %>%
  group_by(country, fao, environment, species, year, Taxon_code) %>% 
    summarize(value = sum(value)) %>% 
  ungroup()


## eliminate country-species data with zero production throughout the time-series (1950-recent)
mar <- mar %>%
  group_by(country, species) %>%
  mutate(total_value = sum(value)) %>%
  filter(total_value > 0) %>%
  select(-total_value) %>%
  ungroup()

```

```{r}

seaweed_culture

seaweed_commod

```

## Exploring new seaweed reference point

**Steps**
1. Per country: divide tonnes production / area of suitability
2. Use the value corresponding to the 95th quantile (across countries and a range of years) of these country data to rescale the data.

```{r}

# ## Filter out seaweeds from tonnes production data
# seaweeds <- read.csv(here('globalprep/np/v2020/output/np_harvest_tonnes.csv')) %>% # using gapfilled layer from np_dataprep.Rmd
#   filter(product == "seaweeds") %>% 
#   left_join(rgns_eez, by = "rgn_id") %>% 
#   dplyr::select(rgn_id, rgn_name, year, tonnes)
# 
# ## Read in list of seaweed suitable countries from Froehlich (2018)
# urlfile = "https://raw.githubusercontent.com/CART-sci/seaweed/master/Data/sw_feasible_countries.csv"
# sw_feas_countries <- read_csv(url(urlfile)) %>% 
#   rename(rgn_name = country)
# 
# ## Divide tonnes production / area of suitability
# seaweed_density <- seaweeds %>% 
#   left_join(sw_feas_countries, by = "rgn_name") %>% 
#   mutate(density = tonnes/feas_area) %>% # tonnes/area(km2)
#   dplyr::select(-X1, -ID)
# 
# summary(seaweed_density)
# 
# ## Find 95th quantile density for each region, year
# quant_95 <- seaweed_density %>% 
#   group_by(rgn_name, year) %>% 
#   summarize(quantDensity = quantile(density, c(0.95), na.rm=TRUE)) %>% 
#   data.frame()
# 
# ## Rescale the density using 95th quantile value across countries and years
# seaweed_rescale <- seaweed_density %>% 
#   mutate(density_rescaled = density/quant_95$quantDensity) %>% 
#   mutate(density_rescaled_capped = ifelse(density_rescaled > 1, 1, density_rescaled))
# 
# hist(seaweed_rescale$density_rescaled_capped)
# 
# # Scrapping this

```

