---
title: 'OHI 2020: Natural Products: Ornamentals data prep'
author: "*Compiled on `r date()` by `r Sys.info()['user']`*"
output: 
  html_document:
    toc: true
    number_sections: true
    theme: cerulean
    highlight: haddock
    includes: 
      in_header: '../../../workflow/templates/ohi_hdr.html'
  pdf_document:
    toc: true
---

[REFERENCE RMD FILE: http://ohi-science.org/ohiprep_v2020/globalprep/np/v2020/np_dataprep.html]

# Summary
This analysis converts FAO mariculture data into one of the data layers used to calculate OHI 2020 global natural products (NP) scores. We will conduct the overall NP data prep on seaweeds, fish oil/fish meal (FOFM), and ornamentals, so that we can produce $ values for weighting later on, however, our final layer from this data prep will only consist of seaweeds.

# Updates from previous assessment
Instead of using all NP commodity categories (seaweed, FOFM, coral, sponge, and ornamentals), we are only prepping seaweeds in this data prep. Overall, we are excluding corals and sponges from our calculations for the NP goal. FOFM is calcualted from Watson fisheries data, seaweed production are calculated from FAO mariculture datasets, while ornamentals production is calculated from FAO commodities datasets. New year of FAO mariculture data (1950-2018). 

***

# Data Source 

## Production data

**Reference**:      
http://www.fao.org/fishery/statistics/software/fishstatj/en#downlApp
Release date: March 2019
FAO Global Aquaculture Production Quantity 1950_2018
FAO metadata found [here](http://www.fao.org/fishery/statistics/global-aquaculture-production/en)

**Downloaded**: 5/11/2020

**Description**:  Quantity (tonnes) of mariculture for each country, species, year.

**Time range**: 1950-2018 

## Sustainability data

**Reference**:      
Mariculture Sustainability Index (Trujillo 2008)

Trujillo, Pablo. 2008. “Using a Mariculture Sustainability Index to Rank Countries’ Performance.” In Fisheries Centre Research Reports, edited by Jackie Alder and Daniel Pauly. University of British Columbia, Vancouver, Canada: Fisheries Centre Research Reports. https://circle.ubc.ca/handle/2429/40933. 

Original data table is located in the Trujillo 2008 paper saved in the OHI Zotero database. Report starts on page 28 of the link provided in the reference. 

**Description**:  

Original MSI rescaled from 1-10 to 0-1 for mariculture taxa groups.

***
  
# Methods
```{r setup, warning=FALSE, message=FALSE}

knitr::opts_chunk$set(eval=FALSE)

## load libraries, set directories
library(ohicore)  #devtools::install_github('ohi-science/ohicore@dev')
library(dplyr)
library(stringr)
library(tidyr)
library(zoo)  
library(ggplot2)
library(here)
library(tidyverse)
library(plotly)
library(readr)

## Load FAO-specific user-defined functions
source(here('workflow/R/fao_fxn.R')) # function for cleaning FAO files
source(here('workflow/R/common.R')) # directory locations
source(here('globalprep/np/v2020/R/np_fxn.R'))
source(here('globalprep/mar/v2020/mar_fxs.R')) # functions specific to mariculture dealing with compound countries

```

# Import Raw Data: FAO Mariculture data

Mariculture production in tonnes.

```{r}

mar <- read.csv(file.path(dir_M, 'git-annex/globalprep/_raw_data/FAO_mariculture/d2020/FAO_GlobalAquacultureProduction_Quantity_1950_2018.csv'), check.names=FALSE, stringsAsFactors=FALSE) ; head(mar) 

```

# Wrangle: 

## Tidy mariculture data

Filter freshwater mariculture, make long format, and clean FAO codes.

```{r}

mar <- mar %>%
  rename(country = `Country (Country)`,
         FAO_name = `ASFIS species (ASFIS species)`, 
         fao = `FAO major fishing area (FAO major fishing area)`, 
         environment = `Environment (Environment)`)
table(mar$environment)  

## Include only marine environments
mar <- mar %>%
filter(environment %in% c("Brackishwater", "Marine"))  

## Convert to long format and clean FAO codes:
mar <- mar %>%
  select(-Unit) 

mar <- mar %>%
  gather(key="year", value="value", num_range("",1950:2018)) %>%
    fao_clean_data() 

```

## Update seaweed species information

Filter out seaweed species from 'raw/species_list.csv' (originally from 'globalprep/mar/v2020/raw'), rename columns, and assign proportions to "include" column determined by research of non-human food vs human food seaweed species cultivated in mariculture. For NP, we are only including non-human food seaweed species. Since some species are used for both non-human food and human food purposes, a proportion is assigned based on research and best guess.

```{r}

## Read in 'species_list.csv' (originally from 'globalprep/mar/v2020/raw'). Filter for 'Taxon_code = AL' (only seaweed species). Rename 'exclude' columns to 'include' since we're now including seaweed species that were excluded in the MAR dataprep (not primarily used as human food). Therefore, "0" means exclude completely (0%), and "1" means include completely (100%). Four scenarios for comparison purposes: 1) "include" - best guess of non-human food species; 2) "include_no_food" - if species is used for human food at all, exclude it; 3) "include_all_seaweed" - including all seaweeds in NP goal (including seaweeds used primarily for human food), 4) "include_all_nei" - including all seaweeds not elsewhere included in the dataset.
seaweed_sp <- read.csv('raw/species_list.csv', stringsAsFactors=FALSE) %>% 
  filter(Taxon_code == 'AL') %>% 
  rename(include = exclude, 
         include_all_seaweed = exclude_no_seaweed, 
         include_all_nei = exclude_no_nei) %>% 
  mutate(include_no_food = include) %>% 
  dplyr::select(1:2, include_no_food, everything())
  
## Save in 'globalprep/np/v2020/explore/raw' as 'species_list_np_seaweeds.csv'. Export and edit csv file directly to go through each species and confirm if appropriate to include in NP goal, adding to 'notes_2' column why each species is included/excluded and what proportion (with a reference).
write.csv(seaweed_sp,"raw/species_list_np_seaweeds.csv", row.names = FALSE)

```

Changes made to species_list_np_seaweeds.csv:

FAO_name                   | Change in "include" value | Explanation summary
-------------------------- | ------------------------- | -------------------
[Chondracanthus chamissoi] | 0 --> 0.8                 | Used in carrageenan
[Porphyra columbina]       | 0 --> 0.01                | Porphyra spp. are produced almost exclusively for direct human consumption
[Spirulina maxima]         | 0 --> 0.3                 | More than 70% of Spirulina market is for human consumption, mainly as health food
Aquatic plants nei         | no change (0)             | The important seaweeds for human consumption are already listed
Babberlocks                | no change (0)             | Only used for human food according to FAO
Bright green nori          | 0 --> 0.01                | Used for abalone feed
Dark green nori            | 0 --> 0.01                | Used for abalone feed
Sea lettuces nei           | 0 --> 0.5                 | Used for feed in aquarium trade
Brown seaweeds             | 0 --> 0.5                 | Main uses: food and alginate
Caulerpa seaweeds          | 0 --> 0.01                | Used for abalone feed
[Meristotheca senegalense] | no change (0)             | Only used for human food
Dulse                      | 0 --> 0.1                 | Found to be a valuable food for red abalone
Elkhorn sea moss           | 1 --> 0.99                | Primarily used to produce carrageenan, but also edible
Eucheuma seaweeds nei      | 1 --> 0.9                 | Also eaten raw in some countries
Fragile codium             | 1 --> 0.1                 | Medicinal uses
Fusiform sargassum         | no change (0)             | Only used for human food
Gelidium seaweeds          | 1 --> 0.8                 | Primarily used for agar, but also consumed as food
Giant kelp                 | 1 --> 0.9                 | Primarily used for alginate, but also consumed as food
Gracilaria seaweeds        | 1 --> 0.75                | Primarily used for agar, but also consumed as food (Ogo)
Green laver                | 0 --> 0.01                | Used for abalone feed
Harpoon seaweeds           | 1 --> 0.75                | Also used for human food (ingredient in poke)
Japanese isinglass         | 1 --> 0.95                | Primarily used for agar, but also consumed as food, especially in Korea
Japanese kelp              | 0 --> 0.1                 | Primarily used for human food (kombu), but also for alginate
Kelp nei                   | 1 --> 0.5                 | Main uses: food and alginate
Laver (Nori)               | 0 --> 0.01                | Used for abalone feed
Mozuku                     | 0 --> 0.01                | Used for homeopathic purposes
Nori nei                   | 0 --> 0.01                | Used for abalone feed
Red seaweeds               | 1 --> 0.75                | Main uses: food and 2 sources of hydrocolloids (agar and carrageenan)
Sea belt                   | 1 --> 0.95                | Primarily used for agar, but food market growing
Seaweeds nei               | 0 --> 0.15                | Most seaweeds are used for human food
Spiny eucheuma             | 1 --> 0.9                 | Primarily used for carrageenan, but also consumed as food
Tangle                     | 1 --> 0.99                | Primarily used for alginate, but also consumed as food
Wakame                     | 0 --> 0.01                | Used for abalone feed
Wakame nei                 | 0 --> 0.01                | Used for abalone feed
Warty gracilaria           | 1 --> 0.95                | Primarily used for agar, but increasingly used for abalone feed
Giant kelps nei            | 1 --> 0.9                 | Primarily used for alginate, but also consumed as food
Green seaweeds             | 1 --> 0.75                | Mixed non-food uses, but also consumed as food
Coarse seagrape            | no change (0)             | Only used for human food
[Sargassum spp]            | 1 --> 0.99                | Primarily used for alginate, but also consumed as food
Spirulina nei.             | 1 --> 0.3                 | More than 70% of Spirulina market is for human consumption, mainly as health food
[Dunaliella salina]        | no change (1)             | Used in cosmetics and dietary supplements
[Capsosiphon fulvescens]   | 1 --> 0.1                 | Used as traditional medicine
Slender wart weed          | 1 --> 0.75                | Primarily used for agar and fish feed, but also consumed as food

More detailed explanations are included in the "notes_2" column of species_list_np_seaweeds_edited.csv. 


## Update species names

Update species name in the `raw/species_list_np_seaweeds_edited.csv` file with names in the `mar` dataset. Simplified the species list and cut the "species" name columns because it wasn't clear what this was trying to accomplish and created potential error.  

```{r}

## Read in edited 'species_list_np_seaweeds_edited.csv'.
seaweeds <- read.csv('raw/species_list_np_seaweeds_edited.csv', stringsAsFactors = FALSE)

## Different versions of the dataset, representing the four different seaweed inclusion methods (include proportions of non-human food species, include only completely non-human food species, include all seaweeds, include all seaweeds nei). Use at the end for data checking. The first one is the one which is saved to the "output" folder. 
seaweeds_sp <- seaweeds %>%
  select(FAO_name, include, alias, Taxon_code)
# seaweeds_sp <- seaweeds %>%
#   select(FAO_name, include_no_food, alias, Taxon_code)
# seaweeds_sp <- seaweeds %>%
#   select(FAO_name, include_all_seaweed, alias, Taxon_code)
# seaweeds_sp <- seaweeds %>%
#   select(FAO_name, include_all_nei, alias, Taxon_code)

new.spp <- setdiff(mar$FAO_name, seaweeds_sp$FAO_name)
new.spp # check: if dim has 0 rows it means all match
# if there is a list of species, hand check species_list.csv to see whether to keep (include seaweeds and species harvested for ornamental/medicinal), check if synonyms match Trujillo names. 
# this is outlined in issue #8 and issue #81 for reference. You need to add the species output from this to species_list_np_seaweeds_edited.csv and fill out the corresponding information. I.e. add FAO_name, include (whether to include in assessment or not. If they are only harvested for human food purposes, do not include), alias, and Taxon_code to species_list_np_seaweeds_edited.csv.
# no seaweed species were listed - v2020

## REMOVE SPECIES not relevant to natural products goal (i.e., human food species)
seaweed_np <- seaweeds_sp %>% 
  left_join(mar, by="FAO_name") %>% 
  filter(include > 0)
# went from 43 species to 39 species (4 completely human food species removed) - v2020
 
## Change names using species alias or FAO species name (global changes)
seaweed_np$species <- ifelse(!is.na(seaweed_np$alias), seaweed_np$alias, seaweed_np$FAO_name) 

## Sum production values for each group to account for duplicate rows after name change (remove NA values)
seaweed_np <- seaweed_np %>%
  filter(!is.na(value)) %>%
  group_by(country, fao, environment, species, year, Taxon_code, include) %>% 
    summarize(value = sum(value)) %>% 
  ungroup()
# went from 39 species to 35 species (lost "Aquatic plants nei", "Bright green nori", "Kelp nei", "Giant kelps nei") due to no production values for those species - v2020
# went from 50 to 49 countries

## Eliminate country-species data with zero production throughout the time-series (1950-recent)
seaweed_np <- seaweed_np %>%
  group_by(country, species) %>%
  mutate(total_value = sum(value)) %>%
  filter(total_value > 0) %>%
  select(-total_value) %>%
  ungroup()

```

## Convert country names to OHI regions

```{r}

## Divide mariculture from countries that we report as separate regions (assume equal production in all regions)
# Netherlands Antilles: Conch restoration among Aruba, Bonaire, Curacao
# Channel Islands: Jersey and Guernsey
# Bonaire/S.Eustatius/Saba
# Yugoslavia SFR: no longer a country after 1992
seaweed_np <- seaweed_np %>%
  mutate(country = ifelse(country=="Réunion", "Reunion", country)) %>%  # this one is hard to get right; v2020: last year it was "R\xe9union", but this year it was "Réunion"
  mar_split()  # function in mar_fxs.R

mar_rgn <- name_2_rgn(df_in = seaweed_np, 
                       fld_name='country', 
                       flds_unique=c('species', 'fao', 'environment', 'Taxon_code', 'year', 'include')) 


## Sum values of regions with multiple subregions
mar_rgn <- mar_rgn %>%
  group_by(fao, environment, species, year, Taxon_code, rgn_id, include) %>% 
  summarize(value = sum(value)) %>%
  ungroup()
# went from 2979 to 2950 observations - v2020

```

**Take a look at the tidied data for a single year and region**

```{r}

data.frame(filter(mar_rgn, rgn_id==182) %>%
  filter(year==2016) %>%
  arrange(species))

```

# Gapfilling

## Fill in missing years after first year of harvest data with 0 values

Checked to make sure that there weren't instances in which it made more sense to carry the previous year's data forward as a method of gapfilling. This didn't seem to be the case.

```{r}

## Spread mar_rgn to create a value for every year-species-region (if missing it will be given a NA)
mar_rgn_spread <- spread(mar_rgn, year, value)
dim(mar_rgn_spread)

## Turn data frame back into long format
mar_rgn_gf <- gather(mar_rgn_spread, "year", "value", num_range("",1950:2018)) %>%
  arrange(rgn_id, species, year, Taxon_code, fao, environment) 

## NA values are converted to zero
mar_rgn_gf <- mar_rgn_gf %>%
  mutate(year = as.numeric(as.character(year))) %>%
  mutate(value_w_0 = ifelse(is.na(value), 0, value)) %>%
  group_by(fao, environment, species, Taxon_code, rgn_id) %>% 
  mutate(cum_value = cumsum(value_w_0)) %>%
  ungroup() %>%
  filter(cum_value > 0) %>% # eliminates years before mariculture began
  mutate(gap_0_fill = ifelse(is.na(value), "NA_to_zero", "0")) %>% # record gapfill
  mutate(value = ifelse(is.na(value), 0, value)) %>% # finally, convert all NAs in original column to 0
  select(-cum_value, -value_w_0)

```

See how may NA values were converted to 0

```{r}

table(mar_rgn_gf$gap_0_fill)
## 382 of these out of 2299+382 cases had NA converted to 0 - v2020

```

Remove species-region-environment time series with less than four years of seaweeed mariculture production > 0 tonnes (assume these are not established seaweed mariculture programs).

```{r}

mar_rgn_gf = mar_rgn_gf %>% 
  group_by(rgn_id, species, fao, environment) %>%
  mutate (not_0 = length(value[value>0])) %>% # length of vector of years greater than 0
  filter (not_0>3) %>% # filter for groups that have at least four years of seaweed mariculture production 
  ungroup() %>% 
  select(rgn_id, species, fao, environment, year, include, value, Taxon_code, gap_0_fill) 

```

Add a unique identifier per cultivated stock that describes each species, fao region, and environment grouping.

```{r}

## Add a unique identifier per cultivated stock
identifier = mar_rgn_gf %>% 
  select(rgn_id, species, fao, environment) %>% 
  unique() %>% 
  mutate(species_code = 1:n())
# 82 unique identifiers - v2020

mar_rgn_gf = left_join(mar_rgn_gf, identifier)
maric <- mar_rgn_gf

```

# Calculate and save tonnes of seaweed

Find the tonnes per each region/year per each seaweed type (multiplied by "include" proportions).

Used to estimate total seaweed mariculture yield per country.

```{r}

## Multiply "include" column by "value" column to find tonnes per region/year for each seaweed species
maric <- maric %>% 
  mutate(tonnes = include*value)

## Save in 'globalprep/np/v2020/explore/int' as 'np_seaweeds_tonnes.csv' for weighting purposes later on
write.csv(maric,"int/np_seaweeds_tonnes_weighting.csv", row.names = FALSE)

```

# Import data: Trujillo sustainability scores

These data describe the sustainability and genetic escapes for country/species combinations (and, in a couple cases, environment and FAO region combinations). In cases where these data were not available for a specific county/species, we averaged the data across taxonomic groups to gapfill the missing data.

```{r}

## Trujillo sustainability data copied over from 'globalprep/mar/v2020/raw':
sus = read.csv('raw/Truj_label_sust.csv', stringsAsFactors = FALSE, na.strings = NA)

```

# Wrangle

```{r}

## These need to be re-added (get cut when associated with region ids)
sus_no_rgn <- filter(sus, is.na(country))
# seaweed species included in list (species_FAO/species_Truj): Gracilaria seaweeds/Gracilaria seaweeds, Laver (Nori)/Laver (Nori), Nori nei/Laver (Nori), Brown seaweeds/Brown seaweeds, Edible seaweeds/Edible seaweeds - v2020

```

## Convert country names to OHI region names.

```{r}

sus_rgn <- name_2_rgn(df_in = sus, 
                       fld_name='country', 
                       flds_unique=c('species_fao', 'fao', 'environment', 'species_Truj'))

sus_rgn <- bind_rows(sus_rgn, sus_no_rgn) %>%
  unique()

```

**Check that the non-matches between Trujillo sustainability FAO spp (sus_rgn) and the FAO mariculture species in the wrangled FAO Aquaculture Production data table (maric) are not due to spelling errors or slightly different names. We want to include as many species that have sustainability scores as possible**

```{r}

## Make sure same species are spelled the same in the two data tables (e.g. check that there are no extra spaces)
sort(setdiff(sus_rgn$species_fao, maric$species)) # species that no longer have mariculture industry or are not included due to being freshwater, etc.
# Brown seaweeds (Pac) - this is fine, since 'sus_rgn' also has an identical listing for "Brown seaweeds"
# Edible seaweeds - missing completely from 'maric', but that's fine since edible seaweeds are excluded from NP

sort(setdiff(maric$species, sus_rgn$species_fao)) # FAO species with no Trujillo data - there will probably be a long list
# [1] "[Capsosiphon fulvescens]" "Caulerpa seaweeds"        "Dark green nori"          "Elkhorn sea moss"         "Eucheuma seaweeds nei"   
# [6] "Fragile codium"           "Gelidium seaweeds"        "Giant kelp"               "Green laver"              "Harpoon seaweeds"        
# [11] "Japanese isinglass"       "Japanese kelp"            "Mozuku"                   "Red seaweeds"             "Sea belt"                
# [16] "Sea lettuces nei"         "Seaweeds nei"             "Spiny eucheuma"           "Wakame"                   "Wakame nei"              
# [21] "Warty gracilaria"        

## Hand check each of the species output here. I mainly check to make sure there aren't obvious cases when the species is the same, but the names are presented slightly differently in the lists. For example, some species might have an extra space in the name somewhere. If any of these obvious differences exist, we need to go in and change the names to match. 
## Its not too surprising (unfortunately) that the FAO list has far more species than the sustainability list. But, I like to go over them to make sure that we are getting sustainability scores for as many species as we can.

```

# FAO mariculture and sustainability scores

Append sustainability score to the FAO mariculture data.

The following joins the sustainability scores to regions/species that have Trujillo data.
```{r}

table(sus_rgn$match_type)

## Join taxa specific to country/species/environment
c_sp_env = sus_rgn %>% 
  filter(match_type == 'c_sp_env') %>% 
  select(rgn_id, species=species_fao, environment, Sust_c_sp_env = Maric_sustainability)

mar_sus <- maric %>%
  left_join(c_sp_env, by= c("species", "environment", "rgn_id"))

## Join taxa specific to country/species/fao region
c_sp_fao = sus_rgn %>% 
  filter(match_type == 'c_sp_fao') %>% 
  select(rgn_id, species=species_fao, fao, Sust_c_sp_fao = Maric_sustainability)

mar_sus <- mar_sus %>%
  left_join(c_sp_fao, by= c("species", "fao", "rgn_id"))

```

**Take a look at the data thus far**
```{r}

head(data.frame(filter(mar_sus, rgn_id==14 & species == "Laver (Nori)")))
head(data.frame(filter(mar_sus, !is.na(Sust_c_sp_fao))))

```

```{r}

## Join taxa specific to country/species
c_sp = sus_rgn %>% 
  filter(match_type == 'c_sp') %>% 
  select(rgn_id, species=species_fao, Sust_c_sp = Maric_sustainability)

mar_sus <- mar_sus %>%
  left_join(c_sp, by= c("species", "rgn_id"))

```

**Now look at a summary after appending all the Trujillo data**

```{r}

summary(mar_sus)
# Sust_c_sp: 1902 NAs - v2020

```

Merge the three Trujillo type categories into a single sustainability score column in the following order:

1. Sust_c_sp_env: taxa specific to country/species/environment (smallest taxonomic level)
2. Sust_c_sp_fao:  taxa specific to country/species/fao region
3. Sust_c_sp: taxa specific to country/species (highest taxonomic level)

For example, if Sust_c_sp_env is missing, use Sust_c_sp_fao and so on. 

```{r}

mar_sus = mar_sus %>% 
  mutate(Sust = ifelse(!is.na(Sust_c_sp_env), Sust_c_sp_env, Sust_c_sp_fao)) %>%
  mutate(Sust = ifelse(is.na(Sust), Sust_c_sp, Sust)) %>%
  select(-Sust_c_sp_env, -Sust_c_sp_fao, -Sust_c_sp)

summary(mar_sus)
# still 1902 NAs in 'Sust' colum - v2020

```


This joins the sustainability data that is gapfilled either at the species level (average of specific species/genera across regions) or at a higher course taxonomic levels and documents which data are gapfilled and how.

```{r}

## Select observations gapfilled at the species/genera level:
gf_sp_sus <- sus_rgn %>% 
  filter(gapfill != "actuals" & match_type == "species") %>%
  select(species = species_fao, gapfill, Sust_gf_sp = Maric_sustainability)
# Laver (Nori), Nori nei, Brown seaweeds

## check that there are no duplicated species_fao
gf_sp_sus[duplicated(gf_sp_sus$species), ]

## Match gapfilling values by species
mar_sus_gf = mar_sus %>%
  left_join(gf_sp_sus, by = 'species')
# only seaweed species that have values in Sust and Sust_gf_sp: Laver (Nori), Gracilaria seaweeds, Brown seaweeds - v2020
# 1717 NAs in gf_sp_sus - v2020

## Select observations gapfilled at the coarse taxon level:
gf_taxon_sus <- sus_rgn %>% 
  filter(gapfill != "actuals" & match_type == "taxon") %>%
  select(Taxon_code=taxon, Sust_gf_taxon = Maric_sustainability)
# 'AL' Taxon_code (algae) = 0.83 Sust_gf_taxon score

## Match gapfilling values by species
mar_sus_gf = mar_sus_gf %>%
  left_join(gf_taxon_sus, by = c('Taxon_code'))

```

**Take a look at the wrangled data**
```{r}

summary(mar_sus_gf) # should be no NA's in Sust_gf_taxon column
table(mar_sus_gf$gapfill)

```

Obtain a sustainability score for each record, and a book-keeping column of whether it's actual or gap-filled

For missing sustainability scores:

1. Use species-level Sustainability score
2. If no species-level scores, gapfill with taxon-level sustainability average

```{r}

mar_sus_final = mar_sus_gf %>% 
  mutate(gapfill = ifelse(!is.na(Sust), "none", gapfill)) %>%
  mutate(Sust = ifelse(is.na(Sust), Sust_gf_sp, Sust)) %>% # gapfill with species level
  mutate(gapfill = ifelse(is.na(Sust), "taxon_average", gapfill)) %>% # add in taxon gapfill record
  mutate(Sust = ifelse(is.na(Sust), Sust_gf_taxon, Sust)) %>% # gapfill with taxon level
  mutate(taxa_code = paste(species, species_code, sep="_")) %>%
  select(rgn_id, species, species_code, taxa_code, taxa_group=Taxon_code, year, gapfill_sus = gapfill, gapfill_fao = gap_0_fill, tonnes, Sust)

```

# Save Data:

```{r}

## Save seaweed mariculture harvest tonnes data ("tonnes" column already incorporated include proportions)
seaweed_harvest_tonnes <- mar_sus_final %>%
  dplyr::select(rgn_id, taxa_code, year, tonnes)

anyDuplicated(seaweed_harvest_tonnes) # check for duplication

write.csv(seaweed_harvest_tonnes, 'int/np_seaweed_harvest_tonnes.csv', row.names=F)

## Multiply tonnes by sustainability coefficient to get sustainably harvested tonnes and save
seaweed_harvest_tonnes_sust <- mar_sus_final %>% 
  mutate(tonnes_sust = tonnes*Sust) %>% 
  select(rgn_id, taxa_code, year, tonnes_sust)

write.csv(seaweed_harvest_tonnes_sust, 'int/np_seaweed_harvest_tonnes_sust.csv', row.names=F)

```

# Summarize values

Summarize the sustainably harvested tonnes per region per year. And, do some error checking.

```{r}
seaweed_harvest_tonnes_sust <- read_csv("int/np_seaweed_harvest_tonnes_sust.csv")

seaweed_sust_sum <- seaweed_harvest_tonnes_sust %>%
  group_by(rgn_id, year) %>%
  summarise(tonnes_sust = sum(tonnes_sust, na.rm = TRUE))
# no change from seaweed_harvest_tonnes_sust

## Error-checking and table exports to see if there are duplicates
stopifnot(sum(duplicated(seaweed_sust_sum[ , c('rgn_id', 'taxa_code', 'year')])) == 0)

```

# Calculate Rolling Averages
Determine rolling averages for sustainably harvested tonnes in order to determine peak values.

```{r}

# Find max year in the summarized data table
year_max <- max(seaweed_harvest_tonnes_sust$year)

roll_tonnes_sust <- seaweed_harvest_tonnes_sust %>%
  arrange(rgn_id, year) %>%
  group_by(rgn_id,) %>%
  mutate(tonnes_rollmean = rollapply(tonnes_sust, width=4, FUN=mean, align='right', partial=TRUE, na.rm=FALSE)) %>%
  rename(tonnes_orig = tonnes_sust) %>% # prevent overwriting of reported and gapfilled values
  mutate(tonnes = ifelse(!is.na(tonnes_rollmean), tonnes_rollmean, tonnes_orig)) %>%
  select(rgn_id, year, tonnes, tonnes_orig)

```

# Score Harvest Relative to Peaks

Score harvest (tonnes) relative to peaks. Output values as .csvs. Perform this using a for loop.

```{r}

#buffer  <-  0.35 # 35% buffer (from OHI Methods) ## commented out because we excluded this for v2020... 

## Find peak harvest per region-product and apply conservative buffer (scale down)
peak_tonnes_sust <- roll_tonnes_sust %>%
    group_by(rgn_id) %>%
    #mutate(tonnes_peak = max(tonnes, na.rm=T)  * (1 - buffer)) %>%
    mutate(tonnes_peak = max(tonnes, na.rm = T)) %>%
    ungroup() 

## Determine relative status:
smooth_tonnes_sust <- peak_tonnes_sust %>% 
    mutate(tonnes_rel = ifelse(tonnes >= tonnes_peak, 1, tonnes / tonnes_peak))
  
```

# Save data layer

```{r}

## Write entire data frame to .csv:
write.csv(smooth_tonnes_sust, 'int/np_seaweed_harvest_smoothed_data.csv', row.names = FALSE, na = '')

smooth_tonnes_sust <- read_csv("int/np_seaweed_harvest_smoothed_data.csv")
## Save scores data for the seaweed layer 
tonnes_seaweed_rel <- smooth_tonnes_sust %>%
  dplyr::select(rgn_id, year, score = tonnes_rel)

write.csv(tonnes_seaweed_rel, 'output/np_seaweed_scores.csv', row.names = FALSE, na = '')

```

