---
title: "OHI 2020: Food Provision/Fisheries, Identify FAO and OHI regions for RAM stocks"
author: "*Compiled on `r date()` by `r Sys.info()['user']`*"
output: 
  html_document:
    code_folding: show
    toc: true
    toc_depth: 1
    toc_float: yes
    number_sections: false
    theme: cerulean
    highlight: haddock
    includes: 
      in_header: '../../../workflow/templates/ohi_hdr.html'
pdf_document:
  toc: true
editor_options: 
  chunk_output_type: console
---

```{r}
knitr::opts_chunk$set(echo = TRUE, eval=FALSE)
```

## Summary

Generates a dataset describing the OHI and FAO major fishing regions that correspond to each stock's range.  This combines last year's dataset of stock-spatial information from RAM v4.44 database and Christopher M. Free's spatial work at Rutgers as well as newly added stocks from RAM v4.491 with manually assigned ohi and fao region ids. Dataset only contains stocks with B/Bmsy data.


## Setup

```{r}

## Libraries
library(dplyr)
library(tidyr)
library(readr)
library(sf)
library(ggplot2)
library(stringr)
library(here) 
setwd(here::here("globalprep","fis","v2020"))

source('../../../workflow/R/common.R')

```

## Identify Newly Added Stocks

Compare stocks found in previous version of RAM data to new data to see what additional stocks have been added.

Old Stocks: Used `timeseries` table, because in RAM v3.80, this table only contained single assessment for each unique stock
```{r}

## old stocks
load(file.path(dir_M, "git-annex/globalprep/_raw_data/RAM/d2019/RLSADB v4.44/DB Files With Assessment Data/DBdata.RData"))

ram_bmsy_old <- timeseries_values_views %>%
dplyr::select(stockid, stocklong, year, TBdivTBmsy, SSBdivSSBmsy, TBdivTBmgt, SSBdivSSBmgt) %>%
  mutate(ram_bmsy = 
           ifelse(!is.na(TBdivTBmsy), TBdivTBmsy, SSBdivSSBmsy)) %>%
  mutate(ram_bmsy =
           ifelse(is.na(TBdivTBmsy) & is.na(SSBdivSSBmsy), TBdivTBmgt, ram_bmsy)) %>%
  mutate(ram_bmsy = 
           ifelse(is.na(TBdivTBmsy) & is.na(SSBdivSSBmsy) & is.na(TBdivTBmgt), SSBdivSSBmgt, ram_bmsy)) %>%
  dplyr::filter(year > 1979) %>%
  filter(!is.na(ram_bmsy)) %>%
  dplyr::select(stockid, stocklong, year, ram_bmsy)

## check number of unique entries
old_stockid <- unique(ram_bmsy_old$stockid) # 442 unique entries

```

New Stocks: Used `timeseries_values_views` table, because in RAM v4.44, this is the table that contains the most recent assessment for each unique stock. The `timeseries` table has all assessments conducted so there are multiple `assessid` per unique `stockid`.
```{r}
## new stocks
load(file.path(dir_M, "git-annex/globalprep/_raw_data/RAM/d2020/RAMLDB v4.491/DB Files With Assessment Data/R Data/DBdata[asmt][v4.491].RData"))

ram_bmsy <- timeseries_values_views %>%
dplyr::select(stockid, stocklong, year, TBdivTBmsy, SSBdivSSBmsy, TBdivTBmgt, SSBdivSSBmgt) %>%
  mutate(ram_bmsy = 
           ifelse(!is.na(TBdivTBmsy), TBdivTBmsy, SSBdivSSBmsy)) %>%
  mutate(ram_bmsy =
           ifelse(is.na(TBdivTBmsy) & is.na(SSBdivSSBmsy), TBdivTBmgt, ram_bmsy)) %>%
  mutate(ram_bmsy = 
           ifelse(is.na(TBdivTBmsy) & is.na(SSBdivSSBmsy) & is.na(TBdivTBmgt), SSBdivSSBmgt, ram_bmsy)) %>% 
  filter(!is.na(ram_bmsy)) %>% 
  dplyr::filter(year > 1979) %>%
  dplyr::select(stockid, stocklong, year, ram_bmsy)

## check number of unique entries
new_stockid <- unique(ram_bmsy$stockid) # 443 unique entries
```


### Investigate Differences
There were 4 stocks included in the new data but not the old data.
These include: 

* "BTSHRIMPNAUST" "COD1f-XIV"     "GTPRAWNNAUST"  "SARDWSE"      

There were 3 stocks present in the old data but not the new.
These include:

* "CODIf-XIV"               "REDDEEPDP-1-2-V-XII-XIV" "RSARDINWA"    - in OLD RAM 4.44

```{r}

## find newly added stocks, when using timeseries_values_views
newStocks <- setdiff(new_stockid, old_stockid) # there are 4 new stocks
unusedStocks <- setdiff(old_stockid, new_stockid) # these are no longer included in the RAM data, will be deleted

```


Subset for just the additional stocks in RAM v4.491 data
```{r}

ram_bmsy_new <- ram_bmsy %>% 
  filter(stockid %in% newStocks)

id_new <- ram_bmsy_new %>% 
  select(stockid, stocklong) %>% 
  distinct()

```


## Add Stock Info to Old RAM_fao_ohi_rgns

Each stock needs the corresponding ohi and fao regions associated with its range.  For the most part, we can use the data from last year, and add the data for the new regions.  However, the old data uses "assessid", but we now need to replace this with the stock id and stock long because assessid is no longer used in the new version of the data.  The stock long and stock id  is obtained from the old RAM B/Bmsy data and merged with the RAM_fao_ohi_rgn data (done below).
```{r}
## Grab last year's fao-ohi-assessid data table
## (we need to get the stock id from the old RAM data)
RAM_fao_ohi_rgns <- read.csv("../v2019/int/RAM_fao_ohi_rgns_final.csv") 

## create unique values of assessid, stock id, and stock long
id_old <- ram_bmsy_old %>% 
  select(stockid, stocklong) %>% 
  distinct()

## join tables
RAM_rgns_old <- RAM_fao_ohi_rgns %>% 
  full_join(id_old, by=c("stockid", "stocklong")) 

```


## remove stocks that are no longer in the RAM data
(but were in the previous version)

```{r}
filter(RAM_rgns_old, stockid %in% unusedStocks) # quick check

RAM_rgns_old <- RAM_rgns_old %>%
  filter(!(stockid %in% unusedStocks))
```

## Combine New and Old Stock Data

Essentially adds new stocks to the bottom of the old stocks data table.  The regions of the new stocks are blank at this point. 
```{r}

## Make sure these new species aren't in the old data
setdiff(id_new$stockid, RAM_rgns_old$stockid)


RAM_rgns_new <- RAM_rgns_old %>% 
  full_join(id_new, by = c("stockid","stocklong"))

```

## Fill Region IDs for New Stocks

Investigate what FAO/OHI region each stock belongs to: 
* Full intermediate stock data table [here](https://github.com/OHI-Science/ohiprep_v2020/master/globalprep/fis/v2019/int/RAM_fao_ohi_rgns.csv)
* Primarily used www.fishsource.org to find stock distribution information. ICES sub-region codes found here: https://www.researchgate.net/figure/Stock-units-defined-by-ICES-for-white-anglerfish-southern-stock-in-Divisions-VIIIc-and_fig1_31198841 and http://gis.ices.dk/sf/
* Referenced F_CODE/FAO_AREAS (filtered for MAJOR in F_LEVEL) in ohi_fao_rgns shapefile (fis/v2017/int) from mazu (can use other map visuals to identify stock extent). 
* Searched and tested key words in data table viewer (`DT::datatable`) to use in `str_detect`. See which stocks have the same FAO id assignment(s).

Some regions have more than 1 `fao_id` or `rgn_id` (e.g. Alaska skate Bering Sea and Aleutian Islands,	Kamchatka flounder Bering Sea and Aleutian Islands have `fao_id` 61 and 67).

Note: In `str_detect`, it doesn't always work when you type out "Bering Sea | Gulf of Alaska$". Using `paste` with the `collapse = "|"` argument seems to work best.

```{r}

# Atlantic cod NAFO 1f and ICES 14 has ohi rgn *145* and fao rgn 21, 27 - -
# Brown tiger shrimp Northern Australia fao rgn 57, 71, and ohi rgn *16* - - 
# Grooved tiger prawn Northern Australia fao rgn 57, 71, and ohi rgn *16*- - 
# Norway lobster Adriatic Sea fao rgn 37, ohi rgn *184*, *187* - 
# Sand eel Sandeel Area 4 fao rgn 27, ohi rgn *223* - - 
# Sardine Western half of Southeast Australia ohi rgn *16*, fao rgn 57 - - 
# Spurdog Northeast Atlantic fao rgn 21, ohi rgn *163* - - 
# 	Pacific Ocean perch Haida Gwaii is in fao rgn 67, ohi rgn *218* - -
# Orange roughy Northwest Chatham Rise is in fao rgn 81, ohi rgn *162* - - 


## Subset for rows with missing FAO and OHI ids
RAM_filt <- RAM_rgns_new %>% 
  filter(is.na(fao_id))

RAM <- RAM_filt %>%
  ## ADDING FAO REGION ID
  mutate(fao_id = 
           case_when(
             str_detect(stocklong, paste(c("Atlantic cod", "Sand eel", "Boarfish VI"), collapse = "|")) ~ 27,
             str_detect(stocklong, paste(c("Northern Australia", "Southeast Australia"), collapse = "|")) ~ 57,
             str_detect(stocklong, paste(c("Northwest Chatham Rise"), collapse = "|")) ~ 81, 
             str_detect(stocklong, paste(c("Haida Gwaii"), collapse = "|")) ~ 67,
             str_detect(stocklong, paste(c("Spurdog"), collapse = "|")) ~ 21,
             str_detect(stocklong, paste(c("Norway lobster"), collapse = "|")) ~ 37
           )) %>%
  mutate(fao_id_2 =
           case_when(
             str_detect(stocklong, paste(c("Atlantic cod"), collapse = 
                        "|")) ~ 21,
             str_detect(stocklong, paste(c("Northern Australia"), collapse = "|")) ~ 71
           )) %>%
  ## ADDING OHI REGION ID
  mutate(rgn_id = 
           case_when(
             str_detect(stocklong, paste(c("Atlantic cod"), collapse = "|")) ~ 145,
             str_detect(stocklong, paste("Northwest Chatham Rise", collapse = "|")) ~ 162,
             str_detect(stocklong, paste(c("Haida Gwaii"), collapse = "|")) ~ 218, 
             str_detect(stocklong, paste(c("Spurdog"), collapse = "|")) ~ 163,
             str_detect(stocklong, paste(c("Australia"), collapse = "|")) ~ 16, 
             str_detect(stocklong, paste(c("Sandeel"), collapse = "|")) ~ 223,
             str_detect(stocklong, paste(c("Norway"), collapse = "|")) ~ 184
             
           )) %>%
  mutate(rgn_id_2 =
           case_when(
             str_detect(stocklong, paste(c("Norway"), collapse = "|")) ~ 187
           ))
             
DT::datatable(RAM,rownames = F)

```

## Tidy Final Datatable

### Gather FAO and OHI ids into long format

```{r}

RAM_temp <- RAM %>% 
  gather(fao, fao_values, contains("fao_id")) %>% 
  rename(fao_id = fao_values) %>% 
  select(-fao) %>% 
  gather(ohi, ohi_values, contains("rgn_id")) %>% 
  rename(rgn_id = ohi_values) %>% 
  select(-ohi) %>% 
  filter(!is.na(fao_id), !is.na(rgn_id)) %>%
  arrange(stockid, fao_id, rgn_id)

```

### Fix Incorrect FAO-OHI Matches

Each row must have both an FAO id and a rgn id. Fix duplicate/incorrect fao-ohi matches by using ohi_fao_rgns spatial file
```{r}

## Spatial file with fao and ohi regions, F_CODE is the FAO id
fao_ohi <- st_read(dsn = file.path(dir_M, "git-annex/globalprep/fis/v2017/int"),
                   layer = "ohi_fao_rgns")
st_geometry(fao_ohi) <- NULL # removes geometry

fao_ohi_id <- fao_ohi %>%
  select(rgn_id, fao_id = F_CODE) %>% 
  arrange(rgn_id) %>% 
  mutate(fao_id = as.numeric(as.character(fao_id))) %>% 
  mutate(rgn_id = as.numeric(as.character(rgn_id))) %>% 
  distinct()

## Filter for correct fao-ohi pairs in the RAM regions table 
RAM_rgns_new_final <- fao_ohi_id %>% 
  left_join(RAM_temp, by = c("rgn_id", "fao_id")) %>% 
  filter(!is.na(stockid)) %>% # keep matches only
 arrange(stocklong, fao_id)


```

Compare with unfiltered RAM regions table to check on fao-ohi pairs that were dropped
```{r}
## Number of unique stocks after joining
nrow(RAM_rgns_new_final) #13
## Number of unique stocks before joining
RAM_temp <- RAM_temp %>% select(rgn_id, fao_id, stockid, stocklong) %>% distinct() %>% arrange(stocklong, fao_id)
nrow(RAM_temp) #13

## add identifier for the two RAM table versions
RAM_temp$idtemp <- 1:nrow(RAM_temp)
RAM_rgns_new_final$idtemp2 <- 1:nrow(RAM_rgns_new_final)

## view the ohi-fao pairs that would be removed
combine <- RAM_temp %>% 
  full_join(RAM_rgns_new_final, by = c("rgn_id", "fao_id", "stockid", "stocklong")) %>% 
  filter(is.na(idtemp2)) ## 0 for 2020

```

Tidy up RAM data table.  After the final check (located below), I hand added some new regions for some of the stocks. 
```{r}

write.csv(RAM_rgns_new_final, "int/RAM_new_stocks.csv", row.names=FALSE)
## check that there are still 9 unique stock ids
length(unique(RAM_rgns_new_final$stockid))

```

Get RAM_area_m2 for new stocks 
Here we identify the FAO/OHI regions where each new RAM stock is located, and grab their RAM_area_m2. This involved the following steps:

1. Create a intersection map that identifies each FAO/OHI region.
2. Overlay each of the new RAM stocks on the region map to determine where they fall.  There were 2 general challenges to doing this.  A few of the stocks did not have a complete dataframe in the shapefile and could not be joined to the other data.  A few stocks had intersecting polygons and consequently would not run.  In the first case, I added these data by hand later in the process.  For the second case, I fixed the intersections and added later.  

```{r RAM to OHI_FAO rgn, eval = FALSE}

ram_bmsy_gf <- read_csv(file.path("int/ram_stock_bmsy_gf.csv")) %>%
  left_join(metadata, by = "stockid") %>%
  dplyr::select(assessid,stockid, year, ram_bmsy, gapfilled, method)


RAM_spatial_dir <- file.path(dir_M, "git-annex/globalprep/_raw_data/RAM/d2017/ramldb_boundaries/ramldb_boundaries")

ram_sf <- list.files(RAM_spatial_dir, pattern = "shp")

ram_sf <- gsub(".shp", "", ram_sf)

ram_sf <- intersect(ram_bmsy_gf$assessid, ram_sf) # get these spatial data

## get a sample file to convert other spatial data
tmp <- read_sf(dsn = RAM_spatial_dir, layer = ram_sf[1])

## Overlay stock data on fao-ohi region data

fao_ohi <- st_read(file.path(dir_M, "git-annex/globalprep/fis/v2017/int"),
        layer = "ohi_fao_rgns")

ram_sf_df <- data.frame(ram_sf, stringsAsFactors=FALSE) # make ram_sf a dataframe so that we can filter for new stocks added from management values...

new_species_list <- read_csv("int/RAM_new_stocks.csv")

ram_sf_df_new <- ram_sf_df %>%
  filter(grepl(paste(new_species_list$stockid, collapse = "|"), ram_sf)) ## 2020 - 2 of the new stocks for this year are located in Chris Free's RAM LDB boundaries 
ram_sf_df_new_vec <- c(ram_sf_df_new$ram_sf)


stock_fao_ohi <- NA
for(stock in ram_sf_df_new_vec) {    #stock = ram_sf_df_mgt$ram_sf[50]
    cat(paste0(stock, "\n"))
    tmp_poly <- read_sf(dsn = RAM_spatial_dir, layer = stock)
    tmp_fao_ohi <- st_intersection(fao_ohi, tmp_poly)
    if(sum(is.na(stock_fao_ohi))==1){
      stock_fao_ohi <- tmp_fao_ohi
    }else
    {stock_fao_ohi <- rbind(stock_fao_ohi, tmp_fao_ohi)}
}



stock_fao_ohi_shps <- stock_fao_ohi[st_dimension(stock_fao_ohi) == 2,]

stock_fao_ohi_shps <- stock_fao_ohi_shps[!is.na(st_dimension(stock_fao_ohi_shps)), ] 

## Fix GEOMETRYCOLLECTION features, http://r-spatial.org/r/2017/03/19/invalid.html

type <- st_is(stock_fao_ohi_shps, "GEOMETRYCOLLECTION")

stock_fao_ohi_shps[type, ] <- st_buffer(stock_fao_ohi_shps[type, ], 0.0)

#check:
type <- st_is(stock_fao_ohi_shps, "GEOMETRYCOLLECTION")


## Get areas in case we want to later weight the data
stock_fao_ohi_shps$RAM_area_m2 <- st_area(stock_fao_ohi_shps)

st_geometry(stock_fao_ohi_shps) <- NULL


stock_fao_ohi_new <- stock_fao_ohi_shps %>%
  dplyr::select(type_w_ant, rgn_ant_id, F_CODE, assessid, RAM_area_m2) %>%
  group_by(rgn_ant_id, F_CODE, assessid) %>%
  summarize(RAM_area_m2 = sum(RAM_area_m2)) %>%
  rename(rgn_id = rgn_ant_id, fao_id = F_CODE) %>%
  ungroup()

## Need to wrangle this so that we have stockid and stocklong (from 2020 ram data v4.491) instead of assessid (from 2017 ram data v3.80). Then we will join our new stocks with these so that we can get an estimate of area! 
stock_fao_ohi_new$fao_id <- as.numeric(as.character(stock_fao_ohi_new$fao_id))

new_species_list <- read_csv("int/RAM_new_stocks.csv")

stock_fao_ohi_new_final <- stock_fao_ohi_new %>%
  mutate(stockid = gsub("-1.*", "", assessid)) %>% #remove everything after "-1"
  mutate(stockid = gsub("-2.*", "", stockid)) %>% #remove everything after "-2"
  mutate(stockid = gsub("^[^-]*-","",stockid)) %>% #remove everything before the first "-"
  left_join(new_species_list, by = c("stockid", "rgn_id", "fao_id")) %>% # join on stockid to get the stocklong names we need...
 dplyr::select(-assessid, -idtemp2, -RAM_area_m2.y) %>% #delete assessid
  rename("RAM_area_m2" = "RAM_area_m2.x") %>% # keep only the rgn id and fao id we determined earlier for the new stocks present
ungroup()

write.csv(stock_fao_ohi_new_final, "int/RAM_fao_ohi_rgns_new_areas.csv", row.names=FALSE)

## Now add these to RAM_fao_ohi_rgns.csv file that you created in fao_ohi_rgns.Rmd

RAM_fao_ohi_rgns <- read_csv("int/RAM_fao_ohi_rgns.csv")

### Now we need to manually add the rest of the new species by looking each of them up individually and creating a spreadsheet with columns rgn_id, fao_id, RAM_area_m2 (these will all be NA), stockid, and stocklong. Make this csv file and upload to mazu. The code below will help as a check to make sure the rgn ids are in the correct fao rgn.

## Spatial file with fao and ohi regions, F_CODE is the FAO id
fao_ohi <- st_read(dsn = file.path(dir_M, "git-annex/globalprep/fis/v2017/int"),
                   layer = "ohi_fao_rgns")
st_geometry(fao_ohi) <- NULL # removes geometry
fao_ohi_id <- fao_ohi %>%
  select(rgn_id, fao_id = F_CODE) %>%
  arrange(rgn_id) %>%
  mutate(fao_id = as.numeric(as.character(fao_id))) %>%
  mutate(rgn_id = as.numeric(as.character(rgn_id))) %>%
  distinct()

manual_additions <- data.frame(stockid = sort(setdiff(new_species_list$stockid, stock_fao_ohi_new_final$stockid))) %>%
  left_join(new_species_list, by = "stockid")


write_csv(manual_additions, file.path("int/manual_new_area_additions.csv"))

## after manually adding the rest (just changing column names to match above, and adding NAs for RAM_area_m2), reupload into "int" folder as manual_new_area_additions_corr.csv

manual_additions_corr <- read_csv("int/manual_new_area_additions_corr.csv")

## add the manual additions areas to our new areas FAO_OHI rgns dataset

RAM_fao_ohi_rgns_new <- rbind(stock_fao_ohi_new_final, manual_additions_corr)

## write this to int folder
write_csv(RAM_fao_ohi_rgns_new, file.path("int/RAM_fao_ohi_rgns_new_areas_final.csv"))
```

### Final check
The goal here is to make sure we are capturing all the ohi/fao regions that the stock is caught in. I'm just going focus on the newly added stocks. After examining this table, I added in some more regions relevant to particular stocks, and saved the changes as "RAM_new_stocks_hand_additions.csv" (and then reran through the code with the new file.).
```{r, eval=FALSE}

ram <- read.csv("int/RAM_fao_ohi_rgns_new_areas_final.csv")

ram$idtemp2 <- 1:nrow(ram)

ram_sp <- ram %>%
  left_join(data.frame(metadata), by = "stockid") %>%
  dplyr::select(rgn_id, fao_id, stockid, stocklong=stocklong.x, scientificname, commonname, idtemp2)

# Watson species, sci name (read in the datatable that includes TaxonKey)
wat_sp <- read_csv(file.path(dir_M,'git-annex/globalprep/fis/v2020/int/stock_catch_by_rgn_taxa.csv')) %>% 
  dplyr:: select(rgn_id, fao_id=fao_rgn, scientificname=TaxonName, stock_id) %>%
  unique() 

wat_sp_w_RAM <- filter(wat_sp, scientificname %in% ram_sp$scientificname)

combined <- left_join(wat_sp_w_RAM, ram_sp, by=c("rgn_id", "fao_id", "scientificname")) %>%
  arrange(scientificname, stockid, fao_id, rgn_id,)

write.csv(combined, "int/watson_RAM_rgn_compare.csv", row.names=FALSE)

combined <- read_csv(file.path("int/watson_RAM_rgn_compare.csv"))


#### The way to hand check these are to look at stock_ids which have an idtemp2. I.e. Pandalus_borealis_27, filter for this stock_id, and hand check the rgn_id's that have NAs as stockids to determine if these rgn_id's should be hand added to our RAM_new_stocks according to the maps on fishsource.org

#2020
## need to hand add 176, 177, 179, 182, 183, 223, 227, 228 for boarfish -	Capros_aper-27
## need to hand add 144 to PANDALI - Pandalus_borealis-27
## need to hand add 219 to PANDALI - Pandalus_borealis-21
## need to hand add 55, 59, 143, 175, 176, 177, 178, 179, 222, 223, 227, 228 for 	PLAICCELT - European Plaice ICES VIIf-g


```

### Combine with full dataset

Combine newly added stocks with ohi and fao region information to the old stock data table. Make sure there are no NAs!
```{r, eval=FALSE}

RAM_final <- bind_rows(read_csv("int/RAM_new_stocks.csv"),
                       read_csv("int/RAM_new_stocks_hand_additions.csv", col_types = "ddccd")) %>% 
  select(-idtemp2)

RAM_fao_ohi_rgn <- RAM_rgns_old %>% 
  full_join(RAM_final, by = c("rgn_id", "fao_id", "stockid", "stocklong")) %>%
  select(-RAM_area_m2.y, RAM_area_m2 = RAM_area_m2.x)


write.csv(RAM_fao_ohi_rgn, "int/RAM_fao_ohi_rgns.csv", row.names=FALSE)


#check again last year 
#old_2018_RAM_fao_ohi_rgns <- read_csv("../v2018/int/RAM_fao_ohi_rgns.csv")
```

